---
title: "ASUbuildR"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    resize_reload: false
runtime: shiny
editor_options: 
  markdown: 
    wrap: 72
---


<style>
.dataTables_wrapper {
  width: fit-content !important;
  max-width: 100%;
}

.dataTables_scroll {
  width: fit-content !important;
  max-width: 100%;
}

.dataTables_scrollBody {
  height: auto !important;
  max-width: 100%;
}

table.dataTable {
  width: auto !important;
}

table.dataTable th,
table.dataTable td {
  white-space: nowrap;
}
</style>

<script>
(function registerAsuTooltipState() {
  if (!window.Shiny || !window.HTMLWidgets) {
    window.setTimeout(registerAsuTooltipState, 50);
    return;
  }
  if (window._asuTooltipStateHandlerInstalled) return;
  window._asuTooltipStateHandlerInstalled = true;
  window._asuTooltipState = window._asuTooltipState || {};
  window._asuTooltipInstallTimers = window._asuTooltipInstallTimers || {};

  function escapeHtml(value) {
    return String(value == null ? "" : value)
      .replace(/&/g, "&amp;")
      .replace(/</g, "&lt;")
      .replace(/>/g, "&gt;")
      .replace(/"/g, "&quot;")
      .replace(/'/g, "&#039;");
  }

  function formatNumber(value, digits) {
    var number = Number(value);
    return Number.isFinite(number)
      ? number.toLocaleString(undefined, { maximumFractionDigits: digits })
      : "";
  }

  function installTooltip(mapId) {
    var widget = window.HTMLWidgets.find("#" + mapId);
    var map = widget && widget.getMap ? widget.getMap() : null;
    if (!map || !map.getLayer || !map.getLayer("basemap")) {
      return false;
    }
    if (map._asuDynamicTooltipInstalled) return true;
    map._asuDynamicTooltipInstalled = true;

    var popup = new maplibregl.Popup({
      closeButton: false,
      closeOnClick: false,
      maxWidth: "400px"
    });
    map.on("mousemove", "basemap", function(e) {
      if (!e.features || !e.features.length) return;
      var props = e.features[0].properties || {};
      var geoid = String(props.GEOID == null ? "" : props.GEOID);
      var state = window._asuTooltipState[mapId] || {};
      var asunum = Object.prototype.hasOwnProperty.call(state, geoid)
        ? state[geoid]
        : (props.asunum == null ? 0 : props.asunum);
      var html =
        "<strong>ASU Number: </strong>" + escapeHtml(asunum) + "<br>" +
        "<strong>Tract GEOID: </strong>" + escapeHtml(geoid) + "<br>" +
        "<strong>Tract Population: </strong>" + formatNumber(props.tract_pop_cur, 0) + "<br>" +
        "<strong>Tract Labor Force: </strong>" + formatNumber(props.tract_ASU_clf, 0) + "<br>" +
        "<strong>Tract Unemployment Rate: </strong>" + formatNumber(props.tract_ASU_urate, 2) + "%<br>" +
        "<strong>Tract Unemployment: </strong>" + formatNumber(props.tract_ASU_unemp, 0) + "<br>";
      map.getCanvas().style.cursor = "pointer";
      popup.setLngLat(e.lngLat).setHTML(html).addTo(map);
    });
    map.on("mouseleave", "basemap", function() {
      map.getCanvas().style.cursor = "";
      popup.remove();
    });
    return true;
  }

  function ensureTooltip(mapId) {
    if (installTooltip(mapId) || window._asuTooltipInstallTimers[mapId]) return;
    window._asuTooltipInstallTimers[mapId] = window.setInterval(function() {
      if (installTooltip(mapId)) {
        window.clearInterval(window._asuTooltipInstallTimers[mapId]);
        delete window._asuTooltipInstallTimers[mapId];
      }
    }, 500);
  }

  window.Shiny.addCustomMessageHandler("asu-tooltip-state", function(message) {
    var mapId = message.id;
    var state = window._asuTooltipState[mapId] ||
      (window._asuTooltipState[mapId] = {});
    (message.updates || []).forEach(function(update) {
      state[String(update.geoid)] = update.asunum;
    });
    ensureTooltip(mapId);
  });
  // A state switch replaces the widget after its assignment message may have
  // arrived. Attach the tooltip to the new map as soon as its layers are ready.
  window.jQuery(document).on("shiny:value.asuTooltip", function(event) {
    if (event.name === "initial_map" || event.name === "edit_map") {
      window.setTimeout(function() { ensureTooltip(event.name); }, 0);
    }
  });
})();
</script>

```{r python_preamble, include=FALSE}
library(reticulate)
library(processx)

# Always define this first so options() won't error
python_ready <- FALSE
env_name <- "asu-cpsat"

# Try to activate the conda env and check for OR-Tools, but never abort knitting
try({
  envs_df <- reticulate::conda_list()
  if (env_name %in% envs_df$name) {
    # Force this interpreter via RETICULATE_PYTHON so a project-local .venv/renv
    # (auto-detected by reticulate) cannot silently override use_condaenv(),
    # even when use_condaenv() is called with required = TRUE.
    py_path <- envs_df$python[envs_df$name == env_name][1]
    Sys.setenv(RETICULATE_PYTHON = py_path)
    reticulate::use_condaenv(env_name, required = FALSE)

    if (reticulate::py_module_available("ortools")) {
      python_ready <- TRUE
      message("[env] Python environment ready")
    } else {
      message("[env] Conda env found, but 'ortools' is missing")
    }
  } else {
    message("[env] Python environment not set up; run ASUbuildR::setup_asu_python()")
  }
}, silent = TRUE)

# Store the flag for later chunks
options(asu_python_ready = python_ready)
```

```{r setup, include=FALSE}
# Figure out full paths to src/ C++ files, going up two levels
src_cpp_files <- c("choose_best_neighbor.cpp",
                   "build_edges.cpp",
                   "choose_best_drop.cpp")

src_paths <- normalizePath(file.path("..", "..", "src", src_cpp_files),
                           winslash = "/", mustWork = FALSE)

# Check that all source files are there
in_repo <- all(file.exists(src_paths))

maybe_sourceCpp <- function(fun, path) {
  if (!exists(fun, mode = "function") && file.exists(path)) {
    message("Compiling ", basename(path))
    Rcpp::sourceCpp(path)
  }
}

if (in_repo) {
  message("→ Dev mode: compiling C++ helpers from src/")
  maybe_sourceCpp("choose_best_neighbor", src_paths[1])
  maybe_sourceCpp("build_edges",          src_paths[2])
  maybe_sourceCpp("choose_best_drop",     src_paths[3])
} else {
  message("→ Installed package mode (shared library already loaded)")

}

# ---- dev mode: source helper functions relative to this file ----------
is_dev_mode <- file.exists("../../DESCRIPTION") && dir.exists("../../R")

if (is_dev_mode) {
  message("→ Dev mode: sourcing from ../../R/")
  r_dir <- normalizePath("../../R", winslash = "/", mustWork = TRUE)

  source(file.path(r_dir, "run_tract_hunter.R"))
  source(file.path(r_dir, "run_asu_original.r"))
  source(file.path(r_dir,"setup_asu_python.R"))
  source(file.path(r_dir, "read_asu_warm_start.R"))
  source(file.path(r_dir, "asu_summary.R"))
  source(file.path(r_dir, "neighbor_cache.R"))
  source(file.path(r_dir, "browser_data.R"))
  source(file.path(r_dir, "solver_checkpoint.R"))
  source(file.path(r_dir, "solver_jobs.R"))
  source(file.path(r_dir, "tract_geography.R"))
} else {
  asu_read_warm_start <- getFromNamespace("asu_read_warm_start", "ASUbuildR")
  asu_summarize_tracts <- getFromNamespace("asu_summarize_tracts", "ASUbuildR")
  asu_cached_neighbors <- getFromNamespace("asu_cached_neighbors", "ASUbuildR")
  for (helper in c("asu_load_tract_geography", "asu_join_tract_data", "asu_read_upload", "asu_state_slice", "asu_display_geometry", "asu_display_bounds",
                   "asu_payload_log", "asu_install_payload_logging", "asu_browser_source", "asu_map_source",
                   "asu_checkpoint_run_dir", "asu_service_checkpoints", "asu_job_root", "asu_job_run_dir",
                   "asu_job_prepare", "asu_job_launch", "asu_job_handle", "asu_job_list", "asu_job_cleanup",
                   "asu_job_state", "asu_job_recover", "asu_job_read"))
    assign(helper, getFromNamespace(helper, "ASUbuildR"))
  message("→ Installed package mode: functions should already be available")
}

`%||%` <- function(a, b) if (!is.null(a) && length(a)) a else b


# ── Use conda environment & import Python module ───────────────────
# This is best-effort: if Python/Conda or the 'asu-cpsat' env is missing,
# the dashboard must still start and load correctly (only the CP-SAT
# algorithm option becomes unavailable). Never let this stop the document.
asu_py <- NULL
if (isTRUE(python_ready)) {
  tryCatch({
    reticulate::use_condaenv(env_name, required = TRUE)
    py_config()  # should now show the env python

    # Import the bundled solver module
    asu_py <- import_from_path(
      module  = "asu_cpsat",
      path    = system.file("python", package = "ASUbuildR"),
      convert = TRUE
    )
    stopifnot(py_has_attr(asu_py, "build_many_asus_cpsat"))
  }, error = function(e) {
    message("[env] Could not load the CP-SAT Python module: ", conditionMessage(e))
    python_ready <<- FALSE
    asu_py <<- NULL
  })
}
options(asu_python_ready = python_ready)


## ---- tigris cache ----------------------------------------------------
## put cached TIGER/Line ZIPs in a per-user cache folder that is always
## writable, even after the package is installed read-only.

# 1. choose sensible user-cache paths (rappdirs works everywhere)
cache_dir <- rappdirs::user_cache_dir("ASUbuildR", "tigris")

# 2. create it if it doesn't exist
if (!dir.exists(cache_dir))
  dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

options(
  tigris_cache_dir = cache_dir,
  tigris_use_cache = TRUE,
  tigris_class     = "sf",
  download.file.method = "libcurl"

)

options(shiny.maxRequestSize = 1000 * 1024^2)  # Set to 50MB or your desired limit

# Record serialized output sizes, never dataset contents.
asu_install_payload_logging(session)

# Initialize reactive values
uploaded_data <- shiny::reactiveVal(NULL)
display_state <- shiny::reactiveVal(NULL)
upload_status <- shiny::reactiveVal("Choose an ASU workbook to begin.")
state <- shiny::reactiveVal(NULL)
tract_year <- shiny::reactiveVal(NULL)
tigris_year <- shiny::reactiveVal(NULL)
tract_data <- shiny::reactiveVal(NULL)
asu_data <- shiny::reactiveVal(NULL)
asu_tracts <- shiny::reactiveVal(NULL)
asu_summary <- shiny::reactiveVal(NULL)
full_data <- shiny::reactiveVal(NULL)
full_data_reset <- shiny::reactiveVal()
selected_tracts <- shiny::reactiveVal(NULL)
nb_adjacency    <- shiny::reactiveVal(NULL) # island-fixed full-state neighbor graph
# New England Optional year selection logic
output$is_new_england <- shiny::reactive({
  if (is.null(state())) {
    return(FALSE)
  }

  # New England state FIPS codes
  new_england_fips <- c("09", "23", "25", "33", "44", "50") # CT, ME, MA, NH, RI, VT

  any(state() %in% new_england_fips)
})

# ---- CP-SAT logging + paths ----
cpsat_log_buf    <- shiny::reactiveVal("")
cpsat_solver_proc<- shiny::reactiveVal(NULL)
cpsat_run_dir    <- shiny::reactiveVal(NULL)
cpsat_df_csv     <- shiny::reactiveVal(NULL)
cpsat_nb_json    <- shiny::reactiveVal(NULL)
cpsat_out_json   <- shiny::reactiveVal(NULL)
cpsat_progress_json <- shiny::reactiveVal(NULL)
cpsat_log_file   <- shiny::reactiveVal(NULL)
cpsat_log_saved  <- shiny::reactiveVal(NULL)
cpsat_stop_file  <- shiny::reactiveVal(NULL)
cpsat_stop_requested <- shiny::reactiveVal(FALSE)
cpsat_skip_file  <- shiny::reactiveVal(NULL)
cpsat_running <- shiny::reactiveVal(FALSE)
cpsat_stage_text <- shiny::reactiveVal("idle")
cpsat_active_strategy <- shiny::reactiveVal(NULL)

append_cpsat_log <- function(..., .persist = TRUE) {
  raw <- paste0(..., collapse = "")
  lines <- strsplit(raw, "\n", fixed = TRUE)[[1]]
  stage_lines <- grep("\\[STAGE\\]", lines, value = TRUE)
  if (length(stage_lines) > 0) {
    latest_stage <- stage_lines[length(stage_lines)]
    latest_stage <- sub("^.*\\[STAGE\\]\\s*", "", latest_stage)
    cpsat_stage_text(trimws(latest_stage))
  }

  line <- paste0(format(Sys.time(), "%H:%M:%S "), raw)
  old  <- cpsat_log_buf()
  new  <- if (nzchar(old)) paste(old, line, sep = "\n") else line
  if (nchar(new) > 12000) new <- substr(new, nchar(new) - 11999, nchar(new))
  cpsat_log_buf(new)
  lf <- cpsat_log_file()
  if (.persist && !is.null(lf)) try(cat(line, "\n", file = file.path(dirname(lf), "events.log"), append = TRUE), silent = TRUE)
}

# keep log output live even if hidden
output$cpsat_log_tail <- shiny::renderText(cpsat_log_buf())
shiny::outputOptions(output, "cpsat_log_tail", suspendWhenHidden = FALSE)
output$cpsat_stage <- shiny::renderText(cpsat_stage_text())
shiny::outputOptions(output, "cpsat_stage", suspendWhenHidden = FALSE)
output$show_cpsat_log_flag <- shiny::renderText({
  if (isTRUE(cpsat_running())) "true" else "false"
})
shiny::outputOptions(output, "show_cpsat_log_flag", suspendWhenHidden = FALSE)

# Disconnect detaches only the viewer; the supervisor and solver own their files.
session$onSessionEnded(function() {
  folder <- shiny::isolate(cpsat_run_dir())
  if (!is.null(folder)) try(cat(format(Sys.time(), "%Y-%m-%d %H:%M:%S %z"),
    " [session] Browser session ended; viewer did not terminate the job.\n",
    file = file.path(folder, "events.log"), append = TRUE), silent = TRUE)
})

# ---- New England year override logic ---------------------------------
# Make it available to the UI
shiny::outputOptions(output, "is_new_england", suspendWhenHidden = FALSE)

# Tract Hunter staged state -----------------------------------------
th_state <- shiny::reactiveVal(NULL)
th_tract_list <- shiny::reactiveVal(NULL)
th_bls_df <- shiny::reactiveVal(NULL)

# Nationwide data remains authoritative; the display selector changes the view only.
available_display_states <- shiny::reactive({
  data <- if (!is.null(full_data())) full_data() else uploaded_data()
  states <- sort(unique(substr(as.character(data$GEOID), 1L, 2L)))
  choices <- stats::setNames(states, states)
  if (length(states) > 1L) choices <- c(choices, "All states (nationwide)" = "ALL")
  choices
})
shiny::observe({
  choices <- available_display_states()
  if (!length(choices)) return()
  chosen <- shiny::isolate(display_state())
  if (is.null(chosen) || !chosen %in% choices) {
    chosen <- unname(choices[1L])
    display_state(chosen)
  }
  for (id in c("display_state_upload", "display_state_initial", "display_state_edit"))
    shiny::updateSelectInput(session, id, choices = choices, selected = chosen)
})
for (picker in c("display_state_upload", "display_state_initial", "display_state_edit")) local({
  id <- picker
  shiny::observeEvent(input[[id]], {
    chosen <- input[[id]]
    if (length(chosen) == 1L && chosen %in% available_display_states() &&
        !identical(chosen, display_state())) {
      display_state(chosen)
      selected_tracts(NULL)
      highlighted_tracts(character(0))
      shiny::updateSelectInput(session, "select_asunum", selected = "None")
      for (other in setdiff(c("display_state_upload", "display_state_initial", "display_state_edit"), id))
        shiny::updateSelectInput(session, other, selected = chosen)
    }
  }, ignoreInit = TRUE)
})
display_data <- shiny::reactive({
  shiny::req(full_data(), display_state())
  asu_state_slice(full_data(), display_state())
})

# Invalidate geometry only when its actual content/attributes change, never on
# assignment edits. State switches and same-GEOID replacement geometry are safe.
map_revision <- shiny::reactiveVal(0L)
map_signature <- NULL
geoid_set <- shiny::reactiveVal(character(0))
shiny::observe({
  if (is.null(full_data())) {
    if (!is.null(map_signature)) {
      map_signature <<- NULL
      geoid_set(character(0))
      map_revision(shiny::isolate(map_revision()) + 1L)
    }
    return()
  }
  data <- display_data()
  signature <- data[, c("GEOID", "tract_pop_cur", "tract_ASU_clf",
                        "tract_ASU_urate", "tract_ASU_unemp")]
  if (!identical(map_signature, signature)) {
    map_signature <<- signature
    geoid_set(sort(data$GEOID))
    map_revision(shiny::isolate(map_revision()) + 1L)
  }
})
# Retain the last two display geometries so state/nation round trips avoid
# repeating nationwide simplification. Exact GEOIDs and geometry invalidate it.
display_geometry_cache <- list()
base_geom <- shiny::reactive({
  map_revision()
  data <- dplyr::select(shiny::isolate(display_data()), GEOID)
  area <- shiny::isolate(display_state())
  cached <- display_geometry_cache[[area]]
  if (!is.null(cached) && identical(cached$input, data)) {
    display_geometry_cache <<- c(display_geometry_cache[area],
      display_geometry_cache[setdiff(names(display_geometry_cache), area)])
    return(cached$output)
  }
  geometry <- if (identical(area, "ALL")) {
    shiny::withProgress(message = "Preparing all-state map", detail = "Simplifying display geometry", value = 0.1,
      asu_display_geometry(data))
  } else asu_display_geometry(data)
  display_geometry_cache <<- c(stats::setNames(list(list(input=data, output=geometry)), area),
    display_geometry_cache[setdiff(names(display_geometry_cache), area)])
  display_geometry_cache <<- head(display_geometry_cache, 2L)
  geometry
})
map_geom <- shiny::reactive({
  attrs <- display_data() |>
    sf::st_drop_geometry() |>
    dplyr::select(GEOID, asunum, tract_pop_cur, tract_ASU_clf, tract_ASU_urate, tract_ASU_unemp)
  dplyr::inner_join(base_geom(), attrs, by = "GEOID")
})
map_source <- shiny::reactive({
  revision <- map_revision()
  data <- shiny::isolate(map_geom())
  shiny::req(nrow(data) > 0L)
  national <- identical(shiny::isolate(display_state()), "ALL")
  result <- tryCatch({
    if (national) shiny::withProgress(message = "Preparing all-state map",
      detail = "Building or loading cached PMTiles; the first load may take several minutes", value = 0.5,
      asu_map_source(data, session, revision, require_tiles = TRUE))
    else asu_map_source(data, session, revision)
  }, error = identity)
  shiny::validate(shiny::need(!inherits(result, "error"),
    if (inherits(result, "error")) conditionMessage(result) else ""))
  result
})

highlighted_tracts <- shiny::reactiveVal(character(0))

# Keep the shared ASU summary synchronized with the authoritative assignment
# table. This covers solver completion, manual edits, reset, and RDS loading.
shiny::observeEvent(full_data(), {
  current_summary <- asu_summarize_tracts(full_data(), by_asu = TRUE) |>
    dplyr::select(-`Unemployment Rate Exact`)
  asu_summary(current_summary)
}, ignoreNULL = TRUE)

output$asu <- shiny::renderTable({
  shiny::req(asu_summary())
  asu_summary()
}, digits = 3)

# Load static data
excel_names <- c("record",
                 "geoid",
                 "st_fips",
                 "cnty_fips",
                 "tract_fips",
                 "name",
                 "tract_pop_dec",
                 "tract_pop_cur",
                 "tract_emp",
                 "tract_unemp",
                 "tract_urate",
                 "tract_urate_error",
                 "cnty_pop_dec",
                 "cnty_pop_cur",
                 "cnty_emp",
                 "cnty_unemp",
                 "cnty_urate",
                 "cnty_urate_error",
                 "pop_shr",
                 "emp_shr",
                 "unemp_shr",
                 "laus_primary",
                 "cnty_emp_ave",
                 "cnty_unemp_ave",
                 "tract_ASU_clf",
                 "tract_ASU_emp",
                 "tract_ASU_unemp",
                 "tract_ASU_urate")

#-------- Map Rendering Helpers --------------------------------
# Stable ASU colors. Gray is reserved for unassigned tracts, so do not add
# neutral colors to this palette. The first seven entries preserve the colors
# previously supplied by RColorBrewer's Accent palette; its gray eighth entry
# is intentionally replaced with teal.
ASU_COLORS <- c(
  "#7FC97F", "#BEAED4", "#FDC086", "#FFFF99", "#386CB0",
  "#F0027F", "#BF5B17", "#00A6A6", "#E41A1C", "#984EA3",
  "#4DAF4A", "#FF7F00", "#A65628", "#1F78B4", "#E7298A",
  "#66A61E", "#7570B3", "#D95F02", "#1B9E77", "#C51B7D"
)

asu_palette <- function(n) {
  n <- max(0L, as.integer(n))
  if (n == 0L) character(0) else rep_len(ASU_COLORS, n)
}

# Helper function to calculate color scheme (used for full layer rebuilds)
calculate_color_scheme <- function(data) {
  number_asus <- max(data$asunum, na.rm = TRUE)
  number_categories <- min(c(number_asus, 8))
  
  if (number_categories == 1) {
    asu_breaks <- 1
    actual_categories <- 1
    brewer_pal <- asu_palette(1L)
  } else {
    asu_breaks <- seq(from = 1, to = number_asus, length.out = number_categories)
    actual_categories <- max(3, number_categories)
    full_brewer_pal <- asu_palette(actual_categories)
    brewer_pal <- full_brewer_pal[1:number_categories]
  }
  
  return(list(breaks = asu_breaks, colors = brewer_pal))
}

# Build a per-GEOID -> color lookup from current asunum assignments.
# Used by update_map_efficiently() so the paint expression is keyed on GEOID
# (a stable feature property that never changes after the initial source load)
# rather than on asunum (which IS stale in the cached MapLibre source).
calculate_tract_colors <- function(data) {
  n_asu <- suppressWarnings(max(data$asunum, na.rm = TRUE))
  n_asu <- if (is.finite(n_asu) && n_asu > 0L) as.integer(n_asu) else 1L

  pal <- asu_palette(n_asu)

  # pal_ext[1] = unassigned gray (asunum 0); the remaining colors are ASUs 1..n.
  pal_ext <- c("#E0E0E0", pal)
  idx <- as.integer(data$asunum) + 1L   # asunum 0 -> index 1, asunum 1 -> index 2, …
  idx[is.na(idx) | idx < 1L | idx > length(pal_ext)] <- 1L

  list(
    values = as.character(data$GEOID),
    stops  = pal_ext[idx]
  )
}

tract_tooltip <- function() {
  mapgl::concat(
    "<strong>ASU Number: </strong>",
    mapgl::get_column("asunum"), "<br>",
    "<strong>Tract GEOID: </strong>", mapgl::get_column("GEOID"), "<br>",
    "<strong>Tract Population: </strong>", mapgl::number_format(mapgl::get_column("tract_pop_cur"), style = "decimal", maximum_fraction_digits = 0), "<br>",
    "<strong>Tract Labor Force: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_clf"), style = "decimal", maximum_fraction_digits = 0), "<br>",
    "<strong>Tract Unemployment Rate: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_urate"), style = "decimal", maximum_fraction_digits = 2), "%<br>",
    "<strong>Tract Unemployment: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_unemp"), style = "decimal", maximum_fraction_digits = 0), "<br>"
  )
}

# Keep dynamic ASU values in a small browser-side lookup used by both map
# tooltips. Only GEOIDs whose assignment changed are sent to the browser.
asu_tooltip_assignments <- shiny::reactiveVal(
  stats::setNames(integer(0), character(0))
)

push_asu_tooltip_updates <- function(data) {
  if (is.null(data) || nrow(data) == 0 ||
      !all(c("GEOID", "asunum") %in% names(data))) return(invisible(FALSE))

  data <- asu_state_slice(data, display_state())
  if (!nrow(data)) return(invisible(FALSE))
  geoids <- as.character(data$GEOID)
  values <- suppressWarnings(as.integer(data$asunum))
  values[is.na(values) | values < 0L] <- 0L
  current <- stats::setNames(values, geoids)
  previous <- asu_tooltip_assignments()
  old_values <- unname(previous[geoids])
  changed <- is.na(old_values) | old_values != values

  if (any(changed)) {
    changed_idx <- which(changed)
    updates <- lapply(changed_idx, function(i) {
      list(geoid = geoids[i], asunum = unname(values[i]))
    })
    for (map_id in c("initial_map", "edit_map")) {
      session$sendCustomMessage(
        "asu-tooltip-state",
        list(id = map_id, updates = updates)
      )
    }
  }
  asu_tooltip_assignments(current)
  invisible(TRUE)
}

shiny::observeEvent(list(full_data(), display_state()), {
  push_asu_tooltip_updates(full_data())
}, ignoreNULL = TRUE)

# Helper: set filter + visibility on the 'selected' or 'highlighted' overlay layer.
# In mapgl 0.5.0, set_filter() composes user+base via ["all", ...], so we MUST
# NOT set an initial filter on these layers. Use visibility="none" as the off-state
# and pair every set_filter call with a matching set_layout_property visibility toggle.
apply_overlay_filter <- function(proxy, layer_id, ids) {
  ids <- ids[!is.na(ids) & (display_state() == "ALL" | substr(ids, 1L, 2L) == display_state())]
  if (length(ids) == 0) {
    proxy |>
      mapgl::set_filter(layer_id, NULL) |>
      mapgl::set_layout_property(layer_id, "visibility", "none")
  } else {
    filter_expr <- if (length(ids) == 1) {
      list("==", list("get", "GEOID"), as.character(ids))
    } else {
      list("in", list("get", "GEOID"), list("literal", as.character(ids)))
    }
    proxy |>
      mapgl::set_filter(layer_id, filter_expr) |>
      mapgl::set_layout_property(layer_id, "visibility", "visible")
  }
}

update_map_efficiently <- function(map_id = "edit_map", paint_only = FALSE) {
  data <- map_geom()
  if (is.null(data) || nrow(data) == 0) return()

  cm <- calculate_tract_colors(data)

  sel <- selected_tracts()   %||% character(0)
  hl  <- highlighted_tracts() %||% character(0)

  proxy <- mapgl::maplibre_proxy(map_id)


  proxy <- proxy |>
    mapgl::set_paint_property(
      "basemap", "fill-color",
      mapgl::match_expr(
        column  = "GEOID",
        values  = cm$values,
        stops   = cm$stops,
        default = "#E0E0E0"
      )
    )

  if (map_id == "edit_map") {
    proxy <- apply_overlay_filter(proxy, "selected",    sel)
    proxy <- apply_overlay_filter(proxy, "highlighted", hl)
  }
  proxy
}

# Geometry changes rebuild the shared HTTP source through map_revision().
update_map_with_data <- function(map_id = "edit_map", new_data = NULL) {
  update_map_efficiently(map_id, paint_only = TRUE)
}

fix_nb_islands <- function(nb, intpt_lat, intpt_lon) {
  g     <- igraph::graph_from_adj_list(nb)
  comps <- igraph::components(g)
  if (comps$no <= 1L) return(nb)
  main_id  <- which.max(comps$csize)
  main_idx <- which(comps$membership == main_id)
  for (comp_id in setdiff(unique(comps$membership), main_id)) {
    island_idx <- which(comps$membership == comp_id)
    combis     <- expand.grid(island = island_idx, main = main_idx)
    combis$dist <- sqrt(
      (intpt_lat[combis$island] - intpt_lat[combis$main])^2 +
      (intpt_lon[combis$island] - intpt_lon[combis$main])^2
    )
    closest <- combis[which.min(combis$dist), ]
    nb[[closest$island]] <- c(nb[[closest$island]], closest$main)
    nb[[closest$main]]   <- c(nb[[closest$main]],   closest$island)
  }
  nb
}

# Check whether asu_geoids form a single connected component in stored_nb.
asu_is_contiguous <- function(asu_geoids, stored_nb) {
  idx <- which(stored_nb$geoids %in% asu_geoids)
  if (length(idx) <= 1L) return(TRUE)
  local_nb <- lapply(seq_along(idx), function(i) {
    nbrs  <- stored_nb$nb[[idx[i]]]
    local <- which(idx %in% nbrs)
    if (length(local) == 0L) integer(0L) else as.integer(local)
  })
  g <- igraph::graph_from_adj_list(local_nb, mode = "all")
  igraph::components(g)$no == 1L
}

```

# Data Initialization
## Controls {data-width="150"}
This screen allows you to import the BLS-provided ASU file, generally named like _NV_asu23.xlsx_ which will be the only external data import step necessary. The script will read the file, extract the state, and determine the population year from the Excel file.

You can specify which year of Census tract boundaries to use with the selector below. This is important because Census tract boundaries may change between years, and you'll want to use the appropriate year that matches the GEOIDs in your BLS file.

If the necessary data is not in columns **A through AB** within the Excel file, this process will not work correctly.
Once the file is loaded, a preview of the data will appear on the right, the state and year information will display below, and the analyst may proceed to the next tab (Load Initial ASU).
Note: at this time, ASUbuildR assumes a state is using county-based census tracts. If a state uses another geography such as NECTAs as the basis for its ASU tract file, this process will require some modification.
```{r}
shiny::h4("Persistent solver jobs")
shiny::selectInput("saved_job", "Running job", choices = c("Choose a job" = ""))
shiny::actionButton("refresh_jobs", "Refresh jobs")
shiny::actionButton("attach_job", "Attach to job")
shiny::helpText("Jobs continue after this tab closes. Attach to monitor results or use Stop. Job files are kept on the machine running R.")
shiny::fileInput("file", "Choose Excel File", accept = c(".xls", ".xlsx"))
shiny::selectInput("display_state_upload", "Display area:", choices = character(0))
shiny::textOutput("upload_status")
shiny::helpText("Choose a state or All states (nationwide) for maps and tract tables. Display selection does not change solver scope.")
# Add numeric input for tract year
shiny::numericInput("tract_year_input",
                     "Census Tract Year:",
                     value = 2024,
                     min = 2010,
                     max = 2025,
                     step = 1)
shiny::helpText("Select the year of Census tract boundaries to use.
                 Connecticut workbooks using legacy county GEOIDs automatically use matching 2021 tract geometry.")
shiny::textOutput("selected_state")
shiny::textOutput("tract_pop_year")
shiny::textOutput("selected_tract_year")
```

## Output {data-width="850"}

```{r}
DT::DTOutput("data_preview")
output$upload_status <- shiny::renderText(upload_status())
shiny::observeEvent(input$file, {
  shiny::req(input$file)
  if (isTRUE(cpsat_running())) {
    shiny::showNotification("Stop the active solve before replacing the uploaded data.", type = "error")
    return()
  }
  upload_status("Reading and validating workbook...")
  result <- tryCatch(shiny::withProgress(message = "Loading ASU workbook", value = 0, {
    shiny::incProgress(0.2, detail = "Reading all tract rows")
    result <- asu_read_upload(input$file$datapath)
    shiny::incProgress(0.8, detail = "Validation complete")
    result
  }), error = function(e) {
    upload_status(paste("Upload rejected:", conditionMessage(e)))
    shiny::showNotification(conditionMessage(e), type = "error", duration = NULL)
    NULL
  })
  if (is.null(result)) return()
  # Publish atomically after validation; no expensive outputs see partial data.
  full_data(NULL)
  full_data_reset(NULL)
  nb_adjacency(NULL)
  selected_tracts(NULL)
  highlighted_tracts(character(0))
  th_state(NULL)
  th_tract_list(NULL)
  th_bls_df(NULL)
  asu_summary(NULL)
  uploaded_data(result$data)
  state(sort(unique(result$data$st_fips)))
  tract_year(result$population_year)
  display_state(state()[1L])
  upload_status(sprintf("Loaded %s tracts across %d states.",
    format(nrow(result$data), big.mark = ","), length(state())))
})
output$data_preview <- DT::renderDT({
  shiny::req(uploaded_data(), display_state())
  DT::datatable(asu_state_slice(uploaded_data(), display_state()), rownames = FALSE,
                options = list(pageLength = 10, scrollX = TRUE, deferRender = TRUE))
}, server = TRUE)
output$selected_state <- shiny::renderText({
  shiny::req(state(), display_state())
  paste0(if (identical(display_state(), "ALL")) "Display: All uploaded states" else paste0("Display FIPS: ", display_state()),
         "; solver scope: ", length(state()), " uploaded state(s)")
})
output$tract_pop_year <- shiny::renderText({
  paste0("Population year (detected from Excel): ", tract_year())
})

# Update tigris_year reactiveVal when input changes
shiny::observeEvent(input$tract_year_input, {
  yr <- suppressWarnings(as.integer(input$tract_year_input))
  if (is.na(yr) || yr < 2010 || yr > 2025) {
    output$selected_tract_year <- shiny::renderText({
      paste0("Invalid year entered (", input$tract_year_input, "); must be 2010-2025. Using previous value: ", tigris_year() %||% "not set")
    })
    return()
  }
  tigris_year(as.character(yr))
  output$selected_tract_year <- shiny::renderText({
    paste0("Census tract boundary year (user selected): ", tigris_year())
  })
}, ignoreInit = FALSE)
```

# Load Initial ASU

## Controls {data-width="200"}

```{r}
shiny::selectInput("display_state_initial", "Display area:", choices = character(0))
# ---- New England year override controls ------------------------------
shiny::conditionalPanel(
  condition = "output.is_new_england === true",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::h5("New England Year Override"),
    shiny::radioButtons(
      inputId = "ne_year_choice",
      label   = "Override tract year for New England:",
      choices = list(
        "Use selected year from Data Initialization" = "detected",
        "Use 2021 for New England states" = "2021"
      ),
      selected = "detected",
      inline = FALSE
    ),
    shiny::p(
      style = "font-size: 0.9em; color: #666;",
      "This override applies only to New England states. Connecticut legacy county GEOIDs automatically use 2021 geometry; other states use the selected year."
    )
  )
)

# ---- algorithm picker -------------------------------------------------
shiny::radioButtons(
  inputId = "asu_algo",
  label   = "Choose algorithm",
  choices = c(
    "Tract Hunter"      = "mine",
    "Simple Snake"      = "orig",
    "CP-SAT (OR-Tools)" = "cpsat"
  ),
  inline  = TRUE
)


# -----------------------------------------------------------------------
# main "Run" button (always visible)
shiny::div(style = "margin-top: 0.5rem;",
    shiny::actionButton(
      inputId = "process",
      label   = "Load tracts & initialise",
      class   = "btn-primary"
    )
)

# -----------------------------------------------------------------------
# extra Tract-Hunter controls (only when algo == 'mine')
shiny::conditionalPanel(
  condition = "input.asu_algo === 'mine'",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::div(
      class = "btn-group",           # keep them on one line
      style = "width: 100%;",
      shiny::actionButton("th_pass",    "Run Hunter",     class = "btn-secondary"),
      shiny::actionButton("th_combine", "Combine Groups",   class = "btn-secondary")
    )
  )
)

# Conditional panel for CP-SAT algorithm options
shiny::conditionalPanel(
  condition = "input.asu_algo === 'cpsat'",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::radioButtons(
      "cpsat_strategy",
      "CP-SAT strategy",
      choices = c(
        "Partitioning strategy (potential multiple ASUs)" = "partition",
        "Component-first global (automatic ASU count)" = "component_global",
        "Legacy single-ASU solve" = "single",
        "Split saved ASUs (joint model)" = "split"
      ),
      selected = "single"
    ),
    shiny::checkboxInput("cpsat_use_warm_start", "Use saved ASUs as a warm start", value = FALSE),
    shiny::conditionalPanel(
      condition = "input.cpsat_use_warm_start === true || input.cpsat_strategy === 'split'",
      shiny::fileInput("cpsat_warm_start", "Warm-start RDS", accept = ".rds", multiple = FALSE),
      shiny::helpText("Matches GEOID plus asunum/asu_id to the current input data. Saved groups must meet current thresholds and connectivity. Max ASUs includes these groups except in Component-first global, which has no group-count cap. Assignments remain changeable.")
    ),
    shiny::conditionalPanel(
      condition = "input.cpsat_strategy === 'split'",
      shiny::helpText("Requires a warm-start RDS regardless of the checkbox. Tries imported ASUs once, highest captured unemployment first, using articulation points and narrow graph corridors as split hints. Each child must contain a parent tract, qualify independently, and avoid touching all other ASUs. Unassigned tracts are available; only strict combined unemployment gains are accepted."),
      shiny::helpText("Increase Max ASUs above the imported count to leave room for children. Each split may create two or three children, limited further by the available slots. Solves run sequentially with all workers. Each starts with a cut pass: at most 100 rounds or 25 stalled-upper-bound rounds. Each quiet cut round waits for optimality or its five-second solver limit, unless interrupted; discovering a cut does not end the round. Time and incumbent-stall limits apply to the subsequent exact flow solve per parent. No merging or polishing follows.")
    ),
    shiny::conditionalPanel(
      condition = "input.cpsat_strategy === 'partition'",
      shiny::selectInput(
        "cpsat_partition_seed_strategy", "Initial ASU seed method",
        choices = c("Connectivity-free components" = "connectivity_free",
                    "Surplus pruning" = "surplus_prune"),
        selected = "connectivity_free"
      ),
      shiny::numericInput("cpsat_expansion_incumbent_stall_seconds",
                          "Expansion incumbent stall limit (sec)", value = 300, min = 0, step = 30),
      shiny::checkboxInput("cpsat_final_consolidation", "Consolidate touching ASUs at the end", value = TRUE),
      shiny::checkboxInput("cpsat_polish_consolidated_asus", "Polish consolidated groups", value = TRUE),
      shiny::helpText("Final consolidation preserves every selected tract. Optional polishing uses the final-polish budget and general stall limit, preserves selected tracts, then consolidates any new contacts. Stop still allows the quick consolidation but skips additional solves."),
      shiny::helpText("Finish retained-component repair and partition expansion solves after this many seconds without an incumbent improvement. 0 disables this limit; the solve time budget still applies. Final polishing uses the general stall limit."),
      shiny::helpText("Surplus pruning starts with all tracts and removes the lowest-surplus tracts while preserving population. A removal may split a component only if every resulting component meets the population threshold; each is then pruned independently toward the rate threshold. Stalled components are repaired using only retained tracts. Valid seeds then enter outward expansion and polishing, where pruned tracts become available again."),
      shiny::helpText("Partitioning expands each ASU individually, sequentially, using all workers and a 1800-second limit per ASU."),
      shiny::helpText("In partitioning, touching ASUs are jointly reoptimized with all reachable unassigned tracts instead of automatically merged. Roots and boundaries can move; the solver may keep separate groups or consolidate them. Combined unemployment cannot decrease. Graph cuts and a connectivity pre-pass are enabled automatically for these touching solves. Each touching-cluster solve uses all workers and the expansion/polish budget (1800 seconds in this dashboard); the pre-pass shares that budget.")
    ),
    shiny::numericInput("cpsat_tau",       "UR threshold (τ)",          value = 0.0645, min = 0, max = 1, step = 0.0001),
    shiny::numericInput("cpsat_pop",       "Population threshold",      value = 10000,  min = 0, step = 100),
    shiny::conditionalPanel(
      condition = "input.cpsat_strategy !== 'component_global'",
      shiny::numericInput("cpsat_max", "Max ASUs", value = 30, min = 1, step = 1),
      shiny::numericInput("cpsat_incumbent_stall_seconds", "Incumbent stall limit (sec)", value = 300, min = 0, step = 30),
      shiny::numericInput("cpsat_parallel_asus", "Concurrent ASU solves", value = 1, min = 1, step = 1),
      shiny::helpText("The solve time limit applies per main window, not to the whole run. Partition expansion and final polishing use their separate budgets.")
    ),
    shiny::numericInput("cpsat_timelimit", "Solve time limit (sec)", value = 18000, min = 0, step = 60),
    shiny::conditionalPanel(
      condition = "input.cpsat_strategy === 'component_global'",
      shiny::helpText("Experimental global selection: all tracts remain available and qualifying connected components become ASUs automatically. No Max ASUs cap, per-ASU concurrency, or incumbent-stall cutoff applies. The solve time limit is a whole-search budget including model preparation, cut generation, and local repair; input validation, shutdown, and final result writing can add overhead."),
      shiny::helpText("Batched component cuts, regional eligibility constraints, and the best certified objective ceiling strengthen later rounds. Short neighborhood repairs run periodically and after stalled improvement, then unrestricted global search resumes. No ASU is permanently frozen, and local repair bounds never establish global optimality."),
      shiny::helpText("Saved ASUs supply a valid starting solution and a soft hint, not fixed boundaries. Quiet cut rounds run for at most five solver seconds. The log reports validated captured unemployment and a global upper bound/gap. A time-limited result is not necessarily optimal, and speed is not guaranteed. Stop returns the best valid selection; Skip advances to the next cut round.")
    ),
    shiny::numericInput("cpsat_workers",   "Total CP-SAT workers (detected cores - 2)", value = max(1, parallel::detectCores(logical = FALSE)-2), min = 1, step = 1),
    shiny::numericInput("cpsat_relgap",        "Relative gap (optional)",    value = NULL, min = 0, step = 0.0001),
    shiny::uiOutput("cpsat_log_link")
  )
)



# ---- dynamic algorithm description -----------------------------------
output$algo_blurb <- shiny::renderUI({
  choice <- input$asu_algo %||% "orig"

  if (identical(choice, "cpsat")) {
    ## ───────── CP-SAT (OR-Tools) METHOD ─────────
    python_status <- if (isTRUE(getOption("asu_python_ready", FALSE))) {
      "<p style=\"color:#2e7d32;\"><strong>Python/OR-Tools environment detected.</strong></p>"
    } else {
      "<p style=\"color:#b71c1c;\"><strong>Python/OR-Tools environment not detected.</strong> Run <code>ASUbuildR::setup_asu_python()</code> in R, then restart the dashboard. Until then, use Tract Hunter or Simple Snake.</p>"
    }
    shiny::HTML(
      paste0(python_status, "<p><strong>CP-SAT (OR-Tools) workflow</strong></p>
       <p>
         This option builds ASUs by solving a mixed-integer model over tract
         adjacency. The app will:
       </p>
       <ul>
         <li>Download Census tract shapes for the selected year and build
             neighbour lists (contiguity).</li>
         <li>Merge the BLS upload you provided (population, labor force,
             employment, unemployment, rate).</li>
         <li>Run the CP-SAT solver with your parameters:
             <code>UR threshold (τ)</code>, <code>Population threshold</code>,
             <code>Max ASUs</code>, <code>Time limit</code>, optional
             <code>Relative gap</code>, <code>Total CP-SAT workers</code>, and
             <code>Concurrent ASU solves</code>. Component-first global instead
             chooses its own ASU count and uses one whole-search time budget.</li>
       </ul>
       <p>
         While solving, a live <em>CP-SAT log</em> panel appears (top-right)
         showing progress. Single-ASU and partition expansion incumbents are
         checked every minute. Component-first global publishes assignments
         only when validated captured unemployment improves. Subtle green/red fills show additions/removals
         relative to each solve's starting hint; these temporary previews clear
         before results are committed. When the solver finishes, the map and summary table
         update automatically; tracts with <code>asunum = 0</code> are not in
         any ASU.
       </p>
       <p>
         <strong>Requirements:</strong> a working Python environment with
         OR-Tools. If it isn't available, you'll see a notice and can use
         Tract Hunter or Simple Snake instead. Uploaded GEOIDs must match the
         tract geometry; legacy Connecticut county GEOIDs use 2021 geometry automatically.
       </p>
       <p>
         Click <em>Load tracts &amp; initialise</em> to start. You can refine
         results on the <em>Modify ASU Selections</em> tab once the map renders.
       </p>")
    )

  } else if (identical(choice, "mine")) {
    ## ───────── TRACT-HUNTER METHOD ─────────
    shiny::HTML(
      "<p><strong>Tract Hunter workflow</strong></p>
       <p>Because Tract Hunter works in stages, the interface exposes those
          stages as separate buttons:</p>
       <ul>
         <li><em>Load tracts &amp; initialise</em><br>
             ▸ downloads shapefiles, merges the BLS file, then runs the
             <strong>Seed + Expand</strong> step automatically. This should only be run once.</li>
         <li><em>Run Hunter</em><br>
             ▸ performs the optimisation/trade pass that adds
             high-unemployment neighbours and drops low-unemployment ones until
             every ASU is as close as possible to the
             <code>6.45%</code> threshold.</li>
         <li><em>Combine Groups</em><br>
             ▸ merges adjacent ASUs that share a boundary. Connected ASUs allow for more tracts to be added, so whenever ASUs are merged it is advisable to run another pass with the Tract Hunter.</li>
       </ul>
       <p>You can stop after any step to inspect the map, or run the buttons
          in sequence for the full, automatic workflow. We suggest iteratively running the Run Hunter and Combine Groups steps until Total Unemployment in ASUs no longer increases. Each step updates the
          colour map and the unemployment totals shown on the right.</p>"
    )

  } else {
    ## ───────── SIMPLE SNAKE (ORIGINAL) METHOD ─────────
    shiny::HTML(
      "<p>
        This button will launch two processes, which will take up to 5&nbsp;minutes
        to complete. For smaller states, it may take under 15&nbsp;seconds.
      </p>
      <p>
        First, the application will download census shapefiles by tract for the
        appropriate state and year. Next, after combining the BLS file with these
        shapefiles, it will iteratively search for contiguous geographic regions
        until the combined unemployment rate in each region is as close to
        6.45% as possible.
      </p>
      <p>
        When this process is complete, a map will appear to the right, and a
        summary of the created ASU regions will appear below. It is important to
        note these will still include areas that do not qualify as an independent
        ASU (there is no check for population), but it gives the analyst a
        starting place for analysis.
      </p>
      <p>
        Once this step is done, you may proceed to the next step.
      </p>"
    )
  }
})


shiny::uiOutput("algo_blurb")

shiny::htmlOutput("total_unemp")

shiny::tableOutput("asu")
#shiny::verbatimTextOutput("cpsat_log_tail")

palette_logic <- function(n_asu) {
  if (n_asu == 0) {
    list(
      fill = "#E0E0E0",
      legend_vals = "(none)",
      legend_cols = "#E0E0E0"
    )
  } else if (n_asu == 1) {
    list(
      fill = mapgl::step_expr(
        column = "asunum",
        base = "#E0E0E0",
        stops = asu_palette(1L),
        values = 1,
        na_color = "#E0E0E0"
      ),
      legend_vals = 1,
      legend_cols = asu_palette(1L)
    )
  } else {
    legend_vals <- seq_len(n_asu)
    legend_cols <- asu_palette(n_asu)

    list(
      fill = mapgl::match_expr(
        column = "asunum",
        values = legend_vals,
        stops = legend_cols
      ),
      legend_vals = legend_vals,
      legend_cols = legend_cols
    )
  }
}



# A viewer can attach after a browser disconnect or an R/dashboard restart.
attach_cpsat_job <- function(run_dir) {
  initial_sf <- readRDS(file.path(run_dir, "input.rds"))
  merged_df <- sf::st_drop_geometry(initial_sf)
  input_df <- utils::read.csv(file.path(run_dir, "df.csv"), colClasses = c(geoid = "character"))
  if (!identical(as.character(merged_df$GEOID), input_df$geoid)) stop("Job GEOID order mismatch")
  graph <- jsonlite::fromJSON(file.path(run_dir, "nb.json"), simplifyVector = FALSE)
  nb_adjacency(list(geoids = merged_df$GEOID,
                   nb = lapply(graph, function(v) as.integer(unlist(v)) + 1L)))
  full_data(initial_sf)
  full_data_reset(initial_sf)
  selected_tracts(NULL)
  highlighted_tracts(character(0))
  state(sort(unique(substr(initial_sf$GEOID, 1L, 2L))))
  if (is.null(display_state()) || !display_state() %in% c(state(), "ALL")) display_state(state()[1L])
  cpsat_log_buf("")
  cpsat_run_dir(run_dir)
  cpsat_active_strategy(asu_job_read(file.path(run_dir, "job.json"))$strategy)
  cpsat_df_csv(file.path(run_dir, "df.csv"))
  cpsat_nb_json(file.path(run_dir, "nb.json"))
  cpsat_out_json(file.path(run_dir, "out.json"))
  cpsat_progress_json(file.path(run_dir, "progress.json"))
  cpsat_log_file(file.path(run_dir, "solver.log"))
  cpsat_log_saved(file.path(run_dir, "solver.log"))
  cpsat_stop_file(file.path(run_dir, "stop.flag"))
  cpsat_skip_file(file.path(run_dir, "skip.flag"))
  cpsat_stop_requested(file.exists(cpsat_stop_file()))
  cpsat_solver_proc(asu_job_handle(run_dir))
  cpsat_running(TRUE)
  cpsat_stage_text(asu_job_state(run_dir)$status)
  append_cpsat_log(paste0("[session] Attached to persistent job: ", run_dir))
    apply_asu_assignments <- function(asu_values, update_reset = FALSE, map_only = FALSE) {
      asu_id_vals <- suppressWarnings(as.integer(asu_values))
      if (length(asu_id_vals) != nrow(merged_df) || any(is.na(asu_id_vals))) {
        return(FALSE)
      }

      if (isTRUE(map_only)) {
        color_df <- data.frame(
          GEOID = as.character(merged_df$GEOID),
          asunum = ifelse(asu_id_vals < 0L, 0L, as.integer(asu_id_vals))
        )
        push_asu_tooltip_updates(color_df)
        color_df <- asu_state_slice(color_df, display_state())
        cm <- calculate_tract_colors(color_df)
        fill_expr <- mapgl::match_expr(
          column  = "GEOID",
          values  = cm$values,
          stops   = cm$stops,
          default = "#E0E0E0"
        )
        for (map_id in c("initial_map", "edit_map")) {
          try(
            mapgl::maplibre_proxy(map_id) |>
              mapgl::set_paint_property("basemap", "fill-color", fill_expr),
            silent = TRUE
          )
        }
        return(TRUE)
      }

      # The persistent input already contains clean original geometry/columns.
      live_sf <- initial_sf |>
        dplyr::mutate(asunum = pmax(0L, asu_id_vals))

      full_data(live_sf)
      if (isTRUE(update_reset)) {
        full_data_reset(live_sf)
      }
      asu_summary(
        live_sf |>
          sf::st_drop_geometry() |>
          dplyr::filter(asunum > 0) |>
          dplyr::group_by(asunum) |>
          dplyr::summarise(
            Tracts              = dplyr::n(),
            Population          = sum(tract_pop_cur,   na.rm = TRUE),
            `Labor Force`       = sum(tract_ASU_clf,   na.rm = TRUE),
            Unemployment        = sum(tract_ASU_unemp, na.rm = TRUE),
            `Unemployment Rate` = round((Unemployment/`Labor Force`)*100, 5),
            .groups = "drop"
          )
      )
      output$asu <- shiny::renderTable(asu_summary(), digits = 3)
      update_map_efficiently("initial_map")
      update_map_efficiently("edit_map")
      TRUE
    }

    # The Python builder starts with -1 for every tract. Seed the comparison
    # with that same state so its INIT snapshot does not trigger a redundant
    # all-grey repaint. Later phase-only snapshots are skipped the same way.
    last_progress_asu <- shiny::reactiveVal(rep.int(-1L, nrow(merged_df)))
    last_exploring_geoids <- shiny::reactiveVal(character(0))
    last_exploring_candidate_geoids <- shiny::reactiveVal(character(0))
    last_exploring_added_geoids <- shiny::reactiveVal(character(0))
    last_exploring_removed_geoids <- shiny::reactiveVal(character(0))

    # stream logs and finalize when done
    if (!is.null(session$userData$cpsatObs)) {
      try(session$userData$cpsatObs$destroy(), silent = TRUE)
    }
    tmr <- shiny::reactiveTimer(1000)
    session$userData$cpsatObs <- shiny::observe({
      tmr()
      px <- cpsat_solver_proc()
      shiny::req(!is.null(px))

      out <- try(px$read_output_lines(), silent = TRUE)
      err <- try(px$read_error_lines(),  silent = TRUE)
      if (!inherits(out,"try-error") && length(out)) append_cpsat_log(paste(out, collapse="\n"), .persist = FALSE)
      if (!inherits(err,"try-error") && length(err)) append_cpsat_log(paste(err, collapse="\n"), .persist = FALSE)

      # Checkpoints are saved by independent R subprocesses, never by this viewer.

      pf <- cpsat_progress_json()
      if (!is.null(pf) && file.exists(pf)) {
        progress_payload <- try(jsonlite::fromJSON(pf), silent = TRUE)
        if (!inherits(progress_payload, "try-error") && !is.null(progress_payload$asu_id)) {
          progress_asu <- suppressWarnings(as.integer(progress_payload$asu_id))
          valid_progress <- length(progress_asu) == nrow(merged_df) &&
            !any(is.na(progress_asu))

          if (valid_progress && !identical(progress_asu, last_progress_asu())) {
            # The JSON is cheap to read on every tick. The assignment comparison
            # updates the map, committed data, and summary table only when tract
            # membership or compact ASU numbering actually changed. Expansion
            # results must be available for saving before polishing finishes.
            if (isTRUE(apply_asu_assignments(
              progress_asu,
              update_reset = FALSE,
              map_only = FALSE
            ))) {
              last_progress_asu(progress_asu)
            }
          }

          overlay_specs <- list(
            list(
              payload = "exploring_idx",
              layer = "exploring",
              previous = last_exploring_geoids
            ),
            list(
              payload = "exploring_candidate_idx",
              layer = "exploring_candidate",
              previous = last_exploring_candidate_geoids
            ),
            list(
              payload = "exploring_added_idx",
              layer = "exploring_added",
              previous = last_exploring_added_geoids
            ),
            list(
              payload = "exploring_removed_idx",
              layer = "exploring_removed",
              previous = last_exploring_removed_geoids
            )
          )
          for (overlay_spec in overlay_specs) {
            overlay_idx <- suppressWarnings(as.integer(
              progress_payload[[overlay_spec$payload]] %||% integer(0)
            ))
            # Suppress retained/candidate outlines, including older snapshots.
            if (overlay_spec$layer %in% c("exploring", "exploring_candidate")) {
              overlay_idx <- integer(0)
            }
            overlay_idx <- unique(overlay_idx[
              !is.na(overlay_idx) & overlay_idx >= 0L &
                overlay_idx < nrow(merged_df)
            ])
            overlay_geoids <- sort(as.character(
              merged_df$GEOID[overlay_idx + 1L]
            ))
            if (!identical(overlay_geoids, overlay_spec$previous())) {
              for (map_id in c("initial_map", "edit_map")) {
                try(
                  apply_overlay_filter(
                    mapgl::maplibre_proxy(map_id),
                    overlay_spec$layer,
                    overlay_geoids
                  ),
                  silent = TRUE
                )
              }
              overlay_spec$previous(overlay_geoids)
            }
          }
        }
      }

      if (!px$is_alive()) {
        for (layer_id in c(
          "exploring", "exploring_candidate",
          "exploring_added", "exploring_removed"
        )) {
          for (map_id in c("initial_map", "edit_map")) {
            try(
              apply_overlay_filter(
                mapgl::maplibre_proxy(map_id),
                layer_id,
                character(0)
              ),
              silent = TRUE
            )
          }
        }
        last_exploring_geoids(character(0))
        last_exploring_candidate_geoids(character(0))
        last_exploring_added_geoids(character(0))
        last_exploring_removed_geoids(character(0))

        # drain remaining
        out_all <- try(px$read_all_output_lines(), silent = TRUE)
        err_all <- try(px$read_all_error_lines(),  silent = TRUE)
        if (!inherits(out_all,"try-error") && length(out_all)) append_cpsat_log(paste(out_all, collapse="\n"), .persist = FALSE)
        if (!inherits(err_all,"try-error") && length(err_all)) append_cpsat_log(paste(err_all, collapse="\n"), .persist = FALSE)

        status <- px$get_exit_status()
        append_cpsat_log(paste0("[run] Exit status: ", status))
        cpsat_stage_text(asu_job_state(run_dir)$status)
        session$userData$cpsatObs$destroy()

        if (identical(status, 0L) && file.exists(cpsat_out_json())) {
          resj <- jsonlite::fromJSON(cpsat_out_json())
          if (!is.null(resj$asu_id) && length(resj$asu_id) == nrow(merged_df)) {
            if (isTRUE(apply_asu_assignments(resj$asu_id, update_reset = TRUE))) {
              append_cpsat_log("[run] Solver finished OK; map/table updated.")
              if (identical(cpsat_active_strategy(), "component_global")) {
                append_cpsat_log(sprintf(
                  "[STAGE] COMPONENT_GLOBAL_RESULT status=%s optimal=%s valid_unemp=%s upper_bound=%s absolute_gap=%s relative_gap=%s",
                  as.character(resj$status %||% "UNKNOWN"),
                  if (isTRUE(resj$optimal)) "true" else "false",
                  as.character(resj$total_unemp %||% "NA"),
                  as.character(resj$upper_bound %||% "NA"),
                  as.character(resj$absolute_gap %||% "NA"),
                  as.character(resj$relative_gap %||% "NA")))
              }
            } else {
              append_cpsat_log("[run][warn] Output JSON had invalid assignment shape.")
            }
          } else {
            append_cpsat_log("[run][warn] Output JSON invalid or length mismatch.")
          }
        } else {
          append_cpsat_log("[run][error] Solver failed or out.json missing.")
        }

        lf <- cpsat_log_file()
        if (!is.null(lf) && file.exists(lf)) cpsat_log_saved(lf)
        cpsat_log_file(NULL)
        sf <- cpsat_stop_file()
        # Keep stop.flag as part of the durable job history.
        cpsat_stop_file(NULL)
        cpsat_stop_requested(FALSE)
        kf <- cpsat_skip_file()
        # Keep any remaining skip flag for diagnostics.
        cpsat_skip_file(NULL)
        cpsat_progress_json(NULL)
        cpsat_running(FALSE)
      }
    })
}

shiny::observe({
  input$refresh_jobs
  shiny::invalidateLater(5000, session)
  choices <- asu_job_list()
  selected <- shiny::isolate(input$saved_job)
  if (length(selected) != 1L || !selected %in% unname(choices)) selected <- ""
  shiny::updateSelectInput(session, "saved_job",
    choices = c("Choose a job" = "", choices), selected = selected)
})
shiny::observeEvent(input$attach_job, {
  folder <- input$saved_job
  shiny::req(length(folder) == 1L, nzchar(folder))
  # Never accept arbitrary filesystem paths supplied by a browser client.
  if (!folder %in% unname(asu_job_list())) {
    shiny::showNotification("This job is no longer running. Refresh the job list.", type = "warning")
    return()
  }
  tryCatch(attach_cpsat_job(folder), error = function(e)
    shiny::showNotification(conditionMessage(e), type = "error", duration = NULL))
})

shiny::observeEvent(input$process, {
  ## 0) Basic checks ------------------------------------------------------
  shiny::req(uploaded_data(), state(), tigris_year())

  selected_year <- tigris_year()
  new_england_2021 <- identical(input$ne_year_choice, "2021")

  # ensure 'res' exists even if something returns early
  res <- NULL

  # Detect algorithm before withProgress to avoid overlapping progress bars
  algo_choice <- shiny::isolate(input$asu_algo)
  if (is.null(algo_choice) || identical(algo_choice, "")) algo_choice <- "orig"

  # --- CP-SAT branch (has its own withProgress) --------------------------
if (identical(algo_choice, "cpsat")) {
  jobs <- asu_job_list()
  if (any(vapply(unname(jobs), function(folder) isTRUE(asu_job_state(folder)$alive), logical(1)))) {
    shiny::showModal(shiny::modalDialog(title = "A solver job is already running",
      "Use Refresh jobs and Attach to job to monitor or stop it. Disconnecting does not stop a job.",
      easyClose = TRUE))
    return()
  }
  if (!isTRUE(getOption("asu_python_ready", FALSE))) {
    shiny::showModal(shiny::modalDialog(
      title = "Python Environment Not Ready",
      "The CP-SAT (OR-Tools) algorithm requires a configured Python environment. Run ASUbuildR::setup_asu_python() in R, then restart the dashboard. In the meantime, choose Tract Hunter or Simple Snake.",
      easyClose = TRUE
    ))
    return()
  }
  session <- shiny::getDefaultReactiveDomain()
  cpsat_running(TRUE)
  cpsat_stage_text("initializing run")
  append_cpsat_log("[run] Button clicked (CP-SAT).")
  # Let this observer finish its first reactive flush so the lightweight log
  # panel reaches the browser before TIGER downloads, geometry preparation, or
  # either MapLibre widget begins. Start the expensive work immediately after
  # that flush rather than holding the log until this whole branch yields.
  session$onFlushed(function() {
  shiny::isolate({
  shiny::withProgress(message = "Running OR-Tools CP-SAT…", value = 0, {
    # ---- 1) Preconditions & parameters --------------------------------
    shiny::req(uploaded_data(), state(), tract_year())
    sel_year <- selected_year
    st_fips  <- state()
    tau      <- if (is.null(input$cpsat_tau))       0.0645 else as.numeric(input$cpsat_tau)
    pop_thr  <- if (is.null(input$cpsat_pop))       10000  else as.integer(input$cpsat_pop)
    max_asus <- if (is.null(input$cpsat_max))       30     else as.integer(input$cpsat_max)
    tlimit   <- if (is.null(input$cpsat_timelimit)) 1200   else as.integer(input$cpsat_timelimit)
    incumbent_stall_seconds <- if (is.null(input$cpsat_incumbent_stall_seconds) || is.na(input$cpsat_incumbent_stall_seconds)) 300L else max(0L, as.integer(input$cpsat_incumbent_stall_seconds))
    workers  <- if (is.null(input$cpsat_workers))   max(1L, parallel::detectCores()-5L) else as.integer(input$cpsat_workers)
    parallel_asus <- if (is.null(input$cpsat_parallel_asus)) 1L else max(1L, as.integer(input$cpsat_parallel_asus))
    cpsat_strategy <- if (is.null(input$cpsat_strategy) || identical(input$cpsat_strategy, "")) "partition" else as.character(input$cpsat_strategy)
    use_partitioning <- identical(cpsat_strategy, "partition")
    use_split <- identical(cpsat_strategy, "split")
    use_component_global <- identical(cpsat_strategy, "component_global")
    if (use_component_global) {
      parallel_asus <- 1L
      incumbent_stall_seconds <- 0L
      append_cpsat_log("[run] component-global mode: automatic ASU count; whole-search time budget; no incumbent-stall cutoff")
    }
    partition_seed_strategy <- if (use_partitioning && identical(input$cpsat_partition_seed_strategy, "surplus_prune")) "surplus_prune" else "connectivity_free"
    expansion_incumbent_stall_seconds <- if (is.null(input$cpsat_expansion_incumbent_stall_seconds) || is.na(input$cpsat_expansion_incumbent_stall_seconds)) 300L else max(0L, as.integer(input$cpsat_expansion_incumbent_stall_seconds))
    if (use_partitioning) append_cpsat_log("[run] partition expansion mode: sequential")
    rel_gap  <- if (!is.null(input$cpsat_relgap) && !is.na(input$cpsat_relgap)) as.numeric(input$cpsat_relgap) else NA_real_
    append_cpsat_log(sprintf("[run] params: strategy=%s tau=%.6f pop=%d max=%s limit=%ds stall=%ds workers=%d concurrent_asus=%d rel_gap=%s",
           cpsat_strategy,
           tau, pop_thr, if (use_component_global) "automatic" else as.character(max_asus), tlimit, incumbent_stall_seconds, workers, parallel_asus,
                             if (is.na(rel_gap)) "None" else as.character(rel_gap)))
    # ---- 2) Build tracts + neighbors ----------------------------------
    shiny::incProgress(0.15, detail = "Downloading tracts & building neighbors")
    
    # 1. Collapse the vector of states into a single string for the log (e.g., "32, 06")
    st_fips_str <- paste(st_fips, collapse = ", ")
    append_cpsat_log(sprintf("[NB] Downloading TIGER tracts for state(s) %s (year=%s)…", st_fips_str, sel_year))
    
    # Validate coverage before calculating adjacency or launching Python.
    tr_sf <- tryCatch(asu_load_tract_geography(
      uploaded_data(), sel_year, new_england_2021,
      log = function(...) append_cpsat_log(paste0(...))), error = identity)
    if (inherits(tr_sf, "error")) {
      append_cpsat_log(paste0("[geography][error] ", conditionMessage(tr_sf)))
      shiny::showModal(shiny::modalDialog(title = "Tract geography mismatch",
        conditionMessage(tr_sf), easyClose = TRUE))
      cpsat_stage_text("geography validation failed")
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    tr_sf$continuous <- asu_cached_neighbors(tr_sf, {
      raw_nb <- lapply(sfdep::st_contiguity(sf::st_geometry(tr_sf)), function(v)
        if (length(v) == 1L && v == 0L) integer(0) else as.integer(v))
      fix_nb_islands(raw_nb, as.numeric(tr_sf$INTPTLAT), as.numeric(tr_sf$INTPTLON))
    }, mode = "queen-island-v1", log = function(...) append_cpsat_log(paste0(...)))
    
    # Join with uploaded BLS
    cols_needed <- c("GEOID","tract_ASU_unemp","tract_ASU_emp","tract_pop_cur","tract_ASU_clf","tract_ASU_urate")
    miss <- setdiff(cols_needed, names(uploaded_data()))
    if (length(miss)) {
      shiny::showModal(shiny::modalDialog(
        title="Input problem", paste("Uploaded data missing:", paste(miss, collapse=", ")), easyClose=TRUE
      ))
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    
    merged_sf <- tryCatch(asu_join_tract_data(tr_sf, uploaded_data()), error = identity)
    if (inherits(merged_sf, "error")) {
      append_cpsat_log(paste0("[input][error] ", conditionMessage(merged_sf)))
      shiny::showModal(shiny::modalDialog(title = "Input problem",
        conditionMessage(merged_sf), easyClose = TRUE))
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    
    # The cached graph already includes island connections.
    nb <- merged_sf$continuous
    nb_adjacency(list(geoids = merged_sf$GEOID, nb = nb))

    # Keep original geometry for saved results, contiguity and solver work.
    # Display simplification happens only for the selected display area.
    merged_df <- merged_sf |> sf::st_drop_geometry()
    append_cpsat_log(sprintf("[NB] Tracts kept: %d (after join)", nrow(merged_df)))

    warm_start <- NULL
    if (use_split || isTRUE(input$cpsat_use_warm_start)) {
      warm_start <- tryCatch({
        if (is.null(input$cpsat_warm_start)) {
          stop(if (use_split) "Split saved ASUs requires a warm-start RDS file." else "Choose a warm-start RDS file or disable warm start.")
        }
        imported <- asu_read_warm_start(
          input$cpsat_warm_start$datapath, merged_df$GEOID, data = merged_df,
          neighbors = lapply(merged_sf$continuous, function(iv) as.integer(iv) - 1L),
          tau = tau, pop_thresh = pop_thr,
          max_asus = if (use_component_global) NULL else max_asus)
        if (use_split && imported$asu_count < 1L) stop("Split saved ASUs requires at least one assigned ASU in the RDS file.")
        imported
      }, error = function(e) e)
      if (inherits(warm_start, "error")) {
        append_cpsat_log(paste0("[warm start] rejected: ", conditionMessage(warm_start)))
        shiny::showModal(shiny::modalDialog(title = "Warm-start problem",
                         conditionMessage(warm_start), easyClose = TRUE))
        cpsat_running(FALSE)
        return(invisible(NULL))
      }
      append_cpsat_log(sprintf("[STAGE] WARM_START loaded=%s asus=%d assigned_tracts=%d matched_tracts=%d ignored_unassigned=%d baseline_unemp=%.0f",
                              input$cpsat_warm_start$name, warm_start$asu_count,
                              warm_start$assigned_tracts, warm_start$matched_tracts,
                              warm_start$ignored_unassigned, warm_start$baseline_unemp))
    }

    clean_sf_columns <- function(x) {
      g <- sf::st_geometry(x)
      df <- x |>
        sf::st_drop_geometry() |>
        dplyr::select(where(~ !is.list(.)))
      sf::st_as_sf(df, geometry = g, crs = sf::st_crs(x))
    }

    # Initialize the stable map source before Python starts. Live progress can
    # then repaint this existing basemap without replacing its geometry or
    # rebuilding any layers.
    initial_sf <- merged_sf |>
      dplyr::mutate(asunum = if (is.null(warm_start)) 0L else pmax(0L, warm_start$ids)) |>
      dplyr::select(-dplyr::any_of("continuous")) |>
      clean_sf_columns()
    full_data(initial_sf)
    
    # Continue with the existing code...
    # Dataframe for Python
    df_py <- merged_df |>
      dplyr::transmute(
        geoid           = GEOID,
        tract_ASU_unemp = as.integer(round(tract_ASU_unemp)),
        tract_ASU_emp   = as.integer(round(tract_ASU_emp)),
        tract_pop2024   = as.integer(round(tract_pop_cur))
      )
    if (!is.null(warm_start)) df_py$warm_start_asu <- warm_start$ids
    
    # Use the island-fixed neighborhood lists for the Python solver
    nb_lists <- lapply(merged_sf$continuous, function(iv) {
      if (length(iv) == 0) list() else as.list(as.integer(iv) - 1L)
    })

    # ---- 3) Persistent job folder + files ------------------------------------
    shiny::incProgress(0.10, detail = "Writing inputs")
    legacy_checkpoint <- !use_partitioning && !use_split && !use_component_global
    run_dir <- asu_job_run_dir()
    dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

    df_csv   <- file.path(run_dir, "df.csv")
    nb_json  <- file.path(run_dir, "nb.json")
    out_json <- file.path(run_dir, "out.json")
    progress_json <- file.path(run_dir, "progress.json")
    runner   <- file.path(run_dir, "runner.py")
    stop_file <- file.path(run_dir, "stop.flag")
    skip_file <- file.path(run_dir, "skip.flag")

    utils::write.csv(df_py, df_csv, row.names = FALSE)
    jsonlite::write_json(nb_lists, nb_json, auto_unbox = TRUE)

    log_file <- file.path(run_dir, "solver.log")
    cpsat_run_dir(run_dir)
    cpsat_df_csv(df_csv)
    cpsat_nb_json(nb_json)
    cpsat_out_json(out_json)
    cpsat_progress_json(progress_json)
    cpsat_log_file(log_file)
    cpsat_log_saved(NULL)
    unlink(progress_json)
    unlink(stop_file)
    unlink(skip_file)
    cpsat_stop_file(stop_file)
    cpsat_stop_requested(FALSE)
    cpsat_skip_file(skip_file)

    append_cpsat_log(paste0("[run] Folder: ", run_dir))
    append_cpsat_log(paste0("[run] df.csv: ", df_csv))
    append_cpsat_log(paste0("[run] nb.json: ", nb_json))

    # ---- 4) Write runner.py -------------------------------------------
    shiny::incProgress(0.05, detail = "Preparing Python runner")
    py_mod_path <- normalizePath(system.file("python", package = "ASUbuildR"),
                                 winslash = "/", mustWork = TRUE)
    runner_code <- sprintf("
import sys, os, json, math
try:
    sys.stdout.reconfigure(line_buffering=True)
except Exception:
    pass
os.environ.setdefault('PYTHONUNBUFFERED','1')
os.environ.setdefault('OMP_NUM_THREADS','1')
os.environ.setdefault('MKL_NUM_THREADS','1')
os.environ.setdefault('OPENBLAS_NUM_THREADS','1')
print('[runner] starting'); print('Python:', sys.version.replace('\\n',' '))

snapshot = os.path.join(os.path.dirname(os.path.abspath(__file__)), 'python')
sys.path.insert(0, snapshot if os.path.isdir(snapshot) else r'%s')
print('[runner] module path added')

from asu_cpsat import build_many_asus_cpsat
from asu_checkpoint import request_rds_checkpoint
import pandas as pd

df = pd.read_csv(r'%s', dtype={'geoid': str})
with open(r'%s','r') as f:
    nb = json.load(f)
print('[runner] inputs: rows=' + str(len(df)) + ' nb_lists=' + str(len(nb)))
checkpoint_folder = %s
def save_checkpoint(phase, ids):
    request_rds_checkpoint(checkpoint_folder, phase, ids)

kwargs = dict(
    df=df, nb=nb,
    legacy_checkpoint_callback=save_checkpoint if checkpoint_folder else None,
    initial_asu_id=df['warm_start_asu'].tolist() if 'warm_start_asu' in df.columns else None,
    split_warm_start=%s,
    component_global=%s,
    tau=%s, pop_thresh=%d, max_asus=%d,
    time_limit=%d, workers=%d, parallel_asus=%d, verbose=True,
    configure_subsolvers=True,
    use_signed_flow=True,
    use_arborescence=False,
    use_root_articulation_implications=True,
    use_tract_first_search=False,
    use_flow_first_search=False,
    use_tract_capacity_search=False,
    # Conservative production profile: retain the proven, low-volume cuts and
    # leave the larger experimental flow/cardinality families opt-in.
    use_flow_capacity_hybrid_search=True,
    use_flow_count_envelope=False,
    use_small_root_separators=True,
    root_separator_max_size=5,
    root_separator_clause_limit=1000,
    root_separator_target_limit=500,
    solution_pool_size=32,
    full_graph_window=True,
    use_bridge_edge_bounds=True,
    use_articulation_edge_bounds=True,
    use_distance_flow_bounds=True,
    use_global_capacity_cardinality_bound=True,
    use_bridge_subtree_pruning=True,
    use_connectivity_free_repair=False,
    connectivity_free_time_limit=120.0,
    harvest_connectivity_free_asus=%s,
    partition_seed_strategy='%s',
    final_consolidation=%s,
    polish_consolidated_asus=%s,
    harvest_all_connectivity_free_components=%s,
    joint_partition_expansion=False,
    use_graph_cut_repair=False,
    graph_cut_repair_time_limit=120,
    standalone_expansion_time_limit=1800,
    incumbent_stall_seconds=%d,
    expansion_incumbent_stall_seconds=%d,
    stop_flag_path=r'%s',
    skip_flag_path=r'%s',
    progress_out_path=r'%s'
)
REL = %s
if REL is not None and not (isinstance(REL, float) and math.isnan(REL)):
    kwargs['rel_gap'] = REL
print('[runner] window scope: ' + ('all tracts; warm-start memberships are changeable' if kwargs['component_global'] else 'parent plus reachable unassigned tracts, excluding other-ASU contacts' if kwargs['split_warm_start'] else 'all remaining tracts'), flush=True)
print('[runner] strategy: ' + ('component_global' if kwargs['component_global'] else 'split' if kwargs['split_warm_start'] else 'partition' if kwargs['harvest_connectivity_free_asus'] else 'single'), flush=True)
print('[runner] search: ' + ('global component cuts; automatic ASU count' if kwargs['component_global'] else 'joint separated children' if kwargs['split_warm_start'] else 'flow-capacity-hybrid'), flush=True)
print('[runner] launching CP-SAT…', flush=True)
res = build_many_asus_cpsat(**kwargs)
print('[runner] solver finished; writing json', flush=True)

asu_id = res['asu_id'] if isinstance(res, dict) else getattr(res,'asu_id')
asu_id = [int(x) for x in asu_id]
if checkpoint_folder:
    save_checkpoint('LEGACY_DONE', asu_id)
output_path = r'%s'
output = {'asu_id': asu_id}
if kwargs['component_global'] and isinstance(res, dict):
    for key in ('status', 'optimal', 'total_unemp', 'upper_bound', 'absolute_gap', 'relative_gap', 'rounds'):
        if key in res:
            output[key] = res[key]
with open(output_path + '.tmp','w') as g:
    json.dump(output, g)
os.replace(output_path + '.tmp', output_path)
print('[runner] OK', flush=True)
",
      py_mod_path, df_csv, nb_json,
        if (legacy_checkpoint) as.character(jsonlite::toJSON(run_dir, auto_unbox = TRUE)) else "None",
        if (use_split) "True" else "False",
        if (use_component_global) "True" else "False",
        format(tau, scientific = FALSE), pop_thr, max_asus, tlimit, workers, parallel_asus,
  if (use_partitioning) "True" else "False",
  partition_seed_strategy,
  if (!identical(input$cpsat_final_consolidation, FALSE)) "True" else "False",
  if (isTRUE(input$cpsat_polish_consolidated_asus)) "True" else "False",
  if (use_partitioning) "True" else "False",
        incumbent_stall_seconds,
        expansion_incumbent_stall_seconds,
        stop_file, skip_file,
      progress_json,
      if (is.na(rel_gap)) "None" else as.character(rel_gap),
      out_json
    )
    writeLines(runner_code, runner)
    append_cpsat_log(paste0("[run] runner.py: ", runner))

    # ---- 5) Locate Python in the conda env ---------------------------------
    shiny::incProgress(0.05, detail = "Locating Python")
    envs <- reticulate::conda_list()
    pybin <- envs$python[envs$name == env_name]
    if (!length(pybin) || !file.exists(pybin)) {
      pybin <- Sys.which("python")
      if (pybin == "") pybin <- Sys.which("python3")
    }
    if (!nzchar(pybin) || !file.exists(pybin)) {
      shiny::showModal(shiny::modalDialog(
        title="Python not found",
        "Could not find the conda environment or python executable.",
        easyClose=TRUE
      ))
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    append_cpsat_log(paste0("[run] Using Python: ", pybin))

    # ---- 6) Launch independent job and attach this viewer ----------------
    asu_job_prepare(run_dir, initial_sf, pybin, py_mod_path, cpsat_strategy,
      settings = list(tau = tau, pop_threshold = pop_thr, workers = workers,
                      time_limit = tlimit, tract_year = sel_year))
    px <- asu_job_launch(run_dir)
    append_cpsat_log(sprintf("[run] Detached job supervisor started (pid=%s).", px$get_pid()))
    attach_cpsat_job(run_dir)

  })

  # stop the parent observer here so non-CP-SAT code below doesn’t run
  })
  }, once = TRUE)
  return(invisible(NULL))
}
# --- end CP-SAT branch --------------------------------------------------

  ## 1 · orig/mine: wrap the long job in a single withProgress ------------
  shiny::withProgress(message = "Running ASU initialisation\u2026", value = 0, {
    ## 1a · Build tracts + neighbours (can take ~10\u201360 s the first time)
    shiny::incProgress(0.10, detail = "Downloading Census shapefiles\u2026")

    tract_list <- tryCatch(asu_load_tract_geography(
      uploaded_data(), selected_year, new_england_2021), error = identity)
    if (inherits(tract_list, "error")) {
      shiny::showModal(shiny::modalDialog(title = "Tract geography mismatch",
        conditionMessage(tract_list), easyClose = TRUE))
      return(invisible(NULL))
    }
    tract_list$row_num <- seq_len(nrow(tract_list))
    tract_list$continuous <- asu_cached_neighbors(tract_list,
      sfdep::st_contiguity(sf::st_geometry(tract_list)))

    # Build and cache island-fixed neighbor graph for contiguity checks.
    nb_adjacency(list(
      geoids = tract_list$GEOID,
      nb     = asu_cached_neighbors(tract_list, fix_nb_islands(
        lapply(tract_list$continuous, function(x)
          if (length(x) == 1L && x == 0L) integer(0L) else as.integer(x)),
        as.numeric(tract_list$INTPTLAT),
        as.numeric(tract_list$INTPTLON)
      ), mode = "queen-island-v1")
    ))

    ## 1b · Merge the BLS data (kept for potential checks/debug)
    shiny::incProgress(0.3, detail = "Merging BLS unemployment data\u2026")
    data_merge <- asu_join_tract_data(tract_list, uploaded_data())

    # (kept for backward-compat, even if not used directly)
    join_flag <- isTRUE(input$join_touching)

    shiny::incProgress(0.05, detail = "Finding initial ASU seeds…")
    if (identical(algo_choice, "orig")) {
      res <- run_asu_original(tract_list, uploaded_data())
    } else {
      shiny::incProgress(0.3, detail = "Seeding & expanding\u2026")
      th_tract_list(tract_list)
      th_bls_df(uploaded_data())
      st <- tract_hunter_seed(th_tract_list(), th_bls_df(), verbose = TRUE)
      th_state(st)
      res <- tract_hunter_finalize(th_state())
    }

    ## 3 · Finish the bar
    shiny::incProgress(1, detail = "Finishing up…")
  }) # ---- withProgress ends here  -------------------------------------

  ## 4 · Store results & update UI ---------------------------------------
  if (!is.null(res)) {
    # These slots are expected from run_asu_original()/tract_hunter_finalize()
    full_data(res$full_data)
    full_data_reset(res$full_data_reset)
    asu_data(res$asu_data)
    asu_tracts(res$asu_tracts)
    asu_summary(res$asu_summary)

    output$asu <- shiny::renderTable(asu_summary(), digits = 3)
    # The global renderMaplibre for output$edit_map will react to full_data()
  }
})


shiny::observeEvent(input$th_pass, {
  shiny::req(th_state())
  shiny::withProgress(message = "Running ASU pass…", value = 0, {
    st <- tract_hunter_asu_pass(th_state(), verbose = TRUE)
  })
  th_state(st)
  res <- tract_hunter_finalize(st)
  full_data(res$full_data)
  full_data_reset(res$full_data_reset)
  asu_data(res$asu_data)
  asu_tracts(res$asu_tracts)
  asu_summary(res$asu_summary)
  output$asu <- shiny::renderTable(asu_summary(), digits = 3)
  
  # Efficient update instead of clearing layers
  update_map_efficiently("initial_map")
  update_map_efficiently("edit_map")
})

shiny::observeEvent(input$th_combine, {
  shiny::req(th_state())
  shiny::withProgress(message = "Combining ASUs…", value = 0, {
    st <- tract_hunter_combine_groups(th_state())
  })
  th_state(st)
  res <- tract_hunter_finalize(st)
  full_data(res$full_data)
  full_data_reset(res$full_data_reset)
  asu_data(res$asu_data)
  asu_tracts(res$asu_tracts)
  asu_summary(res$asu_summary)
  output$asu <- shiny::renderTable(asu_summary(), digits = 3)
  
  # Efficient update instead of clearing layers
  update_map_efficiently("initial_map")
  update_map_efficiently("edit_map")
})

#-------Total Unemployment Box --------------------------------
total_unemp <- shiny::reactive({
  shiny::req(full_data())                                  # wait until available
  full_data() |>
    sf::st_drop_geometry() |>
    dplyr::filter(asunum > 0) |>
    dplyr::summarise(unemp = sum(tract_ASU_unemp, na.rm = TRUE)) |>
    dplyr::pull(unemp)
})

output$total_unemp <- shiny::renderUI({
  shiny::HTML(
    paste0("<strong>Total Unemployment in ASUs:</strong> ",
           scales::comma(total_unemp()))
  )
})

output$total_unemp2 <- shiny::renderUI({
  shiny::HTML(
    paste0("<strong>Total Unemployment in ASUs:</strong> ",
           scales::comma(total_unemp()))
  )
})
```


```{r}
# In your setup chunk, ensure this is working correctly:
output$log_overlay <- shiny::renderUI({
  if (!isTRUE(cpsat_running())) return(NULL)
  
  shiny::absolutePanel(
    id = "cpsat_log_panel",
    top = 80, right = 20, width = 800, height = "70vh", draggable = TRUE,
    style = paste(
      "background:#fff; border:1px solid #ddd;",
      "box-shadow:0 2px 8px rgba(0,0,0,.15);",
      "overflow:auto; resize:both; min-width:320px; min-height:220px;",
      "max-width:calc(100vw - 40px); max-height:calc(100vh - 100px);",
      "padding:10px; z-index:1000;"
    ),
    shiny::div(
      style = "font-family:monospace; font-size:12px; white-space:pre-wrap;",
      shiny::strong("CP-SAT log"), shiny::br(),
      shiny::verbatimTextOutput("cpsat_log_tail", placeholder = FALSE),
      shiny::tags$div(
        style = "margin:8px 0 0 0; padding:6px 8px; background:#f4f7fb; border:1px solid #dbe3ef;",
        shiny::tags$span(style = "font-weight:700;", "Stage: "),
        shiny::textOutput("cpsat_stage", inline = TRUE)
      )
    ),
    shiny::actionButton(
      "stop_solve", if (isTRUE(cpsat_stop_requested())) "Stopping…" else "Stop Solve",
      class = "btn-sm btn-warning", style = "margin-right:4px;"
    ),
    shiny::actionButton(
      "skip_asu", if (identical(cpsat_active_strategy(), "component_global")) "Skip Cut Round" else "Skip to Next ASU",
      class = "btn-sm btn-info", style = "margin-right:4px;"
    ),
    shiny::downloadButton("save_live_log", "Save log as...", class = "btn-sm asu-save-as",
                          `data-asu-save-name` = "asu_solver.log"),
    shiny::actionButton("close_log", "Close", class = "btn-sm")
  )
})

# Signal the running Python solve to stop via its CP-SAT stop_search() callback
# and return its current incumbent, instead of killing the process outright.
shiny::observeEvent(input$stop_solve, {
  shiny::req(!isTRUE(cpsat_stop_requested()))
  sf <- cpsat_stop_file()
  shiny::req(!is.null(sf), isTRUE(cpsat_running()))
  if (!file.exists(sf)) file.create(sf)
  cpsat_stop_requested(TRUE)
  cpsat_stage_text("stop requested")
  append_cpsat_log("[run] Stop requested by user; waiting for solver to return its current incumbent…")
})

# End the current solve and advance according to the attached job's strategy.
shiny::observeEvent(input$skip_asu, {
  kf <- cpsat_skip_file()
  shiny::req(!is.null(kf), isTRUE(cpsat_running()))
  if (!file.exists(kf)) file.create(kf)
  if (identical(cpsat_active_strategy(), "component_global")) {
    cpsat_stage_text("skip cut round requested")
    append_cpsat_log("[run] Skip requested; ending this global cut round, preserving the best valid selection, and continuing the global search.")
  } else {
    cpsat_stage_text("skip current asu requested")
    append_cpsat_log("[run] Skip requested; ending the current solve and retaining eligible incumbent improvements. Bridge mode advances to the next pair; other queued work follows its stage-specific retry policy.")
  }
})

# Add this observer to handle closing the log manually
shiny::observeEvent(input$close_log, {
  px <- cpsat_solver_proc()
  if (!is.null(px) && px$is_alive()) {
    shiny::showNotification("This job is still running. Use Stop to end it; closing the browser only detaches.")
  } else {
    cpsat_running(FALSE)
    cpsat_stage_text("idle")
  }
})

output$cpsat_log_link <- shiny::renderUI({
  lf <- cpsat_log_saved()
  if (is.null(lf)) return(NULL)
  shiny::tagList(
    shiny::hr(style = "margin: 0.4rem 0;"),
    shiny::div(style = "font-size:11px; color:#555; margin-bottom:4px;", "Last log:"),
    shiny::div(style = "font-family:monospace; font-size:10px; word-break:break-all; color:#333; margin-bottom:6px;", lf),
    shiny::div(
      shiny::actionButton("open_log_folder2", "Open folder", class = "btn-xs btn-info", style = "margin-right:4px;"),
      shiny::actionButton("open_log_file2",   "Open log",    class = "btn-xs btn-default"),
      shiny::downloadButton("save_finished_log", "Save log as...", class = "btn-xs asu-save-as",
                            `data-asu-save-name` = "asu_solver.log")
    )
  )
})

shiny::observeEvent(input$open_log_folder2, {
  lf <- cpsat_log_saved()
  if (!is.null(lf)) try(shell.exec(dirname(normalizePath(lf, mustWork = FALSE))), silent = TRUE)
})
shiny::observeEvent(input$open_log_file2, {
  lf <- cpsat_log_saved()
  if (!is.null(lf) && file.exists(lf)) try(shell.exec(normalizePath(lf)), silent = TRUE)
})

# Copy the complete on-disk log, rather than the shortened live display.
cpsat_log_download <- function() shiny::downloadHandler(
  filename = function() "asu_solver.log",
  contentType = "text/plain",
  content = function(file) {
    lf <- cpsat_log_file() %||% cpsat_log_saved()
    shiny::req(!is.null(lf), file.exists(lf))
    if (!file.copy(lf, file, overwrite = TRUE)) stop("Could not prepare the solver log.")
  }
)
output$save_live_log <- cpsat_log_download()
output$save_finished_log <- cpsat_log_download()
shiny::outputOptions(output, "save_live_log", suspendWhenHidden = FALSE)
shiny::outputOptions(output, "save_finished_log", suspendWhenHidden = FALSE)

shiny::observeEvent(input$asu_save_notice, {
  notice <- input$asu_save_notice
  shiny::showNotification(notice$message,
                          type = if (identical(notice$type, "error")) "error" else "message",
                          duration = if (identical(notice$type, "error")) NULL else 8)
})

# Make sure the log text stays updated

```

```{r}
# Add debounced reactive for map updates
map_update_trigger <- shiny::debounce(shiny::reactive({
  list(full_data(), input$update)
}), 200)

# Observer for debounced updates — fires when full_data() or input$update changes.
# Assignment edits update paint only; geometry travels through the shared HTTP source.
shiny::observeEvent(map_update_trigger(), {
  if (!is.null(full_data()) && nrow(full_data()) > 0) {
    update_map_efficiently("edit_map", paint_only = TRUE)
  }
}, ignoreInit = TRUE)
```

```{r}
# Read-only overview map for Load Initial ASU tab; keyed on geoid_set() so it
# only re-renders when the tract set changes, not on every asunum edit.
output$initial_map <- mapgl::renderMaplibre({
  map_revision()
  if (!length(geoid_set())) return(mapgl::maplibre(style = mapgl::carto_style("positron")))
  map_data <- shiny::isolate(map_geom())
  if (is.null(map_data) || nrow(map_data) == 0) {
    return(
      mapgl::maplibre(style = mapgl::carto_style("positron")) |>
      mapgl::add_control("No valid polygon geometry to display.", "top-left")
    )
  }

  map_data <- map_data |> dplyr::mutate(asunum = dplyr::na_if(asunum, 0L))

  colors <- calculate_tract_colors(map_data)

  map <- mapgl::maplibre(style = mapgl::carto_style("positron")) |>
    mapgl::fit_bounds(asu_display_bounds(map_data), animate = FALSE)
  source <- shiny::isolate(map_source())
  map$x$sources <- list(source$source)
  layer_source <- "tracts-source"
  source_layer <- source$source_layer

  map |>
    mapgl::add_fill_layer(
      id                 = "basemap",
      source             = layer_source,
      source_layer       = source_layer,
      fill_color         = mapgl::match_expr(
        column = "GEOID",
        values = colors$values,
        stops = colors$stops,
        default = "#E0E0E0"
      ),
      fill_opacity       = 0.5,
      fill_outline_color = "black",
      tooltip = NULL
    ) |>
    mapgl::add_line_layer(
      id = "exploring",
      source = layer_source,
      source_layer = source_layer,
      line_color = "#000000",
      line_width = 3,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_line_layer(
      id = "exploring_candidate",
      source = layer_source,
      source_layer = source_layer,
      line_color = "#FFD54F",
      line_width = 3,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "exploring_removed",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "#D32F2F",
      fill_opacity = 0.25,
      fill_antialias = FALSE,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "exploring_added",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "#2E7D32",
      fill_opacity = 0.25,
      fill_antialias = FALSE,
      visibility = "none"
    )
}) |>
  shiny::bindCache(map_revision(), cache = "session")
```


## Map {data-width="800"}
```{r}
# Mount the lightweight log output before the map widget so that, when both
# invalidate in the same reactive flush, the browser receives the log first.
shiny::uiOutput("log_overlay")

mapgl::maplibreOutput("initial_map", height = "100vh")

```


# Modify ASU Selections

## Controls {data-width="200" style="height: 100vh; overflow-y: auto; padding-right: 10px;"}

These tools allow you to select, inspect, and change ASU designations for individual Census tracts.  *Selected* Census tracts will be outlined in red, and are the tracts updated when you change the ASU number.  You may also click on the map to select or unselect individual tracts. An ASU value of 0 means the tract is not included in any ASU.

```{r}
shiny::selectInput("display_state_edit", "Display area:", choices = character(0))
#-------Total Unemployment Box --------------------------------
shiny::htmlOutput("total_unemp2")
shiny::renderUI(shiny::HTML("<hr>"))

# Select an ASU — render once, update choices in-place to avoid full widget rebuild
shiny::selectInput("select_asunum", "Select ASU Number:",
              choices = "None", selected = "None")
shiny::renderUI(shiny::HTML("<hr>"))

# Change ASU
shiny::numericInput("new_asu", "Change ASU Value:", value = 0, min = 0)
shiny::actionButton("update", "Update Selected Tracts")
shiny::renderUI(shiny::HTML("<hr>"))

# Clear Selected Tracts
shiny::actionButton("clear", "Clear Selected Tracts")
shiny::renderUI(shiny::HTML("<hr>"))

asu_choices <- shiny::reactive({
  shiny::req(full_data())
  display_data() |> dplyr::pull(asunum) |> unique() |> sort()
}) |> shiny::debounce(400)

# Push choice updates in-place; 400 ms debounce absorbs rapid sequential edits.
shiny::observe({
  shiny::updateSelectInput(session, "select_asunum",
                           choices = c("None", asu_choices()))
}) |> shiny::bindEvent(asu_choices(), ignoreNULL = FALSE, ignoreInit = FALSE)

shiny::renderUI(shiny::HTML("<hr>"))

# Show Lowest UR Tracts in Selection
shiny::selectInput("unemployment_filter", 
             "Highlight selected tracts with:",
             choices = list("Low unemployment" = "low", 
                           "High unemployment" = "high"),
             selected = "low")
shiny::numericInput("percentile", "Percentile Threshold:", value = 5, min = 1, max = 100, step = 1)
shiny::actionButton("filter_percentile", "Highlight Tracts")
shiny::actionButton("clear_percentile", "Clear Highlights")
shiny::renderUI(shiny::HTML("<hr>"))

# Single GROID Selection
shiny::textInput("manual_geoid", "Enter a GEOID:", placeholder = "32031990000")
shiny::actionButton("select_geoid", "Select a Single Tract")
shiny::renderUI(shiny::HTML("<hr>"))

# Reset to Initial Load
shiny::actionButton("reset", "Reset to Initial Data")
```

## Interactive Map {data-width="600"}

```{r}
# The map
  mapgl::maplibreOutput("edit_map", height = "100vh")
  
output$edit_map <- mapgl::renderMaplibre({

  map_revision()
  if (!length(geoid_set())) return(mapgl::maplibre(style = mapgl::carto_style("positron")))
  # isolate() prevents reactive dependency on map_geom()/full_data() so that
  # asunum edits do not invalidate this output and cause MapLibre to reinitialise.
  data <- shiny::isolate(map_geom())
  if (is.null(data) || nrow(data) == 0) return(NULL)

  colors <- calculate_tract_colors(data)
  map <- mapgl::maplibre(style = mapgl::carto_style("positron")) |>
    mapgl::fit_bounds(asu_display_bounds(data), animate = FALSE)
  source <- shiny::isolate(map_source())
  map$x$sources <- list(source$source)
  layer_source <- "tracts-source"
  source_layer <- source$source_layer

  map |>
    mapgl::add_fill_layer(
      id = "basemap",
      source = layer_source,
      source_layer = source_layer,
      fill_color = mapgl::match_expr(
        column = "GEOID",
        values = colors$values,
        stops = colors$stops,
        default = "#E0E0E0"
      ),
      fill_opacity = 0.5,
      fill_outline_color = "black",
      tooltip = NULL
    ) |>
    mapgl::add_line_layer(
      id = "exploring",
      source = layer_source,
      source_layer = source_layer,
      line_color = "#000000",
      line_width = 3,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_line_layer(
      id = "exploring_candidate",
      source = layer_source,
      source_layer = source_layer,
      line_color = "#FFD54F",
      line_width = 3,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "exploring_removed",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "#D32F2F",
      fill_opacity = 0.25,
      fill_antialias = FALSE,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "exploring_added",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "#2E7D32",
      fill_opacity = 0.25,
      fill_antialias = FALSE,
      visibility = "none"
    ) |>
    # Pre-add hidden overlay layers; visibility="none" is the off-state so that
    # mapgl 0.5.0's filter-stack (base+user composed via ["all",...]) is not
    # poisoned by a permanent "__NONE__" base filter.
    mapgl::add_line_layer(
      id = "selected",
      source = layer_source,
      source_layer = source_layer,
      line_color = "red",
      line_width = 4,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "highlighted",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "black",
      fill_opacity = 0.8,
      visibility = "none"
    )
}) |>
  shiny::bindCache(map_revision(), cache = "session")

# OPTIMIZATION 1: Use set_filter instead of clear/add for selection highlighting
shiny::observeEvent(input$edit_map_feature_click, {
  click <- input$edit_map_feature_click
  current_selection <- selected_tracts()

  clicked_geoid <- click$properties$GEOID
  if (is.null(clicked_geoid) || length(clicked_geoid) == 0 || is.na(clicked_geoid)) return()
  if (!clicked_geoid %in% display_data()$GEOID) return()

  # toggle selection
  if (clicked_geoid %in% current_selection) {
    current_selection <- setdiff(current_selection, clicked_geoid)
  } else {
    current_selection <- c(current_selection, clicked_geoid)
  }
  selected_tracts(current_selection)

  # Build filter expression and apply (with visibility toggle)
  apply_overlay_filter(
    mapgl::maplibre_proxy("edit_map"),
    "selected",
    current_selection
  )

  output$selection_table <- shiny::renderTable({ selected_tracts() }, digits = 3)
})


# OPTIMIZATION 2: Debounce rapid interactions
clear_selection_debounced <- shiny::debounce(shiny::reactive({
  input$clear
}), 100)

shiny::observeEvent(clear_selection_debounced(), {
  selected_tracts(NULL)
  
  # Use apply_overlay_filter to hide selection layer
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
  
  # Reset ASU dropdown to "None"
    shiny::updateSelectInput(session, "select_asunum", selected = "None")
})

# OPTIMIZATION 3: Efficient ASU selection with set_filter
shiny::observeEvent(input$select_asunum, {
  shiny::req(map_geom())

  if (input$select_asunum == "None") {
    selected_tracts(NULL)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
    return()
  }

  selected_tract_ids <- map_geom() |>
    dplyr::filter(asunum == as.numeric(input$select_asunum)) |>
    dplyr::pull(GEOID)

  if (length(selected_tract_ids) == 0) {
    selected_tracts(NULL)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
    return()
  }

  selected_tracts(selected_tract_ids)

  apply_overlay_filter(
    mapgl::maplibre_proxy("edit_map"),
    "selected",
    selected_tract_ids
  ) |>
    mapgl::fit_bounds(asu_display_bounds(subset(full_data(), GEOID %in% selected_tract_ids)), animate = TRUE)
})


# Manual GEOID selection with set_filter
shiny::observeEvent(input$select_geoid, {
  shiny::req(map_geom())

  selected_tract <- map_geom() |>
    dplyr::filter(GEOID == input$manual_geoid)

  if (nrow(selected_tract) > 0) {
    shiny::updateSelectInput(session, "select_asunum", selected = "None")

    apply_overlay_filter(
      mapgl::maplibre_proxy("edit_map"),
      "selected",
      selected_tract$GEOID
    ) |>
      mapgl::fit_bounds(asu_display_bounds(selected_tract), animate = TRUE)

    selected_tracts(selected_tract$GEOID)
  }
})


# OPTIMIZATION 4: Memoize expensive calculations
# Memoized filter of selected tracts by percentile
filtered_tracts_memo <- shiny::reactive({
  shiny::req(input$percentile, input$unemployment_filter, full_data(), selected_tracts())

  potential_highlights <- full_data() |>
    dplyr::filter(GEOID %in% selected_tracts(),
                  tract_pop_cur > 0, tract_ASU_clf > 0)

  if (nrow(potential_highlights) == 0) return(character(0))

  if (input$unemployment_filter == "low") {
    threshold <- stats::quantile(potential_highlights$tract_ASU_urate, input$percentile / 100)
    potential_highlights |>
      dplyr::filter(tract_ASU_urate <= threshold) |>
      dplyr::pull(GEOID)
  } else {
    threshold <- stats::quantile(potential_highlights$tract_ASU_urate, 1 - (input$percentile / 100))
    potential_highlights |>
      dplyr::filter(tract_ASU_urate >= threshold) |>
      dplyr::pull(GEOID)
  }
}) |>
  shiny::bindCache(input$percentile, input$unemployment_filter, selected_tracts(), full_data()$asunum)

# Apply highlights only when the button is explicitly clicked
shiny::observe({
  if (is.null(selected_tracts()) || length(selected_tracts()) == 0) {
    shiny::showNotification("Select tracts on the map or via the ASU dropdown first.",
                            type = "warning", duration = 4)
    return()
  }
  filtered_selected <- filtered_tracts_memo()
  highlighted_tracts(filtered_selected)

  if (length(filtered_selected) > 0 && !any(is.na(filtered_selected))) {
    apply_overlay_filter(
      mapgl::maplibre_proxy("edit_map"),
      "highlighted",
      filtered_selected
    )
  } else {
    shiny::showNotification("No tracts matched the filter criteria.", type = "message", duration = 3)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", character(0))
  }
}) |> shiny::bindEvent(input$filter_percentile, ignoreInit = TRUE)

# Clear highlights on demand; do NOT clear on select_asunum change
shiny::observeEvent(input$clear_percentile, {
  highlighted_tracts(character(0))
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", character(0))
}, ignoreInit = TRUE)




# Update with dynamic color changes
shiny::observeEvent(input$update, {
  shiny::req(selected_tracts(), full_data())
  old_data <- full_data()
  new_asu <- as.numeric(input$new_asu)
  
  # Batch update for better performance
  updated_data <- old_data |>
    dplyr::mutate(
      asunum = dplyr::case_when(
        GEOID %in% selected_tracts() ~ new_asu,
        TRUE ~ asunum
      )
    )
  full_data(updated_data)
  selected_tracts(NULL)

  # Paint-only update fires immediately so colors change without waiting for the
  # 200 ms debounce. The shared geometry source remains unchanged.
  cm <- calculate_tract_colors(map_geom())
  proxy <- mapgl::maplibre_proxy("edit_map") |>
    mapgl::set_paint_property(
      "basemap", "fill-color",
      mapgl::match_expr(
        column  = "GEOID",
        values  = cm$values,
        stops   = cm$stops,
        default = "#E0E0E0"
      )
    )
  apply_overlay_filter(proxy, "selected", character(0))
})


refresh_edit_map <- function() {
  shiny::req(map_geom())
  data <- map_geom()
  
  # Use the efficient update function that handles data changes
  update_map_with_data("edit_map", data) |>
    mapgl::fit_bounds(asu_display_bounds(data), animate = TRUE)
}


shiny::observeEvent(input$reset, {
  shiny::req(full_data_reset())
  full_data(full_data_reset())
  selected_tracts(NULL)
  highlighted_tracts(character(0))
  shiny::updateSelectInput(session, "select_asunum", selected = "None")
  
  # Use the efficient refresh function
  refresh_edit_map()
})

# OPTIMIZATION 8: Optimized data loading observer
# Observer for loading the data
shiny::observeEvent(input$load_data, {
  shiny::req(input$load_data)
  
  # Read the RDS file into a data frame
  loaded_data <- tryCatch({
    readRDS(input$load_data$datapath)
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Loading Data",
      "Failed to load the data. Please check the file format.",
      easyClose = TRUE
    ))
    return(NULL)
  })
  
  # If loaded_data is NULL (failed to load), exit the observer
  shiny::req(!is.null(loaded_data))
  
  # Ensure the loaded data is an sf object
  if (!inherits(loaded_data, "sf")) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Loading Data",
      "The loaded data is not a valid sf object.",
      easyClose = TRUE
    ))
    return(NULL)
  }
  
  # Update full_data with the loaded data
  full_data(loaded_data)
  full_data_reset(loaded_data)
  nb_adjacency(NULL)
  display_state(sort(unique(substr(loaded_data$GEOID, 1L, 2L)))[1L])
  
  # Provide feedback to the user
  shiny::showModal(shiny::modalDialog(
    title = "Data Loaded",
    "Your data has been successfully loaded.",
    easyClose = TRUE
  ))
  
  # Use the efficient update function for new data
  # map_revision() creates the selected display source after this observer finishes.
  
  # Clear any existing selections/highlights
  selected_tracts(NULL)
  highlighted_tracts(NULL)
})

# Trigger map resize when panels are resized
shiny::observeEvent(input$map_resize, {
  # Force map to resize/redraw
  mapgl::maplibre_proxy("edit_map") |>
    mapgl::maplibre_resize()
}, ignoreInit = TRUE)
```

## Selection Summary {data-width="200"}

```{r}
preview_data <- shiny::reactive({
  shiny::req(full_data(), length(selected_tracts()) > 0)
  selected_geoids <- as.character(selected_tracts())
  replacement_data <- full_data() |>
    sf::st_drop_geometry() |>
    dplyr::filter(as.character(GEOID) %in% selected_geoids)
  summary_data <- asu_summarize_tracts(replacement_data)
  

  list(data = replacement_data, summary = summary_data)
})

# Conditional display based on selection
shiny::conditionalPanel(
  condition = "output.has_selection == true",
  style = "height: calc(100vh - 20px); display: flex; flex-direction: column;",
  
  shiny::h3(" Summary of Selected Tracts"),
  shiny::hr(),

  # Summary statistics
  shiny::div(
    style = "background-color: #f8f9fa; padding: 15px; border-radius: 5px; margin-bottom: 15px; flex-shrink: 0;",
    shiny::textOutput("summary_count"),
    shiny::br(),
    shiny::textOutput("summary_total_pop"),
    shiny::textOutput("summary_total_labor_force"),
    shiny::textOutput("summary_avg_unemployment_rate"),
    shiny::textOutput("summary_total_unemployment")
  ),
  
  shiny::hr(style = "flex-shrink: 0;"),
  
  # Detailed table
  shiny::h4("Tract Details", style = "flex-shrink: 0;"),
  shiny::div(
    style = "flex: 1; overflow-y: auto; border: 1px solid #dee2e6; border-radius: 5px; min-height: 400px;",
    DT::dataTableOutput("summary_table")
  )
)

# No selection message
shiny::conditionalPanel(
  condition = "output.has_selection == false",
  shiny::div(
    style = "text-align: center; color: #6c757d; font-style: italic; padding: 50px 20px;",
    shiny::h4("No tracts currently selected"),
    shiny::p("Select tracts on the map or use the controls to view summary data here.")
  )
)

# Server logic for conditional display
output$has_selection <- shiny::reactive({
  length(selected_tracts()) > 0
})
shiny::outputOptions(output, "has_selection", suspendWhenHidden = FALSE)

# Summary statistics outputs using your existing preview_data reactive
output$summary_count <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Number of selected tracts:", preview_result$summary$Tracts)
})

output$summary_total_pop <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total population:", format(preview_result$summary$Population, big.mark = ","))
})

output$summary_total_labor_force <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total labor force:", format(preview_result$summary$`Labor Force`, big.mark = ","))
})

output$summary_avg_unemployment_rate <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Aggregate unemployment rate:", preview_result$summary$`Unemployment Rate`, "%")
})

output$summary_total_unemployment <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total unemployment:", format(preview_result$summary$Unemployment, big.mark = ","))
})

# Detailed table — interactive DT with bidirectional map-highlight sync
output$summary_table <- DT::renderDataTable({
  shiny::req(selected_tracts())

  preview_data()$data |>
    dplyr::select(GEOID, asunum, tract_pop_cur, tract_ASU_clf,
                  tract_ASU_urate, tract_ASU_unemp) |>
    dplyr::arrange(GEOID) |>
    dplyr::rename(
      ASU             = asunum,
      Population      = tract_pop_cur,
      `Labor Force`   = tract_ASU_clf,
      `Unemp. Rate %` = tract_ASU_urate,
      `Unemp.`        = tract_ASU_unemp
    ) |>
    DT::datatable(
      selection  = "multiple",
      rownames   = FALSE,
      options    = list(
        pageLength = 25,
        scrollX    = TRUE,
        dom        = "ftp",
        columnDefs = list(list(className = "dt-right", targets = c(1, 2, 3, 4, 5)))
      )
    ) |>
    DT::formatRound(columns = "Unemp. Rate %", digits = 2) |>
    DT::formatCurrency(columns = c("Population", "Labor Force", "Unemp."),
                       currency = "", digits = 0)
}, server = TRUE)

# DT row selected → highlight those tracts on the map
shiny::observeEvent(input$summary_table_rows_selected, ignoreNULL = FALSE, {
  shiny::req(selected_tracts(), preview_data())
  sel_rows  <- input$summary_table_rows_selected
  tbl_geoids <- preview_data()$data |> dplyr::arrange(GEOID) |> dplyr::pull(GEOID)
  geoids <- if (length(sel_rows) > 0) tbl_geoids[sel_rows] else character(0)
  highlighted_tracts(geoids)
  filter_expr <- if (length(geoids) == 0) {
    NULL
  } else if (length(geoids) == 1) {
    list("==", list("get", "GEOID"), as.character(geoids))
  } else {
    list("in", list("get", "GEOID"), list("literal", as.character(geoids)))
  }
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", geoids)
})

# highlighted_tracts() changed externally (percentile filter / map) → reflect in DT
shiny::observe({
  hl <- highlighted_tracts()
  shiny::req(selected_tracts(), preview_data())
  tbl_geoids  <- preview_data()$data |> dplyr::arrange(GEOID) |> dplyr::pull(GEOID)
  row_indices <- which(tbl_geoids %in% hl)
  if (!identical(as.integer(shiny::isolate(input$summary_table_rows_selected)),
                 as.integer(row_indices))) {
    DT::dataTableProxy("summary_table") |> DT::selectRows(row_indices)
  }
}) |> shiny::bindEvent(highlighted_tracts(), ignoreInit = TRUE)
```

# Save and Load Data

## Controls {data-width="150"}

These buttons allow you to save the map in progress so that you can load
it again in the future to continue your work.

Use **Save Data As...** to choose a filename and folder on your computer.
After loading data, go directly to the ASU Modification tab to continue
making edits. The export directory below is used only for LSS and summary CSV files.

```{r}
shiny::includeScript(file.path(dirname(knitr::current_input(dir = TRUE)), "save-as.js"))

# Server-side directory for LSS and CSV exports
shiny::textInput("save_dir_path", "LSS / CSV Export Directory (on the R server):", value = getwd(), width = "100%")
shiny::textOutput("current_save_dir")
shiny::textOutput("dir_status")
shiny::br()

shiny::downloadButton("save_data", "Save Data As...", class = "asu-save-as",
                      `data-asu-save-name` = "saved_data.rds")
shiny::helpText("Choose a filename and folder in the save dialog. If your browser downloads immediately, enable Ask where to save each file in its download settings.")
shiny::fileInput("load_data", "Load Data", accept = c(".rds"))

# Reactive value to store the selected directory
save_directory <- shiny::reactiveVal(getwd())  # Default to current working directory

# Observer to handle directory path changes
shiny::observeEvent(input$save_dir_path, {
  if (!is.null(input$save_dir_path) && input$save_dir_path != "") {
    if (dir.exists(input$save_dir_path)) {
      save_directory(input$save_dir_path)
    }
  }
})

# Display current save directory
output$current_save_dir <- shiny::renderText({
  paste("Current save directory: ", save_directory())
})

output$review_save_dir <- shiny::renderText({
  paste("Current save directory: ", save_directory())
})

# Display directory status
output$dir_status <- shiny::renderText({
  if (dir.exists(save_directory())) {
    "✓ Directory exists and is accessible"
  } else {
    "⚠ Directory does not exist - it will be created when saving"
  }
})

# Prepare an RDS download; the browser chooses the destination.
output$save_data <- shiny::downloadHandler(
  filename = function() "saved_data.rds",
  contentType = "application/octet-stream",
  content = function(file) {
    shiny::req(full_data())
    saveRDS(full_data(), file)
  }
)
shiny::outputOptions(output, "save_data", suspendWhenHidden = FALSE)


```

## Output {data-width="850"}

```{r}
# Empty output section
```

# ASU Review and Finalization

## Controls {data-width="150"}

This screen will allow you to review your selected ASUs. It uses a
default unemployment-rate threshold of 6.45% and a population threshold
of 10,000. Update either value here when the applicable program guidance
changes.

*Generate ASU Summary* will show a table of all ASUs, and whether they
pass or fail the unemployment, population, and contiguity tests. The
qualification calculation uses the unrounded aggregate rate, even though
the displayed rate is rounded. Once generated, the table stays synchronized
with subsequent assignment edits and threshold changes.

Note, you *may* go back and forth between the selection and review
screens.

When ready to create an output file, the *Generate LSS .txt File* button
will create a formatted batch file for upload to LSS.

```{r}
shiny::numericInput("asu_ur", "ASU Unemployment Rate Threshold:", value = 6.45, min = 0)

shiny::numericInput("asu_pop", "ASU Population Threshold:", value = 10000, min = 0)

shiny::actionButton("run_asu_summary", "Generate ASU Summary")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::textOutput("review_save_dir")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::actionButton("lss_txt", "Generate LSS .txt File")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::actionButton("review_txt", "Create Summary CSV")
```

## Output {data-width="850"}

```{r}
DT::dataTableOutput("asu_review")

review_summary_requested <- shiny::reactiveVal(FALSE)

shiny::observeEvent(input$run_asu_summary, {
  review_summary_requested(TRUE)
})

summarized_asu <- shiny::reactive({
  shiny::req(review_summary_requested(), full_data(), input$asu_ur, input$asu_pop)

  asu_sf <- full_data() |>
    dplyr::mutate(.review_asunum = suppressWarnings(as.integer(as.character(asunum)))) |>
    dplyr::filter(!is.na(.review_asunum), .review_asunum > 0L)
  summary_table <- asu_summarize_tracts(full_data(), by_asu = TRUE)

  # Contiguity check: use the island-fixed full-state neighbor graph built at
  # load time so that island tracts connected via centroid distance are handled
  # consistently with how the solver built the ASUs.
  asu_ids <- summary_table$asunum

  nb_check <- nb_adjacency()
  if (is.null(nb_check)) {
    # Fallback for sessions loaded from RDS: rebuild from current geometry.
    all_sf  <- full_data()
    raw_nb  <- asu_cached_neighbors(all_sf, lapply(
      suppressWarnings(sfdep::st_contiguity(all_sf$geometry)),
      function(x) if (length(x) == 1L && x == 0L) integer(0L) else as.integer(x)
    ))
    fixed_nb <- if (all(c("INTPTLAT", "INTPTLON") %in% names(all_sf)))
      asu_cached_neighbors(all_sf,
        fix_nb_islands(raw_nb, as.numeric(all_sf$INTPTLAT), as.numeric(all_sf$INTPTLON)),
        mode = "queen-island-v1")
    else raw_nb
    nb_check <- list(geoids = all_sf$GEOID, nb = fixed_nb)
    nb_adjacency(nb_check)
  }

  contiguous_flag <- vapply(asu_ids, function(id) {
    grp_geoids <- asu_sf |>
      dplyr::filter(.review_asunum == id) |>
      dplyr::pull(GEOID)
    asu_is_contiguous(grp_geoids, nb_check)
  }, logical(1L))

  contiguity_df <- data.frame(asunum = asu_ids, contiguous = contiguous_flag)

  summary_table |>
    dplyr::left_join(contiguity_df, by = "asunum") |>
    dplyr::mutate(
      ur_qualified   = dplyr::if_else(
        !is.na(`Unemployment Rate Exact`) &
          `Unemployment Rate Exact` >= input$asu_ur,
        TRUE, FALSE
      ),
      pop_qualified  = dplyr::if_else(Population >= input$asu_pop, TRUE, FALSE),
      contiguous     = contiguous,
      asu_qualified  = ur_qualified & pop_qualified & contiguous
    ) |>
    dplyr::select(-`Unemployment Rate Exact`)
})

output$asu_review <- DT::renderDataTable({
  tbl <- summarized_asu()
  shiny::req(tbl)

  tbl |>
    dplyr::mutate(
      ur_qualified  = ifelse(ur_qualified,  "TRUE", "FALSE"),
      pop_qualified = ifelse(pop_qualified, "TRUE", "FALSE"),
      contiguous    = ifelse(contiguous,    "TRUE", "FALSE"),
      asu_qualified = ifelse(asu_qualified, "TRUE", "FALSE")
    ) |>
    DT::datatable(
      rownames  = FALSE,
      selection = "none",
      
      options   = list(
        scrollX   = TRUE,
        pageLength = 25,
        dom        = "ftp",
        order      = list(list(0, "asc"))
      )
    ) |>
    DT::formatRound(columns = "Unemployment Rate", digits = 3) |>
    DT::formatCurrency(
      columns  = c("Population", "Labor Force", "Unemployment"),
      currency = "", digits = 0
    ) |>
    DT::formatStyle(
      columns     = c("ur_qualified", "pop_qualified", "contiguous", "asu_qualified"),
      color       = DT::styleEqual(c("TRUE", "FALSE"), c("#2e7d32", "#c62828")),
      fontWeight  = "bold"
    )
}, server = TRUE)

shiny::observeEvent(input$lss_txt, {
  
  shiny::req(full_data())

  # Function to generate text content
  generate_txt <- function() {
    df <- full_data() |>
      sf::st_drop_geometry() |>
      dplyr::select(GEOID, asunum) |>
      dplyr::filter(asunum > 0) |>
      dplyr::mutate(asunum = as.integer(asunum),
             GEOID = paste0("14000US", GEOID),
             asunum = paste0("SU", state(), sprintf("%04d", asunum)),
             GEOID = paste(asunum, GEOID, sep = " + ")) |> 
      dplyr::arrange(asunum, GEOID) |>
      dplyr::select(-asunum)

    txt_content <- utils::capture.output(utils::write.table(df, quote=FALSE, row.names=FALSE, col.names = FALSE))
    txt_content <- paste(txt_content, collapse="\n")
    return(txt_content)
  }

  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    shiny::showModal(shiny::modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  txt_content <- generate_txt()

  # Use the selected save directory
  filePath <- file.path(current_dir, "lss_batch_file.txt")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      shiny::showModal(shiny::modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the file
    writeLines(txt_content, filePath)
    
    # Show success message with file location
    shiny::showModal(shiny::modalDialog(
      title = "File Created Successfully",
      shiny::HTML(paste0(
        "<strong>LSS batch file has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Creating File",
      paste("Failed to create LSS file:", e$message),
      easyClose = TRUE
    ))
  })
})

shiny::observeEvent(input$review_txt, {

  shiny::req(full_data())

  # Function to generate CSV content
  generate_csv <- function() {
    df <- full_data() |>
      sf::st_drop_geometry() |>
      dplyr::arrange(asunum, GEOID)
    
    return(df)
  }
  
  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    shiny::showModal(shiny::modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  df <- generate_csv()

  # Use the selected save directory
  filePath <- file.path(current_dir, "ASU_Review_File.csv")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      shiny::showModal(shiny::modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the CSV file
    utils::write.csv(df, filePath, row.names = FALSE)
    
    # Show success message with file location
    shiny::showModal(shiny::modalDialog(
      title = "File Created Successfully",
      shiny::HTML(paste0(
        "<strong>ASU Review CSV has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Creating File",
      paste("Failed to create CSV file:", e$message),
      easyClose = TRUE
    ))
  })
})

####################
## Future Development

# Export Excel Worksheet with tract details and selections.
# Allow "Save" of data with txt/Excel files so single session is not required.
# Better table formatting.
# highlight high unemployment tracts in non-included areas.
# Check for contiguity of selected ASUs after manual edits.
# Export high-resolution images of ASUs for review.
# Use reactlog package to monitor reactive interactions in the code for cleanup. https://mastering-shiny.org/reactive-graph.html
```

# How to Use ASUbuildR

## User guide

```{r dashboard-user-guide, echo=FALSE}
dashboard_guide_path <- file.path(
  dirname(knitr::current_input(dir = TRUE)), "dashboard-guide.md"
)
shiny::includeMarkdown(dashboard_guide_path)
```
