---
title: "ASUbuildR"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
runtime: shiny
editor_options: 
  markdown: 
    wrap: 72
---


<style>
.dataTables_wrapper {
  width: fit-content !important;
  max-width: 100%;
}

.dataTables_scroll {
  width: fit-content !important;
  max-width: 100%;
}

.dataTables_scrollBody {
  height: auto !important;
  max-width: 100%;
}

table.dataTable {
  width: auto !important;
}

table.dataTable th,
table.dataTable td {
  white-space: nowrap;
}
</style>

```{r python_preamble, include=FALSE}
library(reticulate)
library(processx)

# Always define this first so options() won't error
python_ready <- FALSE
env_name <- "asu-cpsat"

# Try to activate the conda env and check for OR-Tools, but never abort knitting
try({
  envs <- reticulate::conda_list()$name
  if (env_name %in% envs) {
    reticulate::use_condaenv(env_name, required = FALSE)

    if (reticulate::py_module_available("ortools")) {
      python_ready <- TRUE
      message("[env] Python environment ready")
    } else {
      message("[env] Conda env found, but 'ortools' is missing")
    }
  } else {
    message("[env] Python environment not set up; run ASUbuildR::setup_asu_python()")
  }
}, silent = TRUE)

# Store the flag for later chunks
options(asu_python_ready = python_ready)
```

```{r setup, include=FALSE}
# Figure out full paths to src/ C++ files, going up two levels
src_cpp_files <- c("choose_best_neighbor.cpp",
                   "build_edges.cpp",
                   "choose_best_drop.cpp")

src_paths <- normalizePath(file.path("..", "..", "src", src_cpp_files),
                           winslash = "/", mustWork = FALSE)

# Check that all source files are there
in_repo <- all(file.exists(src_paths))

maybe_sourceCpp <- function(fun, path) {
  if (!exists(fun, mode = "function") && file.exists(path)) {
    message("Compiling ", basename(path))
    Rcpp::sourceCpp(path)
  }
}

if (in_repo) {
  message("→ Dev mode: compiling C++ helpers from src/")
  maybe_sourceCpp("choose_best_neighbor", src_paths[1])
  maybe_sourceCpp("build_edges",          src_paths[2])
  maybe_sourceCpp("choose_best_drop",     src_paths[3])
} else {
  message("→ Installed package mode (shared library already loaded)")

}

# ---- dev mode: source helper functions relative to this file ----------
is_dev_mode <- file.exists("../../DESCRIPTION") && dir.exists("../../R")

if (is_dev_mode) {
  message("→ Dev mode: sourcing from ../../R/")
  r_dir <- normalizePath("../../R", winslash = "/", mustWork = TRUE)

  source(file.path(r_dir, "run_tract_hunter.R"))
  source(file.path(r_dir, "run_asu_original.r"))
  source(file.path(r_dir,"setup_asu_python.R"))
} else {
  message("→ Installed package mode: functions should already be available")
}

`%||%` <- function(a, b) if (!is.null(a) && length(a)) a else b


# ── Use conda environment & import Python module ───────────────────
# This is best-effort: if Python/Conda or the 'asu-cpsat' env is missing,
# the dashboard must still start and load correctly (only the CP-SAT
# algorithm option becomes unavailable). Never let this stop the document.
asu_py <- NULL
if (isTRUE(python_ready)) {
  tryCatch({
    reticulate::use_condaenv(env_name, required = TRUE)
    py_config()  # should now show the env python

    # Import the bundled solver module
    asu_py <- import_from_path(
      module  = "asu_cpsat",
      path    = system.file("python", package = "ASUbuildR"),
      convert = TRUE
    )
    stopifnot(py_has_attr(asu_py, "build_many_asus_cpsat"))
  }, error = function(e) {
    message("[env] Could not load the CP-SAT Python module: ", conditionMessage(e))
    python_ready <<- FALSE
    asu_py <<- NULL
  })
}
options(asu_python_ready = python_ready)


## ---- tigris cache ----------------------------------------------------
## put cached TIGER/Line ZIPs in a per-user cache folder that is always
## writable, even after the package is installed read-only.

# 1. choose sensible user-cache paths (rappdirs works everywhere)
cache_dir <- rappdirs::user_cache_dir("ASUbuildR", "tigris")
pmtiles_cache_dir <- rappdirs::user_cache_dir("ASUbuildR", "pmtiles")

# 2. create it if it doesn't exist
if (!dir.exists(cache_dir))
  dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)
if (!dir.exists(pmtiles_cache_dir))
  dir.create(pmtiles_cache_dir, recursive = TRUE, showWarnings = FALSE)

options(
  tigris_cache_dir = cache_dir,
  tigris_use_cache = TRUE,
  tigris_class     = "sf",
  download.file.method = "libcurl"

)

options(shiny.maxRequestSize = 50 * 1024^2)  # Set to 50MB or your desired limit

# Initialize reactive values
uploaded_data <- shiny::reactiveVal(NULL)
state <- shiny::reactiveVal(NULL)
tract_year <- shiny::reactiveVal(NULL)
tigris_year <- shiny::reactiveVal(NULL)
tract_data <- shiny::reactiveVal(NULL)
asu_data <- shiny::reactiveVal(NULL)
asu_tracts <- shiny::reactiveVal(NULL)
asu_summary <- shiny::reactiveVal(NULL)
full_data <- shiny::reactiveVal(NULL)
full_data_reset <- shiny::reactiveVal()
selected_tracts <- shiny::reactiveVal(NULL)
nb_adjacency    <- shiny::reactiveVal(NULL) # island-fixed full-state neighbor graph
pmtiles_server  <- shiny::reactiveVal(NULL)
pmtiles_disabled <- shiny::reactiveVal(
  !isTRUE(getOption("ASUbuildR.use_pmtiles", TRUE)) ||
    !requireNamespace("freestiler", quietly = TRUE)
)


# New England Optional year selection logic
output$is_new_england <- shiny::reactive({
  if (is.null(state())) {
    return(FALSE)
  }

  # New England state FIPS codes
  new_england_fips <- c("09", "23", "25", "33", "44", "50") # CT, ME, MA, NH, RI, VT

  state() %in% new_england_fips
})

# ---- CP-SAT logging + paths ----
cpsat_log_buf    <- shiny::reactiveVal("")
cpsat_solver_proc<- shiny::reactiveVal(NULL)
cpsat_run_dir    <- shiny::reactiveVal(NULL)
cpsat_df_csv     <- shiny::reactiveVal(NULL)
cpsat_nb_json    <- shiny::reactiveVal(NULL)
cpsat_out_json   <- shiny::reactiveVal(NULL)
cpsat_log_file   <- shiny::reactiveVal(NULL)
cpsat_log_saved  <- shiny::reactiveVal(NULL)
cpsat_stop_file  <- shiny::reactiveVal(NULL)
cpsat_stop_requested <- shiny::reactiveVal(FALSE)
cpsat_skip_file  <- shiny::reactiveVal(NULL)
cpsat_running <- shiny::reactiveVal(FALSE)

append_cpsat_log <- function(...) {
  line <- paste0(format(Sys.time(), "%H:%M:%S "), paste0(..., collapse = ""))
  old  <- cpsat_log_buf()
  new  <- if (nzchar(old)) paste(old, line, sep = "\n") else line
  if (nchar(new) > 12000) new <- substr(new, nchar(new) - 11999, nchar(new))
  cpsat_log_buf(new)
  lf <- cpsat_log_file()
  if (!is.null(lf)) try(cat(line, "\n", file = lf, append = TRUE), silent = TRUE)
}

# keep log output live even if hidden
output$cpsat_log_tail <- shiny::renderText(cpsat_log_buf())
shiny::outputOptions(output, "cpsat_log_tail", suspendWhenHidden = FALSE)
output$show_cpsat_log_flag <- shiny::renderText({
  if (isTRUE(cpsat_running())) "true" else "false"
})
shiny::outputOptions(output, "show_cpsat_log_flag", suspendWhenHidden = FALSE)

# Kill the solver process if the user closes the browser before it finishes.
session$onSessionEnded(function() {
  px <- shiny::isolate(cpsat_solver_proc())
  if (!is.null(px) && px$is_alive()) px$kill()

  tile_server <- shiny::isolate(pmtiles_server())
  if (!is.null(tile_server) && requireNamespace("freestiler", quietly = TRUE)) {
    try(freestiler::stop_server(tile_server$port), silent = TRUE)
  }
})

# ---- New England year override logic ---------------------------------
# Make it available to the UI
shiny::outputOptions(output, "is_new_england", suspendWhenHidden = FALSE)

# Tract Hunter staged state -----------------------------------------
th_state <- shiny::reactiveVal(NULL)
th_tract_list <- shiny::reactiveVal(NULL)
th_bls_df <- shiny::reactiveVal(NULL)

prep_map_data <- function(map_data) {
  # Cache the result to avoid repeated processing
  if (nrow(map_data) == 0) return(map_data)
  
  map_data |>
    sf::st_make_valid() |>
    sf::st_simplify(dTolerance = .005, preserveTopology = TRUE) |>  # Simplify geometry
    sf::st_zm(what = "ZM", drop = TRUE) |>
    dplyr::filter(!sf::st_is_empty(geometry)) |>
    dplyr::mutate(geom_type = sf::st_geometry_type(geometry)) |>
    dplyr::filter(geom_type %in% c("POLYGON", "MULTIPOLYGON")) |>
    dplyr::select(-geom_type)
}

# PERFORMANCE: st_make_valid()/st_simplify() over every tract is expensive,
# but the geometry only changes when a new tract set is loaded (new state or
# year). Editing ASU assignments only changes the `asunum` attribute, so
# cache the cleaned geometry keyed on the GEOID set rather than on
# full_data() itself (which changes on every edit and would defeat a
# full_data()-keyed cache, forcing re-simplification on every click).
base_geom <- shiny::reactive({
  shiny::req(full_data())
  full_data() |>
    dplyr::select(GEOID) |>
    prep_map_data()
}) |>
  shiny::bindCache(sort(full_data()$GEOID))

# Cheap non-spatial join of current attributes onto the cached geometry.
# Only include columns needed for rendering and tooltips to minimise GeoJSON size.
map_geom <- shiny::reactive({
  shiny::req(full_data())
  attrs <- full_data() |>
    sf::st_drop_geometry() |>
    dplyr::select(GEOID, asunum, tract_pop_cur, tract_ASU_clf, tract_ASU_urate, tract_ASU_unemp)
  dplyr::inner_join(base_geom(), attrs, by = "GEOID")
})

map_tile_fingerprint <- function(data) {
  attrs <- data |>
    sf::st_drop_geometry() |>
    dplyr::arrange(GEOID)
  fingerprint_file <- tempfile(fileext = ".rds")
  on.exit(unlink(fingerprint_file), add = TRUE)
  saveRDS(attrs, fingerprint_file, version = 2)
  unname(tools::md5sum(fingerprint_file))
}

map_cache_component <- function(value, fallback = "unknown") {
  values <- as.character(value %||% fallback)
  values <- values[!is.na(values) & nzchar(values)]
  if (length(values) == 0L) values <- fallback
  values <- gsub("[^A-Za-z0-9_-]", "_", values)
  paste(sort(unique(values)), collapse = "-")
}

ensure_map_tiles <- function(data) {
  if (isTRUE(pmtiles_disabled()) || is.null(data) || nrow(data) == 0) return(NULL)

  tryCatch({
    state_id <- map_cache_component(state())
    year_id <- map_cache_component(tigris_year())
    fingerprint <- map_tile_fingerprint(data)
    tile_name <- sprintf("tracts_v2_%s_%s_%s.pmtiles", state_id, year_id, fingerprint)
    tile_path <- file.path(pmtiles_cache_dir, tile_name)

    if (!file.exists(tile_path)) {
      temp_path <- tempfile(pattern = "asu_tracts_", tmpdir = pmtiles_cache_dir,
                            fileext = ".pmtiles")
      on.exit(unlink(temp_path), add = TRUE)
      message("[map] Creating PMTiles cache: ", tile_name)
      freestiler::freestile(
        data,
        temp_path,
        layer_name = "tracts",
        tile_format = "mvt",
        min_zoom = 0L,
        max_zoom = 14L,
        simplification = TRUE,
        overwrite = TRUE,
        quiet = TRUE
      )
      if (!file.rename(temp_path, tile_path)) {
        if (!file.copy(temp_path, tile_path, overwrite = TRUE)) {
          stop("Could not move the completed PMTiles archive into the cache.")
        }
      }
      message("[map] PMTiles cache created: ",
              format(file.info(tile_path)$size, big.mark = ","), " bytes")
    } else {
      message("[map] Reusing PMTiles cache: ", tile_name)
    }

    server_info <- pmtiles_server()
    if (is.null(server_info)) {
      server_info <- freestiler::serve_tiles(
        pmtiles_cache_dir,
        port = httpuv::randomPort()
      )
      pmtiles_server(server_info)
    }

    list(
      source_id = "tracts-source",
      source_layer = "tracts",
      url = paste0(server_info$url, "/", tile_name),
      path = tile_path
    )
  }, error = function(e) {
    warning("PMTiles rendering disabled for this session: ", conditionMessage(e),
            call. = FALSE)
    pmtiles_disabled(TRUE)
    NULL
  })
}

map_uses_pmtiles <- function() {
  !isTRUE(pmtiles_disabled()) && !is.null(pmtiles_server())
}

# Reactively tracks the GEOID set but only fires when GEOIDs actually change.
# Because full_data() changes on every ASU edit (asunum values), map_geom()
# is invalidated on every edit too. If output$edit_map
# depend directly on map_geom(), they are invalidated and Shiny re-sends the
# cached HTML to the browser, which re-initialises MapLibre and wipes any
# proxy color updates. Keying on geoid_set() instead means the map outputs
# are only invalidated (and re-sent) when the set of tracts genuinely changes.
geoid_set <- shiny::reactiveVal(character(0))
shiny::observe({
  shiny::req(full_data())
  new_geoids <- sort(full_data()$GEOID)
  if (!identical(geoid_set(), new_geoids)) {
    geoid_set(new_geoids)
  }
})

highlighted_tracts <- shiny::reactiveVal(character(0))

# Load static data
excel_names <- c("record",
                 "geoid",
                 "st_fips",
                 "cnty_fips",
                 "tract_fips",
                 "name",
                 "tract_pop_dec",
                 "tract_pop_cur",
                 "tract_emp",
                 "tract_unemp",
                 "tract_urate",
                 "tract_urate_error",
                 "cnty_pop_dec",
                 "cnty_pop_cur",
                 "cnty_emp",
                 "cnty_unemp",
                 "cnty_urate",
                 "cnty_urate_error",
                 "pop_shr",
                 "emp_shr",
                 "unemp_shr",
                 "laus_primary",
                 "cnty_emp_ave",
                 "cnty_unemp_ave",
                 "tract_ASU_clf",
                 "tract_ASU_emp",
                 "tract_ASU_unemp",
                 "tract_ASU_urate")

#-------- Map Rendering Helpers --------------------------------
# Helper function to calculate color scheme (used for full layer rebuilds)
calculate_color_scheme <- function(data) {
  number_asus <- max(data$asunum, na.rm = TRUE)
  number_categories <- min(c(number_asus, 8))
  
  if (number_categories == 1) {
    asu_breaks <- 1
    actual_categories <- 1
    brewer_pal <- "#7FC97F"
  } else {
    asu_breaks <- seq(from = 1, to = number_asus, length.out = number_categories)
    actual_categories <- max(3, number_categories)
    full_brewer_pal <- RColorBrewer::brewer.pal(actual_categories, "Accent")
    brewer_pal <- full_brewer_pal[1:number_categories]
  }
  
  return(list(breaks = asu_breaks, colors = brewer_pal))
}

# Build a per-GEOID -> color lookup from current asunum assignments.
# Used by update_map_efficiently() so the paint expression is keyed on GEOID
# (a stable feature property that never changes after the initial source load)
# rather than on asunum (which IS stale in the cached MapLibre source).
calculate_tract_colors <- function(data) {
  n_asu <- suppressWarnings(max(data$asunum, na.rm = TRUE))
  n_asu <- if (is.finite(n_asu) && n_asu > 0L) as.integer(n_asu) else 1L

  if (n_asu == 1L) {
    pal <- "#7FC97F"
  } else {
    n_cat <- min(max(n_asu, 3L), 8L)
    pal <- RColorBrewer::brewer.pal(n_cat, "Accent")
  }

  # pal_ext[1] = grey (asunum 0), pal_ext[2..] = palette colors for asunum 1..n
  pal_ext <- c("#E0E0E0", rep_len(pal, n_asu))
  idx <- as.integer(data$asunum) + 1L   # asunum 0 -> index 1, asunum 1 -> index 2, …
  idx[is.na(idx) | idx < 1L | idx > length(pal_ext)] <- 1L

  list(
    values = as.character(data$GEOID),
    stops  = pal_ext[idx]
  )
}

tract_tooltip <- function() {
  mapgl::concat(
    "<strong>ASU Number: </strong>",
    mapgl::get_column("asunum"), "<br>",
    "<strong>Tract GEOID: </strong>", mapgl::get_column("GEOID"), "<br>",
    "<strong>Tract Population: </strong>", mapgl::number_format(mapgl::get_column("tract_pop_cur"), style = "decimal", maximum_fraction_digits = 0), "<br>",
    "<strong>Tract Labor Force: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_clf"), style = "decimal", maximum_fraction_digits = 0), "<br>",
    "<strong>Tract Unemployment Rate: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_urate"), style = "decimal", maximum_fraction_digits = 2), "%<br>",
    "<strong>Tract Unemployment: </strong>", mapgl::number_format(mapgl::get_column("tract_ASU_unemp"), style = "decimal", maximum_fraction_digits = 0), "<br>"
  )
}

# Helper: set filter + visibility on the 'selected' or 'highlighted' overlay layer.
# In mapgl 0.5.0, set_filter() composes user+base via ["all", ...], so we MUST
# NOT set an initial filter on these layers. Use visibility="none" as the off-state
# and pair every set_filter call with a matching set_layout_property visibility toggle.
apply_overlay_filter <- function(proxy, layer_id, ids) {
  ids <- ids[!is.na(ids)]
  if (length(ids) == 0) {
    proxy |>
      mapgl::set_filter(layer_id, NULL) |>
      mapgl::set_layout_property(layer_id, "visibility", "none")
  } else {
    filter_expr <- if (length(ids) == 1) {
      list("==", list("get", "GEOID"), as.character(ids))
    } else {
      list("in", list("get", "GEOID"), list("literal", as.character(ids)))
    }
    proxy |>
      mapgl::set_filter(layer_id, filter_expr) |>
      mapgl::set_layout_property(layer_id, "visibility", "visible")
  }
}

update_map_efficiently <- function(map_id = "edit_map", paint_only = FALSE) {
  data <- map_geom()
  if (is.null(data) || nrow(data) == 0) return()

  cm <- calculate_tract_colors(data)

  sel <- selected_tracts()   %||% character(0)
  hl  <- highlighted_tracts() %||% character(0)

  proxy <- mapgl::maplibre_proxy(map_id)

  if (!paint_only && !map_uses_pmtiles()) {
    # set_source refreshes tooltip feature properties; skip for paint-only edits.
    proxy <- proxy |> mapgl::set_source(layer_id = "basemap", source = data)
  }

  proxy <- proxy |>
    mapgl::set_paint_property(
      "basemap", "fill-color",
      mapgl::match_expr(
        column  = "GEOID",
        values  = cm$values,
        stops   = cm$stops,
        default = "#E0E0E0"
      )
    )

  if (map_id == "edit_map") {
    proxy <- apply_overlay_filter(proxy, "selected",    sel)
    proxy <- apply_overlay_filter(proxy, "highlighted", hl)
  }
  proxy
}

# Enhanced version that can also update the data source if needed
update_map_with_data <- function(map_id = "edit_map", new_data = NULL) {
  data <- if (is.null(new_data)) map_geom() else new_data
  if (is.null(data) || nrow(data) == 0) return()

  if (!is.null(new_data) && map_uses_pmtiles()) {
    proxy <- update_map_efficiently(map_id, paint_only = TRUE)
    return(proxy)
  }
  
  color_scheme <- calculate_color_scheme(data)
  
  # If new data is provided, we need to update the source
  if (!is.null(new_data)) {
    mapgl::maplibre_proxy(map_id) |>
      mapgl::clear_layer("basemap") |>
      mapgl::clear_layer("selected") |>
      mapgl::clear_layer("highlighted") |>
      mapgl::add_fill_layer(
        id = "basemap",
        source = data,
        fill_color = mapgl::step_expr(
          column = "asunum",
          base = "#E0E0E0",
          stops = color_scheme$colors,
          values = color_scheme$breaks,
          na_color = "#E0E0E0"
        ),
        fill_opacity = 0.5,
        fill_outline_color = "black",
        tooltip = tract_tooltip()
      ) |>
      mapgl::add_line_layer(
        id = "selected",
        source = data,
        line_color = "red",
        line_width = 4,
        line_opacity = 1,
        visibility = "none"
      ) |>
      mapgl::add_fill_layer(
        id = "highlighted",
        source = data,
        fill_color = "black",
        fill_opacity = 0.8,
        visibility = "none"
      )
  } else {
    # Just update the paint property
    mapgl::maplibre_proxy(map_id) |>
      mapgl::set_paint_property("basemap", "fill-color",
        mapgl::step_expr(
          column = "asunum",
          base = "#E0E0E0",
          stops = color_scheme$colors,
          values = color_scheme$breaks,
          na_color = "#E0E0E0"
        )
      )
  }
}
# Connect every sub-graph to the largest component via nearest centroid.
fix_nb_islands <- function(nb, intpt_lat, intpt_lon) {
  g     <- igraph::graph_from_adj_list(nb)
  comps <- igraph::components(g)
  if (comps$no <= 1L) return(nb)
  main_id  <- which.max(comps$csize)
  main_idx <- which(comps$membership == main_id)
  for (comp_id in setdiff(unique(comps$membership), main_id)) {
    island_idx <- which(comps$membership == comp_id)
    combis     <- expand.grid(island = island_idx, main = main_idx)
    combis$dist <- sqrt(
      (intpt_lat[combis$island] - intpt_lat[combis$main])^2 +
      (intpt_lon[combis$island] - intpt_lon[combis$main])^2
    )
    closest <- combis[which.min(combis$dist), ]
    nb[[closest$island]] <- c(nb[[closest$island]], closest$main)
    nb[[closest$main]]   <- c(nb[[closest$main]],   closest$island)
  }
  nb
}

# Check whether asu_geoids form a single connected component in stored_nb.
asu_is_contiguous <- function(asu_geoids, stored_nb) {
  idx <- which(stored_nb$geoids %in% asu_geoids)
  if (length(idx) <= 1L) return(TRUE)
  local_nb <- lapply(seq_along(idx), function(i) {
    nbrs  <- stored_nb$nb[[idx[i]]]
    local <- which(idx %in% nbrs)
    if (length(local) == 0L) integer(0L) else as.integer(local)
  })
  g <- igraph::graph_from_adj_list(local_nb, mode = "all")
  igraph::components(g)$no == 1L
}

```

# Data Initialization
## Controls {data-width="150"}
This screen allows you to import the BLS-provided ASU file, generally named like _NV_asu23.xlsx_ which will be the only external data import step necessary. The script will read the file, extract the state, and determine the population year from the Excel file.

You can specify which year of Census tract boundaries to use with the selector below. This is important because Census tract boundaries may change between years, and you'll want to use the appropriate year that matches the GEOIDs in your BLS file.

If the necessary data is not in columns **A through AB** within the Excel file, this process will not work correctly.
Once the file is loaded, a preview of the data will appear on the right, the state and year information will display below, and the analyst may proceed to the next tab (Load Initial ASU).
Note: at this time, ASUbuildR assumes a state is using county-based census tracts. If a state uses another geography such as NECTAs as the basis for its ASU tract file, this process will require some modification.
```{r}
shiny::fileInput("file", "Choose Excel File", accept = c(".xls", ".xlsx"))
# Add numeric input for tract year
shiny::numericInput("tract_year_input",
                     "Census Tract Year:",
                     value = 2024,
                     min = 2010,
                     max = 2025,
                     step = 1)
shiny::helpText("Select the year of Census tract boundaries to use.
                 For New England states, 2021 is the last year with NECTA-based tracts.")
shiny::textOutput("selected_state")
shiny::textOutput("tract_pop_year")
shiny::textOutput("selected_tract_year")
```

## Output {data-width="850"}

```{r}
shiny::tableOutput("data_preview")
shiny::observeEvent(input$file, {
  shiny::req(input$file)
  df <- readxl::read_excel(input$file$datapath, range = "A2:AB25000", col_names = excel_names) |>
    dplyr::filter(!is.na(geoid)) |>
    dplyr::mutate(GEOID = stringr::str_remove(geoid, "14000US")) |>
    dplyr::select(GEOID,
           st_fips,
           cnty_fips,
           tract_fips,
           name,
           tract_pop_cur,
           tract_ASU_clf,
           tract_ASU_emp,
           tract_ASU_unemp,
           tract_ASU_urate) |>
    dplyr::mutate(
      tract_pop_cur = as.integer(round(tract_pop_cur)),
      tract_ASU_clf = as.integer(round(tract_ASU_clf)),
      tract_ASU_emp = as.integer(round(tract_ASU_emp)),
      tract_ASU_unemp = as.integer(round(tract_ASU_unemp))
    )
  uploaded_data(df)
  uploaded_data() |>
    dplyr::pull(st_fips) |>
    unique() |>
    state()
  
  # Get population year from the Excel file header
  readxl::read_excel(input$file$datapath, range = "H1:H1", col_names = "year") |>
    dplyr::pull(year) |>
    stringr::str_remove("tract_pop") |>
    tract_year()
  
  output$data_preview <- shiny::renderTable({
    head(uploaded_data(), 50)
  })
  
  output$selected_state <- shiny::renderText({
    paste0("Selected FIPS is: ", state())
  })
  
  output$tract_pop_year <- shiny::renderText({
    paste0("Population year (detected from Excel): ", tract_year())
  })
})

# Update tigris_year reactiveVal when input changes
shiny::observeEvent(input$tract_year_input, {
  yr <- suppressWarnings(as.integer(input$tract_year_input))
  if (is.na(yr) || yr < 2010 || yr > 2025) {
    output$selected_tract_year <- shiny::renderText({
      paste0("Invalid year entered (", input$tract_year_input, "); must be 2010-2025. Using previous value: ", tigris_year() %||% "not set")
    })
    return()
  }
  tigris_year(as.character(yr))
  output$selected_tract_year <- shiny::renderText({
    paste0("Census tract boundary year (user selected): ", tigris_year())
  })
}, ignoreInit = FALSE)
```

# Load Initial ASU

## Controls {data-width="200"}

```{r}
# ---- New England year override controls ------------------------------
shiny::conditionalPanel(
  condition = "output.is_new_england === true",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::h5("New England Year Override"),
    shiny::radioButtons(
      inputId = "ne_year_choice",
      label   = "Override tract year for New England:",
      choices = list(
        "Use selected year from Data Initialization" = "detected",
        "Force 2021 (last NECTA year)" = "2021"
      ),
      selected = "detected",
      inline = FALSE
    ),
    shiny::p(
      style = "font-size: 0.9em; color: #666;",
      "Note: 2021 is the last year NECTA-based tracts were available for New England states. 
       You may have already selected 2021 in the Data Initialization tab."
    )
  )
)

# ---- algorithm picker -------------------------------------------------
shiny::radioButtons(
  inputId = "asu_algo",
  label   = "Choose algorithm",
  choices = c(
    "Tract Hunter"      = "mine",
    "Simple Snake"      = "orig",
    "CP-SAT (OR-Tools)" = "cpsat"
  ),
  inline  = TRUE
)


# -----------------------------------------------------------------------
# main "Run" button (always visible)
shiny::div(style = "margin-top: 0.5rem;",
    shiny::actionButton(
      inputId = "process",
      label   = "Load tracts & initialise",
      class   = "btn-primary"
    )
)

# -----------------------------------------------------------------------
# extra Tract-Hunter controls (only when algo == 'mine')
shiny::conditionalPanel(
  condition = "input.asu_algo === 'mine'",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::div(
      class = "btn-group",           # keep them on one line
      style = "width: 100%;",
      shiny::actionButton("th_pass",    "Run Hunter",     class = "btn-secondary"),
      shiny::actionButton("th_combine", "Combine Groups",   class = "btn-secondary")
    )
  )
)

# Conditional panel for CP-SAT algorithm options
shiny::conditionalPanel(
  condition = "input.asu_algo === 'cpsat'",
  shiny::tagList(
    shiny::hr(style = "margin: 0.8rem 0;"),
    shiny::numericInput("cpsat_tau",       "UR threshold (τ)",          value = 0.0645, min = 0, max = 1, step = 0.0001),
    shiny::numericInput("cpsat_pop",       "Population threshold",      value = 10000,  min = 0, step = 100),
    shiny::numericInput("cpsat_max",       "Max ASUs",                  value = 30,     min = 1, step = 1),
    shiny::numericInput("cpsat_timelimit", "Time limit (sec) per window", value = 18000, min = 0, step = 60),
    shiny::numericInput("cpsat_workers",   "Total CP-SAT workers (detected cores - 2)", value = max(1, parallel::detectCores(logical = FALSE)-2), min = 1, step = 1),
    shiny::numericInput("cpsat_parallel_asus", "Concurrent ASU solves", value = 1, min = 1, step = 1),
    shiny::numericInput("cpsat_relgap",        "Relative gap (optional)",    value = NULL, min = 0, step = 0.0001),
    shiny::checkboxInput("cpsat_enable_max_nodes", "Limit tracts per ASU (cap + combine)", value = FALSE),
    shiny::conditionalPanel(
      condition = "input.cpsat_enable_max_nodes === true",
      shiny::numericInput("cpsat_max_nodes",           "Max tracts per ASU",                     value = 500, min = 1, step = 1),
      shiny::numericInput("cpsat_combine_timelimit",   "Combine/re-solve time limit (sec, optional)", value = NULL, min = 0, step = 60)
    ),
    shiny::uiOutput("cpsat_log_link")
  )
)



# ---- dynamic algorithm description -----------------------------------
output$algo_blurb <- shiny::renderUI({
  choice <- input$asu_algo %||% "orig"

  if (identical(choice, "cpsat")) {
    ## ───────── CP-SAT (OR-Tools) METHOD ─────────
    python_status <- if (isTRUE(getOption("asu_python_ready", FALSE))) {
      "<p style=\"color:#2e7d32;\"><strong>Python/OR-Tools environment detected.</strong></p>"
    } else {
      "<p style=\"color:#b71c1c;\"><strong>Python/OR-Tools environment not detected.</strong> Run <code>ASUbuildR::setup_asu_python()</code> in R, then restart the dashboard. Until then, use Tract Hunter or Simple Snake.</p>"
    }
    shiny::HTML(
      paste0(python_status, "<p><strong>CP-SAT (OR-Tools) workflow</strong></p>
       <p>
         This option builds ASUs by solving a mixed-integer model over tract
         adjacency. The app will:
       </p>
       <ul>
         <li>Download Census tract shapes for the selected year and build
             neighbour lists (contiguity).</li>
         <li>Merge the BLS upload you provided (population, labor force,
             employment, unemployment, rate).</li>
         <li>Run the CP-SAT solver with your parameters:
             <code>UR threshold (τ)</code>, <code>Population threshold</code>,
             <code>Max ASUs</code>, <code>Time limit</code>, optional
             <code>Relative gap</code>, <code>Total CP-SAT workers</code>, and
             <code>Concurrent ASU solves</code>.</li>
       </ul>
       <p>
         <strong>Limit tracts per ASU (cap + combine):</strong> leave this
         toggle off for unrestricted ASU sizes (default). Enabling it reveals
         two extra controls: <code>Max tracts per ASU</code>, which caps each
         ASU built during the main pass at that many tracts (so the solver
         builds as many separate, non-overlapping capped ASUs as it can), and
         <code>Combine/re-solve time limit</code>, an optional separate CP-SAT
         time budget for the final pass. Once no more capped ASUs can be
         built, touching capped ASUs are automatically combined and re-solved
         (uncapped) by CP-SAT -- repeating until no more touching ASUs remain
         -- to improve the merged result. Leaving the combine time limit blank
         reuses the main <code>Time limit</code> for that pass.
       </p>
       <p>
         While solving, a live <em>CP-SAT log</em> panel appears (top-right)
         showing progress. When the solver finishes, the map and summary table
         update automatically; tracts with <code>asunum = 0</code> are not in
         any ASU.
       </p>
       <p>
         <strong>Requirements:</strong> a working Python environment with
         OR-Tools. If it isn't available, you'll see a notice and can use
         Tract Hunter or Simple Snake instead. For New England states, you may
         need to switch the tract year to 2021 using the option shown above.
       </p>
       <p>
         Click <em>Load tracts &amp; initialise</em> to start. You can refine
         results on the <em>Modify ASU Selections</em> tab once the map renders.
       </p>")
    )

  } else if (identical(choice, "mine")) {
    ## ───────── TRACT-HUNTER METHOD ─────────
    shiny::HTML(
      "<p><strong>Tract Hunter workflow</strong></p>
       <p>Because Tract Hunter works in stages, the interface exposes those
          stages as separate buttons:</p>
       <ul>
         <li><em>Load tracts &amp; initialise</em><br>
             ▸ downloads shapefiles, merges the BLS file, then runs the
             <strong>Seed + Expand</strong> step automatically. This should only be run once.</li>
         <li><em>Run Hunter</em><br>
             ▸ performs the optimisation/trade pass that adds
             high-unemployment neighbours and drops low-unemployment ones until
             every ASU is as close as possible to the
             <code>6.45%</code> threshold.</li>
         <li><em>Combine Groups</em><br>
             ▸ merges adjacent ASUs that share a boundary. Connected ASUs allow for more tracts to be added, so whenever ASUs are merged it is advisable to run another pass with the Tract Hunter.</li>
       </ul>
       <p>You can stop after any step to inspect the map, or run the buttons
          in sequence for the full, automatic workflow. We suggest iteratively running the Run Hunter and Combine Groups steps until Total Unemployment in ASUs no longer increases. Each step updates the
          colour map and the unemployment totals shown on the right.</p>"
    )

  } else {
    ## ───────── SIMPLE SNAKE (ORIGINAL) METHOD ─────────
    shiny::HTML(
      "<p>
        This button will launch two processes, which will take up to 5&nbsp;minutes
        to complete. For smaller states, it may take under 15&nbsp;seconds.
      </p>
      <p>
        First, the application will download census shapefiles by tract for the
        appropriate state and year. Next, after combining the BLS file with these
        shapefiles, it will iteratively search for contiguous geographic regions
        until the combined unemployment rate in each region is as close to
        6.45% as possible.
      </p>
      <p>
        When this process is complete, a map will appear to the right, and a
        summary of the created ASU regions will appear below. It is important to
        note these will still include areas that do not qualify as an independent
        ASU (there is no check for population), but it gives the analyst a
        starting place for analysis.
      </p>
      <p>
        Once this step is done, you may proceed to the next step.
      </p>"
    )
  }
})


shiny::uiOutput("algo_blurb")

shiny::htmlOutput("total_unemp")

shiny::tableOutput("asu")
#shiny::verbatimTextOutput("cpsat_log_tail")

palette_logic <- function(n_asu) {
  if (n_asu == 0) {
    list(
      fill = "#E0E0E0",
      legend_vals = "(none)",
      legend_cols = "#E0E0E0"
    )
  } else if (n_asu == 1) {
    list(
      fill = mapgl::step_expr(
        column = "asunum",
        base = "#E0E0E0",
        stops = "#7FC97F",
        values = 1,
        na_color = "#E0E0E0"
      ),
      legend_vals = 1,
      legend_cols = "#7FC97F"
    )
  } else {
    pal <- RColorBrewer::brewer.pal(min(max(n_asu, 3L), 8L), "Accent")
    legend_vals <- seq_len(n_asu)
    legend_cols <- rep_len(pal, n_asu)

    list(
      fill = mapgl::match_expr(
        column = "asunum",
        values = legend_vals,
        stops = legend_cols
      ),
      legend_vals = legend_vals,
      legend_cols = legend_cols
    )
  }
}



shiny::observeEvent(input$process, {
  ## 0) Basic checks ------------------------------------------------------
  shiny::req(uploaded_data(), state(), tigris_year())

  # Determine year (NE override)
  selected_year <- tigris_year()
  state_val <- state()
  if (!is.null(state_val) && length(state_val) > 0) {
    new_england_fips <- c("09","23","25","33","44","50")
    if (any(state_val %in% new_england_fips) &&
        !is.null(input$ne_year_choice) &&
        input$ne_year_choice == "2021") {
      selected_year <- "2021"
    }
  }

  # ensure 'res' exists even if something returns early
  res <- NULL

  # Detect algorithm before withProgress to avoid overlapping progress bars
  algo_choice <- shiny::isolate(input$asu_algo)
  if (is.null(algo_choice) || identical(algo_choice, "")) algo_choice <- "orig"

  # --- CP-SAT branch (has its own withProgress) --------------------------
if (identical(algo_choice, "cpsat")) {
  if (!isTRUE(getOption("asu_python_ready", FALSE))) {
    shiny::showModal(shiny::modalDialog(
      title = "Python Environment Not Ready",
      "The CP-SAT (OR-Tools) algorithm requires a configured Python environment. Run ASUbuildR::setup_asu_python() in R, then restart the dashboard. In the meantime, choose Tract Hunter or Simple Snake.",
      easyClose = TRUE
    ))
    return()
  }
  session <- shiny::getDefaultReactiveDomain()
  shiny::withProgress(message = "Running OR-Tools CP-SAT…", value = 0, {
    cpsat_running(TRUE)
    append_cpsat_log("[run] Button clicked (CP-SAT).")
    # ---- 1) Preconditions & parameters --------------------------------
    shiny::req(uploaded_data(), state(), tract_year())
    sel_year <- selected_year              # from your earlier NE override logic
    st_fips  <- state()
    tau      <- if (is.null(input$cpsat_tau))       0.0645 else as.numeric(input$cpsat_tau)
    pop_thr  <- if (is.null(input$cpsat_pop))       10000  else as.integer(input$cpsat_pop)
    max_asus <- if (is.null(input$cpsat_max))       30     else as.integer(input$cpsat_max)
    tlimit   <- if (is.null(input$cpsat_timelimit)) 1200   else as.integer(input$cpsat_timelimit)
    workers  <- if (is.null(input$cpsat_workers))   max(1L, parallel::detectCores()-5L) else as.integer(input$cpsat_workers)
    parallel_asus <- if (is.null(input$cpsat_parallel_asus)) 1L else max(1L, as.integer(input$cpsat_parallel_asus))
    rel_gap  <- if (!is.null(input$cpsat_relgap) && !is.na(input$cpsat_relgap)) as.numeric(input$cpsat_relgap) else NA_real_
    max_nodes_enabled <- isTRUE(input$cpsat_enable_max_nodes)
    max_nodes_per_asu <- if (max_nodes_enabled && !is.null(input$cpsat_max_nodes) && !is.na(input$cpsat_max_nodes)) as.integer(input$cpsat_max_nodes) else NA_integer_
    combine_timelimit <- if (max_nodes_enabled && !is.null(input$cpsat_combine_timelimit) && !is.na(input$cpsat_combine_timelimit)) as.integer(input$cpsat_combine_timelimit) else NA_integer_
    append_cpsat_log(sprintf("[run] params: tau=%.6f pop=%d max=%d limit=%ds workers=%d concurrent_asus=%d rel_gap=%s max_nodes_per_asu=%s combine_time_limit=%s",
                 tau, pop_thr, max_asus, tlimit, workers, parallel_asus,
                             if (is.na(rel_gap)) "None" else as.character(rel_gap),
                             if (is.na(max_nodes_per_asu)) "None" else as.character(max_nodes_per_asu),
                             if (is.na(combine_timelimit)) "None" else as.character(combine_timelimit)))
    # ---- 2) Build tracts + neighbors ----------------------------------
    shiny::incProgress(0.15, detail = "Downloading tracts & building neighbors")
    
    # 1. Collapse the vector of states into a single string for the log (e.g., "32, 06")
    st_fips_str <- paste(st_fips, collapse = ", ")
    append_cpsat_log(sprintf("[NB] Downloading TIGER tracts for state(s) %s (year=%s)…", st_fips_str, sel_year))
    
    # 2. Map over the vector of states, combine them, then calculate contiguity
    tr_sf <- purrr::map(st_fips, \(st) {
      tigris::tracts(state = st, year = sel_year, progress_bar = FALSE)
    }) |>
      dplyr::bind_rows() |>
      dplyr::mutate(continuous = sfdep::st_contiguity(geometry))
    
    # Join with uploaded BLS
    cols_needed <- c("GEOID","tract_ASU_unemp","tract_ASU_emp","tract_pop_cur","tract_ASU_clf","tract_ASU_urate")
    miss <- setdiff(cols_needed, names(uploaded_data()))
    if (length(miss)) {
      shiny::showModal(shiny::modalDialog(
        title="Input problem", paste("Uploaded data missing:", paste(miss, collapse=", ")), easyClose=TRUE
      ))
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    
    merged_sf <- tr_sf |>
      dplyr::left_join(uploaded_data(), by="GEOID")
    
    # Apply island fix
    append_cpsat_log("[NB] Applying island fix to connect disconnected components")
    
    # Convert list column to regular lists
    merged_sf$continuous <- lapply(merged_sf$continuous, function(x) {
      # If it's scalar 0, replace with integer(0)
      if (length(x) == 1 && x == 0) return(integer(0))
      as.integer(x)
    })
    
    # Use igraph to identify components
    nb <- merged_sf$continuous
    g <- igraph::graph_from_adj_list(nb)
    comps <- igraph::components(g)
    main_component_id <- which.max(comps$csize)
    main_indices <- which(comps$membership == main_component_id)
    island_indices <- which(comps$membership != main_component_id)
    
    # If there are islands, connect them to the mainland
    if (length(island_indices) > 0) {
      append_cpsat_log(sprintf("[NB] Found %d tracts in islands, connecting to mainland", length(island_indices)))
      
      coords <- data.frame(
        INTPTLAT = as.numeric(merged_sf$INTPTLAT),
        INTPTLON = as.numeric(merged_sf$INTPTLON)
      )
      
      island_component_ids <- setdiff(unique(comps$membership), main_component_id)
      for (island_id in island_component_ids) {
        island_members <- which(comps$membership == island_id)
        # For all members of this island component, get distances to all in main
        combis <- expand.grid(island = island_members, main = main_indices)
        combis$dist <- sqrt(
          (coords$INTPTLAT[combis$island] - coords$INTPTLAT[combis$main])^2 +
          (coords$INTPTLON[combis$island] - coords$INTPTLON[combis$main])^2
        )
        closest <- combis[which.min(combis$dist), ]
        g <- igraph::add_edges(g, c(closest$island, closest$main))
        
        # Update the neighborhood lists with the new connection
        nb[[closest$island]] <- c(nb[[closest$island]], closest$main)
        nb[[closest$main]] <- c(nb[[closest$main]], closest$island)
      }
      
      append_cpsat_log("[NB] Island fix complete - all tracts now connected")
    } else {
      append_cpsat_log("[NB] No islands detected - all tracts already connected")
    }
    
    # Update merged_sf with the fixed neighborhood structure
    merged_sf$continuous <- nb
    nb_adjacency(list(geoids = merged_sf$GEOID, nb = nb))

    # Pre-simplify geometry inside withProgress so map renders immediately after
    # the solver finishes (visible to the user as part of the setup phase).
    append_cpsat_log("[NB] Simplifying tract geometry...")
    sf::st_geometry(merged_sf) <- sf::st_geometry(merged_sf) |>
      sf::st_make_valid() |>
      sf::st_simplify(dTolerance = .005, preserveTopology = TRUE) |>
      sf::st_zm(what = "ZM", drop = TRUE)
    
    merged_df <- merged_sf |> sf::st_drop_geometry()
    append_cpsat_log(sprintf("[NB] Tracts kept: %d (after join)", nrow(merged_df)))
    
    # Continue with the existing code...
    # Dataframe for Python
    df_py <- merged_df |>
      dplyr::transmute(
        geoid           = GEOID,
        tract_ASU_unemp = as.integer(round(tract_ASU_unemp)),
        tract_ASU_emp   = as.integer(round(tract_ASU_emp)),
        tract_pop2024   = as.integer(round(tract_pop_cur))
      )
    
    # Use the island-fixed neighborhood lists for the Python solver
    nb_lists <- lapply(merged_sf$continuous, function(iv) {
      if (length(iv) == 0) list() else as.list(as.integer(iv) - 1L)
    })

    # ---- 3) Temp run folder + files ------------------------------------
    shiny::incProgress(0.10, detail = "Writing inputs")
    run_dir <- file.path(tempdir(), paste0("cpsat_", format(Sys.time(), "%Y%m%d_%H%M%S")))
    dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

    df_csv   <- file.path(run_dir, "df.csv")
    nb_json  <- file.path(run_dir, "nb.json")
    out_json <- file.path(run_dir, "out.json")
    runner   <- file.path(run_dir, "runner.py")
    stop_file <- file.path(run_dir, "stop.flag")
    skip_file <- file.path(run_dir, "skip.flag")

    utils::write.csv(df_py, df_csv, row.names = FALSE)
    jsonlite::write_json(nb_lists, nb_json, auto_unbox = TRUE)

    log_file <- file.path(run_dir, "solver.log")
    cpsat_run_dir(run_dir)
    cpsat_df_csv(df_csv)
    cpsat_nb_json(nb_json)
    cpsat_out_json(out_json)
    cpsat_log_file(log_file)
    cpsat_log_saved(NULL)
    unlink(stop_file)
    unlink(skip_file)
    cpsat_stop_file(stop_file)
    cpsat_stop_requested(FALSE)
    cpsat_skip_file(skip_file)

    append_cpsat_log(paste0("[run] Folder: ", run_dir))
    append_cpsat_log(paste0("[run] df.csv: ", df_csv))
    append_cpsat_log(paste0("[run] nb.json: ", nb_json))

    # ---- 4) Write runner.py -------------------------------------------
    shiny::incProgress(0.05, detail = "Preparing Python runner")
    py_mod_path <- normalizePath(system.file("python", package = "ASUbuildR"),
                                 winslash = "/", mustWork = TRUE)
    runner_code <- sprintf("
import sys, os, json, math
try:
    sys.stdout.reconfigure(line_buffering=True)
except Exception:
    pass
os.environ.setdefault('PYTHONUNBUFFERED','1')
os.environ.setdefault('OMP_NUM_THREADS','1')
os.environ.setdefault('MKL_NUM_THREADS','1')
os.environ.setdefault('OPENBLAS_NUM_THREADS','1')
print('[runner] starting'); print('Python:', sys.version.replace('\\n',' '))

sys.path.insert(0, r'%s')
print('[runner] module path added')

from asu_cpsat import build_many_asus_cpsat
import pandas as pd

df = pd.read_csv(r'%s')
with open(r'%s','r') as f:
    nb = json.load(f)
print('[runner] inputs: rows=' + str(len(df)) + ' nb_lists=' + str(len(nb)))

kwargs = dict(
    df=df, nb=nb,
    tau=%s, pop_thresh=%d, max_asus=%d,
    time_limit=%d, workers=%d, parallel_asus=%d, verbose=True,
    configure_subsolvers=True,
    use_signed_flow=True,
    use_arborescence=False,
    use_root_articulation_implications=True,
    use_tract_first_search=False,
    use_flow_count_envelope=False,
    use_small_root_separators=True,
    root_separator_max_size=6,

    root_separator_clause_limit=5000,
    root_separator_target_limit=1000,
    solution_pool_size=32,
    full_graph_window=True,
    use_bridge_edge_bounds=True,
    use_articulation_edge_bounds=True,
    use_distance_flow_bounds=True,
    use_global_capacity_cardinality_bound=True,
    use_bridge_subtree_pruning=True,
    use_connectivity_free_repair=True,
    connectivity_free_time_limit=30.0,
    harvest_connectivity_free_asus=True,
    standalone_expansion_time_limit=600.0,
    stop_flag_path=r'%s',
    skip_flag_path=r'%s'
)
REL = %s
if REL is not None and not (isinstance(REL, float) and math.isnan(REL)):
    kwargs['rel_gap'] = REL
MAXN = %s
if MAXN is not None and not (isinstance(MAXN, float) and math.isnan(MAXN)):
    kwargs['max_nodes_per_asu'] = int(MAXN)
    kwargs['combine_capped_asus'] = True
COMBINE_TL = %s
if COMBINE_TL is not None and not (isinstance(COMBINE_TL, float) and math.isnan(COMBINE_TL)):
    kwargs['combine_time_limit'] = int(COMBINE_TL)
print('[runner] window scope: all remaining tracts', flush=True)
print('[runner] launching CP-SAT…', flush=True)
res = build_many_asus_cpsat(**kwargs)
print('[runner] solver finished; writing json', flush=True)

asu_id = res['asu_id'] if isinstance(res, dict) else getattr(res,'asu_id')
asu_id = [int(x) for x in asu_id]
with open(r'%s','w') as g:
    json.dump({'asu_id': asu_id}, g)
print('[runner] OK', flush=True)
",
      py_mod_path, df_csv, nb_json,
      format(tau, scientific = FALSE), pop_thr, max_asus, tlimit, workers, parallel_asus,
      stop_file, skip_file,
      if (is.na(rel_gap)) "None" else as.character(rel_gap),
      if (is.na(max_nodes_per_asu)) "None" else as.character(max_nodes_per_asu),
      if (is.na(combine_timelimit)) "None" else as.character(combine_timelimit),
      out_json
    )
    writeLines(runner_code, runner)
    append_cpsat_log(paste0("[run] runner.py: ", runner))

    # ---- 5) Locate Python in the conda env ---------------------------------
    shiny::incProgress(0.05, detail = "Locating Python")
    envs <- reticulate::conda_list()
    pybin <- envs$python[envs$name == env_name]
    if (!length(pybin) || !file.exists(pybin)) {
      pybin <- Sys.which("python")
      if (pybin == "") pybin <- Sys.which("python3")
    }
    if (!nzchar(pybin) || !file.exists(pybin)) {
      shiny::showModal(shiny::modalDialog(
        title="Python not found",
        "Could not find the conda environment or python executable.",
        easyClose=TRUE
      ))
      cpsat_running(FALSE)
      return(invisible(NULL))
    }
    append_cpsat_log(paste0("[run] Using Python: ", pybin))

    # ---- 6) Launch and STREAM output ----------------------------------
    shiny::incProgress(0.35, detail = "Solving (streaming logs)…")
    env_map <- c(
      PYTHONUNBUFFERED = "1",
      OMP_NUM_THREADS  = "1",
      MKL_NUM_THREADS  = "1",
      OPENBLAS_NUM_THREADS = "1"
    )
    px <- processx::process$new(
      command = pybin,
      args    = c(runner),
      stdout  = "|",
      stderr  = "|",
      windows_hide_window = TRUE,
      env     = env_map
    )
    cpsat_solver_proc(px)
    append_cpsat_log(sprintf("[run] Python started (pid=%s). Streaming…", px$get_pid()))

    # stream logs and finalize when done
    if (!is.null(session$userData$cpsatObs)) {
      try(session$userData$cpsatObs$destroy(), silent = TRUE)
    }
    tmr <- shiny::reactiveTimer(300)
    session$userData$cpsatObs <- shiny::observe({
      tmr()
      px <- cpsat_solver_proc()
      shiny::req(!is.null(px))

      out <- try(px$read_output_lines(), silent = TRUE)
      err <- try(px$read_error_lines(),  silent = TRUE)
      if (!inherits(out,"try-error") && length(out)) append_cpsat_log(paste(out, collapse="\n"))
      if (!inherits(err,"try-error") && length(err)) append_cpsat_log(paste(err, collapse="\n"))

      if (!px$is_alive()) {
        # drain remaining
        out_all <- try(px$read_all_output_lines(), silent = TRUE)
        err_all <- try(px$read_all_error_lines(),  silent = TRUE)
        if (!inherits(out_all,"try-error") && length(out_all)) append_cpsat_log(paste(out_all, collapse="\n"))
        if (!inherits(err_all,"try-error") && length(err_all)) append_cpsat_log(paste(err_all, collapse="\n"))

        status <- px$get_exit_status()
        append_cpsat_log(paste0("[run] Exit status: ", status))
        session$userData$cpsatObs$destroy()

        # ---- 7) Load results -> update map/reactives -------------------
        clean_sf_columns <- function(x) {
          g <- sf::st_geometry(x)
          df <- x |>
            sf::st_drop_geometry() |>
            dplyr::select(where(~ !is.list(.)))
          sf::st_as_sf(df, geometry = g, crs = sf::st_crs(x))
        }

        if (identical(status, 0L) && file.exists(cpsat_out_json())) {
          resj <- jsonlite::fromJSON(cpsat_out_json())
          if (!is.null(resj$asu_id) && length(resj$asu_id) == nrow(merged_df)) {
            asu_id <- as.integer(resj$asu_id)

            final_sf <- merged_sf |>
              dplyr::mutate(asunum = ifelse(asu_id < 0L, 0L, as.integer(asu_id))) |>
              dplyr::select(-dplyr::any_of("continuous")) |>
              clean_sf_columns()
            # geometry already simplified in merged_sf (done during withProgress)

            # update reactives used by the rest of the app
            full_data(final_sf)
            full_data_reset(final_sf)
            asu_summary(
              final_sf |>
                sf::st_drop_geometry() |>
                dplyr::filter(asunum > 0) |>
                dplyr::group_by(asunum) |>
                dplyr::summarise(
                  Tracts              = dplyr::n(),
                  Population          = sum(tract_pop_cur,   na.rm = TRUE),
                  `Labor Force`       = sum(tract_ASU_clf,   na.rm = TRUE),
                  Unemployment        = sum(tract_ASU_unemp, na.rm = TRUE),
                  `Unemployment Rate` = round((Unemployment/`Labor Force`)*100, 5),
                  .groups = "drop"
                )
            )
            output$asu <- shiny::renderTable(asu_summary(), digits = 3)
            update_map_efficiently("initial_map")
            update_map_efficiently("edit_map")
            append_cpsat_log("[run] Solver finished OK; map/table updated.")
          } else {
            append_cpsat_log("[run][warn] Output JSON invalid or length mismatch.")
          }
        } else {
          append_cpsat_log("[run][error] Solver failed or out.json missing.")
        }

        lf <- cpsat_log_file()
        if (!is.null(lf) && file.exists(lf)) cpsat_log_saved(lf)
        cpsat_log_file(NULL)
        sf <- cpsat_stop_file()
        if (!is.null(sf)) unlink(sf)
        cpsat_stop_file(NULL)
        cpsat_stop_requested(FALSE)
        kf <- cpsat_skip_file()
        if (!is.null(kf)) unlink(kf)
        cpsat_skip_file(NULL)
        cpsat_running(FALSE)
      }
    })
  })

  # stop the parent observer here so non-CP-SAT code below doesn’t run
  return(invisible(NULL))
}
# --- end CP-SAT branch --------------------------------------------------

  ## 1 · orig/mine: wrap the long job in a single withProgress ------------
  shiny::withProgress(message = "Running ASU initialisation\u2026", value = 0, {
    ## 1a · Build tracts + neighbours (can take ~10\u201360 s the first time)
    shiny::incProgress(0.10, detail = "Downloading Census shapefiles\u2026")

    # use selected_year (handles New England override)
    # Map over each state to fetch tracts, bind into one map, then calculate contiguity
    tract_list <- purrr::map(state(), \(st) tigris::tracts(state = st, year = selected_year)) |> 
      dplyr::bind_rows() |>
      dplyr::mutate(
        row_num    = dplyr::row_number(),
        continuous = sfdep::st_contiguity(geometry)
      )

    # Build and cache island-fixed neighbor graph for contiguity checks.
    nb_adjacency(list(
      geoids = tract_list$GEOID,
      nb     = fix_nb_islands(
        lapply(tract_list$continuous, function(x)
          if (length(x) == 1L && x == 0L) integer(0L) else as.integer(x)),
        as.numeric(tract_list$INTPTLAT),
        as.numeric(tract_list$INTPTLON)
      )
    ))

    ## 1b · Merge the BLS data (kept for potential checks/debug)
    shiny::incProgress(0.3, detail = "Merging BLS unemployment data\u2026")
    data_merge <- tract_list |>
      dplyr::left_join(uploaded_data(), by = "GEOID")

    # (kept for backward-compat, even if not used directly)
    join_flag <- isTRUE(input$join_touching)

    shiny::incProgress(0.05, detail = "Finding initial ASU seeds…")
    if (identical(algo_choice, "orig")) {
      res <- run_asu_original(tract_list, uploaded_data())
    } else {
      shiny::incProgress(0.3, detail = "Seeding & expanding\u2026")
      th_tract_list(tract_list)
      th_bls_df(uploaded_data())
      st <- tract_hunter_seed(th_tract_list(), th_bls_df(), verbose = TRUE)
      th_state(st)
      res <- tract_hunter_finalize(th_state())
    }

    ## 3 · Finish the bar
    shiny::incProgress(1, detail = "Finishing up…")
  }) # ---- withProgress ends here  -------------------------------------

  ## 4 · Store results & update UI ---------------------------------------
  if (!is.null(res)) {
    # These slots are expected from run_asu_original()/tract_hunter_finalize()
    full_data(res$full_data)
    full_data_reset(res$full_data_reset)
    asu_data(res$asu_data)
    asu_tracts(res$asu_tracts)
    asu_summary(res$asu_summary)

    output$asu <- shiny::renderTable(asu_summary(), digits = 3)
    # The global renderMaplibre for output$edit_map will react to full_data()
  }
})


shiny::observeEvent(input$th_pass, {
  shiny::req(th_state())
  shiny::withProgress(message = "Running ASU pass…", value = 0, {
    st <- tract_hunter_asu_pass(th_state(), verbose = TRUE)
  })
  th_state(st)
  res <- tract_hunter_finalize(st)
  full_data(res$full_data)
  full_data_reset(res$full_data_reset)
  asu_data(res$asu_data)
  asu_tracts(res$asu_tracts)
  asu_summary(res$asu_summary)
  output$asu <- shiny::renderTable(asu_summary(), digits = 3)
  
  # Efficient update instead of clearing layers
  update_map_efficiently("initial_map")
  update_map_efficiently("edit_map")
})

shiny::observeEvent(input$th_combine, {
  shiny::req(th_state())
  shiny::withProgress(message = "Combining ASUs…", value = 0, {
    st <- tract_hunter_combine_groups(th_state())
  })
  th_state(st)
  res <- tract_hunter_finalize(st)
  full_data(res$full_data)
  full_data_reset(res$full_data_reset)
  asu_data(res$asu_data)
  asu_tracts(res$asu_tracts)
  asu_summary(res$asu_summary)
  output$asu <- shiny::renderTable(asu_summary(), digits = 3)
  
  # Efficient update instead of clearing layers
  update_map_efficiently("initial_map")
  update_map_efficiently("edit_map")
})

#-------Total Unemployment Box --------------------------------
total_unemp <- shiny::reactive({
  shiny::req(full_data())                                  # wait until available
  full_data() |>
    sf::st_drop_geometry() |>
    dplyr::filter(asunum > 0) |>
    dplyr::summarise(unemp = sum(tract_ASU_unemp, na.rm = TRUE)) |>
    dplyr::pull(unemp)
})

output$total_unemp <- shiny::renderUI({
  shiny::HTML(
    paste0("<strong>Total Unemployment in ASUs:</strong> ",
           scales::comma(total_unemp()))
  )
})

output$total_unemp2 <- shiny::renderUI({
  shiny::HTML(
    paste0("<strong>Total Unemployment in ASUs:</strong> ",
           scales::comma(total_unemp()))
  )
})
```


```{r}
# In your setup chunk, ensure this is working correctly:
output$log_overlay <- shiny::renderUI({
  if (!isTRUE(cpsat_running())) return(NULL)
  
  shiny::absolutePanel(
    id = "cpsat_log_panel",
    top = 80, right = 20, width = 800, height = "70vh", draggable = TRUE,
    style = "background:#fff; border:1px solid #ddd; box-shadow:0 2px 8px rgba(0,0,0,.15); overflow-y:auto; padding:10px; z-index:1000;",
    shiny::div(
      style = "font-family:monospace; font-size:12px; white-space:pre-wrap;",
      shiny::strong("CP-SAT log"), shiny::br(), 
      shiny::verbatimTextOutput("cpsat_log_tail", placeholder = FALSE)
    ),
    shiny::actionButton(
      "stop_solve", if (isTRUE(cpsat_stop_requested())) "Stopping…" else "Stop Solve",
      class = "btn-sm btn-warning", style = "margin-right:4px;"
    ),
    shiny::actionButton(
      "skip_asu", "Skip to Next ASU",
      class = "btn-sm btn-info", style = "margin-right:4px;"
    ),
    shiny::actionButton("close_log", "Close", class = "btn-sm")
  )
})

# Signal the running Python solve to stop via its CP-SAT stop_search() callback
# and return its current incumbent, instead of killing the process outright.
shiny::observeEvent(input$stop_solve, {
  shiny::req(!isTRUE(cpsat_stop_requested()))
  sf <- cpsat_stop_file()
  shiny::req(!is.null(sf), isTRUE(cpsat_running()))
  if (!file.exists(sf)) file.create(sf)
  cpsat_stop_requested(TRUE)
  append_cpsat_log("[run] Stop requested by user; waiting for solver to return its current incumbent…")
})

# Halt only the ASU currently being solved (via stop_search()) and let the
# builder move on to the next ASU window, instead of stopping the whole run.
shiny::observeEvent(input$skip_asu, {
  kf <- cpsat_skip_file()
  shiny::req(!is.null(kf), isTRUE(cpsat_running()))
  if (!file.exists(kf)) file.create(kf)
  append_cpsat_log("[run] Skip requested by user; committing current ASU's incumbent and moving to the next one…")
})

# Add this observer to handle closing the log manually
shiny::observeEvent(input$close_log, {
  cpsat_running(FALSE)
})

output$cpsat_log_link <- shiny::renderUI({
  lf <- cpsat_log_saved()
  if (is.null(lf)) return(NULL)
  shiny::tagList(
    shiny::hr(style = "margin: 0.4rem 0;"),
    shiny::div(style = "font-size:11px; color:#555; margin-bottom:4px;", "Last log:"),
    shiny::div(style = "font-family:monospace; font-size:10px; word-break:break-all; color:#333; margin-bottom:6px;", lf),
    shiny::div(
      shiny::actionButton("open_log_folder2", "Open folder", class = "btn-xs btn-info", style = "margin-right:4px;"),
      shiny::actionButton("open_log_file2",   "Open log",    class = "btn-xs btn-default")
    )
  )
})

shiny::observeEvent(input$open_log_folder2, {
  lf <- cpsat_log_saved()
  if (!is.null(lf)) try(shell.exec(dirname(normalizePath(lf, mustWork = FALSE))), silent = TRUE)
})
shiny::observeEvent(input$open_log_file2, {
  lf <- cpsat_log_saved()
  if (!is.null(lf) && file.exists(lf)) try(shell.exec(normalizePath(lf)), silent = TRUE)
})

# Make sure the log text stays updated

```

```{r}
# Add debounced reactive for map updates
map_update_trigger <- shiny::debounce(shiny::reactive({
  list(full_data(), input$update)
}), 200)

# Observer for debounced updates — fires when full_data() or input$update changes.
# paint_only=TRUE skips set_source so no GeoJSON is shipped on manual edits;
# algorithm completions use explicit update_map_efficiently() calls with the default.
shiny::observeEvent(map_update_trigger(), {
  if (!is.null(full_data()) && nrow(full_data()) > 0) {
    update_map_efficiently("edit_map", paint_only = TRUE)
  }
}, ignoreInit = TRUE)
```

```{r}
# Read-only overview map for Load Initial ASU tab; keyed on geoid_set() so it
# only re-renders when the tract set changes, not on every asunum edit.
output$initial_map <- mapgl::renderMaplibre({
  shiny::req(length(geoid_set()) > 0)
  map_data <- shiny::isolate(map_geom())
  if (is.null(map_data) || nrow(map_data) == 0) {
    return(
      mapgl::maplibre(style = mapgl::carto_style("positron")) |>
      mapgl::add_control("No valid polygon geometry to display.", "top-left")
    )
  }

  tile_source <- ensure_map_tiles(map_data)
  map_data <- map_data |> dplyr::mutate(asunum = dplyr::na_if(asunum, 0L))

  colors <- calculate_tract_colors(map_data)

  map <- mapgl::maplibre(style = mapgl::carto_style("positron")) |>
    mapgl::fit_bounds(map_data, animate = FALSE)
  layer_source <- map_data
  source_layer <- NULL
  if (!is.null(tile_source)) {
    map <- map |>
      mapgl::add_pmtiles_source(
        id = tile_source$source_id,
        url = tile_source$url,
        promote_id = "GEOID"
      )
    layer_source <- tile_source$source_id
    source_layer <- tile_source$source_layer
  }

  map |>
    mapgl::add_fill_layer(
      id                 = "basemap",
      source             = layer_source,
      source_layer       = source_layer,
      fill_color         = mapgl::match_expr(
        column = "GEOID",
        values = colors$values,
        stops = colors$stops,
        default = "#E0E0E0"
      ),
      fill_opacity       = 0.5,
      fill_outline_color = "black",
      tooltip = tract_tooltip()
    )
}) |>
  shiny::bindCache(geoid_set(), cache = "session")
```


## Map {data-width="800"}
```{r}
mapgl::maplibreOutput("initial_map", height = "100vh")

# Log overlay (shows only while running)
shiny::uiOutput("log_overlay")

```


# Modify ASU Selections

## Controls {data-width="200" style="height: 100vh; overflow-y: auto; padding-right: 10px;"}

These tools allow you to select, inspect, and change ASU designations for individual Census tracts.  *Selected* Census tracts will be outlined in red, and are the tracts updated when you change the ASU number.  You may also click on the map to select or unselect individual tracts. An ASU value of 0 means the tract is not included in any ASU.

```{r}
#-------Total Unemployment Box --------------------------------
shiny::htmlOutput("total_unemp2")
shiny::renderUI(shiny::HTML("<hr>"))

# Select an ASU — render once, update choices in-place to avoid full widget rebuild
shiny::selectInput("select_asunum", "Select ASU Number:",
              choices = "None", selected = "None")
shiny::renderUI(shiny::HTML("<hr>"))

# Change ASU
shiny::numericInput("new_asu", "Change ASU Value:", value = 0, min = 0)
shiny::actionButton("update", "Update Selected Tracts")
shiny::renderUI(shiny::HTML("<hr>"))

# Clear Selected Tracts
shiny::actionButton("clear", "Clear Selected Tracts")
shiny::renderUI(shiny::HTML("<hr>"))

asu_choices <- shiny::reactive({
  shiny::req(full_data())
  full_data() |> dplyr::pull(asunum) |> unique() |> sort()
}) |> shiny::debounce(400)

# Push choice updates in-place; 400 ms debounce absorbs rapid sequential edits.
shiny::observe({
  shiny::updateSelectInput(session, "select_asunum",
                           choices = c("None", asu_choices()))
}) |> shiny::bindEvent(asu_choices(), ignoreNULL = FALSE, ignoreInit = FALSE)

shiny::renderUI(shiny::HTML("<hr>"))

# Show Lowest UR Tracts in Selection
shiny::selectInput("unemployment_filter", 
             "Highlight selected tracts with:",
             choices = list("Low unemployment" = "low", 
                           "High unemployment" = "high"),
             selected = "low")
shiny::numericInput("percentile", "Percentile Threshold:", value = 5, min = 1, max = 100, step = 1)
shiny::actionButton("filter_percentile", "Highlight Tracts")
shiny::actionButton("clear_percentile", "Clear Highlights")
shiny::renderUI(shiny::HTML("<hr>"))

# Single GROID Selection
shiny::textInput("manual_geoid", "Enter a GEOID:", placeholder = "32031990000")
shiny::actionButton("select_geoid", "Select a Single Tract")
shiny::renderUI(shiny::HTML("<hr>"))

# Reset to Initial Load
shiny::actionButton("reset", "Reset to Initial Data")
```

## Interactive Map {data-width="600"}

```{r}
# The map
  mapgl::maplibreOutput("edit_map", height = "100vh")
  
output$edit_map <- mapgl::renderMaplibre({

  shiny::req(length(geoid_set()) > 0)
  # isolate() prevents reactive dependency on map_geom()/full_data() so that
  # asunum edits do not invalidate this output and cause MapLibre to reinitialise.
  data <- shiny::isolate(map_geom())
  if (is.null(data) || nrow(data) == 0) return(NULL)

  colors <- calculate_tract_colors(data)
  tile_source <- ensure_map_tiles(data)
  map <- mapgl::maplibre(style = mapgl::carto_style("positron")) |>
    mapgl::fit_bounds(data, animate = FALSE)
  layer_source <- data
  source_layer <- NULL
  if (!is.null(tile_source)) {
    map <- map |>
      mapgl::add_pmtiles_source(
        id = tile_source$source_id,
        url = tile_source$url,
        promote_id = "GEOID"
      )
    layer_source <- tile_source$source_id
    source_layer <- tile_source$source_layer
  }

  map |>
    mapgl::add_fill_layer(
      id = "basemap",
      source = layer_source,
      source_layer = source_layer,
      fill_color = mapgl::match_expr(
        column = "GEOID",
        values = colors$values,
        stops = colors$stops,
        default = "#E0E0E0"
      ),
      fill_opacity = 0.5,
      fill_outline_color = "black",
      tooltip = tract_tooltip()
    ) |>
    # Pre-add hidden overlay layers; visibility="none" is the off-state so that
    # mapgl 0.5.0's filter-stack (base+user composed via ["all",...]) is not
    # poisoned by a permanent "__NONE__" base filter.
    mapgl::add_line_layer(
      id = "selected",
      source = layer_source,
      source_layer = source_layer,
      line_color = "red",
      line_width = 4,
      line_opacity = 1,
      visibility = "none"
    ) |>
    mapgl::add_fill_layer(
      id = "highlighted",
      source = layer_source,
      source_layer = source_layer,
      fill_color = "black",
      fill_opacity = 0.8,
      visibility = "none"
    )
}) |>
  shiny::bindCache(geoid_set(), cache = "session")

# OPTIMIZATION 1: Use set_filter instead of clear/add for selection highlighting
shiny::observeEvent(input$edit_map_feature_click, {
  click <- input$edit_map_feature_click
  current_selection <- selected_tracts()

  clicked_geoid <- click$properties$GEOID
  if (is.null(clicked_geoid) || length(clicked_geoid) == 0 || is.na(clicked_geoid)) return()

  # toggle selection
  if (clicked_geoid %in% current_selection) {
    current_selection <- setdiff(current_selection, clicked_geoid)
  } else {
    current_selection <- c(current_selection, clicked_geoid)
  }
  selected_tracts(current_selection)

  # Build filter expression and apply (with visibility toggle)
  apply_overlay_filter(
    mapgl::maplibre_proxy("edit_map"),
    "selected",
    current_selection
  )

  output$selection_table <- shiny::renderTable({ selected_tracts() }, digits = 3)
})


# OPTIMIZATION 2: Debounce rapid interactions
clear_selection_debounced <- shiny::debounce(shiny::reactive({
  input$clear
}), 100)

shiny::observeEvent(clear_selection_debounced(), {
  selected_tracts(NULL)
  
  # Use apply_overlay_filter to hide selection layer
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
  
  # Reset ASU dropdown to "None"
    shiny::updateSelectInput(session, "select_asunum", selected = "None")
})

# OPTIMIZATION 3: Efficient ASU selection with set_filter
shiny::observeEvent(input$select_asunum, {
  shiny::req(map_geom())

  if (input$select_asunum == "None") {
    selected_tracts(NULL)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
    return()
  }

  selected_tract_ids <- map_geom() |>
    dplyr::filter(asunum == as.numeric(input$select_asunum)) |>
    dplyr::pull(GEOID)

  if (length(selected_tract_ids) == 0) {
    selected_tracts(NULL)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "selected", character(0))
    return()
  }

  selected_tracts(selected_tract_ids)

  apply_overlay_filter(
    mapgl::maplibre_proxy("edit_map"),
    "selected",
    selected_tract_ids
  ) |>
    mapgl::fit_bounds(subset(full_data(), GEOID %in% selected_tract_ids), animate = TRUE)
})


# Manual GEOID selection with set_filter
shiny::observeEvent(input$select_geoid, {
  shiny::req(map_geom())

  selected_tract <- map_geom() |>
    dplyr::filter(GEOID == input$manual_geoid)

  if (nrow(selected_tract) > 0) {
    shiny::updateSelectInput(session, "select_asunum", selected = "None")

    apply_overlay_filter(
      mapgl::maplibre_proxy("edit_map"),
      "selected",
      selected_tract$GEOID
    ) |>
      mapgl::fit_bounds(selected_tract, animate = TRUE)

    selected_tracts(selected_tract$GEOID)
  }
})


# OPTIMIZATION 4: Memoize expensive calculations
# Memoized filter of selected tracts by percentile
filtered_tracts_memo <- shiny::reactive({
  shiny::req(input$percentile, input$unemployment_filter, full_data(), selected_tracts())

  potential_highlights <- full_data() |>
    dplyr::filter(GEOID %in% selected_tracts(),
                  tract_pop_cur > 0, tract_ASU_clf > 0)

  if (nrow(potential_highlights) == 0) return(character(0))

  if (input$unemployment_filter == "low") {
    threshold <- stats::quantile(potential_highlights$tract_ASU_urate, input$percentile / 100)
    potential_highlights |>
      dplyr::filter(tract_ASU_urate <= threshold) |>
      dplyr::pull(GEOID)
  } else {
    threshold <- stats::quantile(potential_highlights$tract_ASU_urate, 1 - (input$percentile / 100))
    potential_highlights |>
      dplyr::filter(tract_ASU_urate >= threshold) |>
      dplyr::pull(GEOID)
  }
}) |>
  shiny::bindCache(input$percentile, input$unemployment_filter, selected_tracts(), full_data()$asunum)

# Apply highlights only when the button is explicitly clicked
shiny::observe({
  if (is.null(selected_tracts()) || length(selected_tracts()) == 0) {
    shiny::showNotification("Select tracts on the map or via the ASU dropdown first.",
                            type = "warning", duration = 4)
    return()
  }
  filtered_selected <- filtered_tracts_memo()
  highlighted_tracts(filtered_selected)

  if (length(filtered_selected) > 0 && !any(is.na(filtered_selected))) {
    apply_overlay_filter(
      mapgl::maplibre_proxy("edit_map"),
      "highlighted",
      filtered_selected
    )
  } else {
    shiny::showNotification("No tracts matched the filter criteria.", type = "message", duration = 3)
    apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", character(0))
  }
}) |> shiny::bindEvent(input$filter_percentile, ignoreInit = TRUE)

# Clear highlights on demand; do NOT clear on select_asunum change
shiny::observeEvent(input$clear_percentile, {
  highlighted_tracts(character(0))
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", character(0))
}, ignoreInit = TRUE)




# Update with dynamic color changes
shiny::observeEvent(input$update, {
  shiny::req(selected_tracts(), full_data())
  old_data <- full_data()
  new_asu <- as.numeric(input$new_asu)
  
  # Batch update for better performance
  updated_data <- old_data |>
    dplyr::mutate(
      asunum = dplyr::case_when(
        GEOID %in% selected_tracts() ~ new_asu,
        TRUE ~ asunum
      )
    )
  full_data(updated_data)
  selected_tracts(NULL)

  # Paint-only update fires immediately so colors change without waiting for the
  # 200 ms debounce. GeoJSON sources refresh afterward; PMTiles remain immutable.
  cm <- calculate_tract_colors(map_geom())
  proxy <- mapgl::maplibre_proxy("edit_map") |>
    mapgl::set_paint_property(
      "basemap", "fill-color",
      mapgl::match_expr(
        column  = "GEOID",
        values  = cm$values,
        stops   = cm$stops,
        default = "#E0E0E0"
      )
    )
  apply_overlay_filter(proxy, "selected", character(0))
})


refresh_edit_map <- function() {
  shiny::req(map_geom())
  data <- map_geom()
  
  # Use the efficient update function that handles data changes
  update_map_with_data("edit_map", data) |>
    mapgl::fit_bounds(data, animate = TRUE)
}


shiny::observeEvent(input$reset, {
  shiny::req(full_data_reset())
  full_data(full_data_reset())
  selected_tracts(NULL)
  highlighted_tracts(character(0))
  shiny::updateSelectInput(session, "select_asunum", selected = "None")
  
  # Use the efficient refresh function
  refresh_edit_map()
})

# OPTIMIZATION 8: Optimized data loading observer
# Observer for loading the data
shiny::observeEvent(input$load_data, {
  shiny::req(input$load_data)
  
  # Read the RDS file into a data frame
  loaded_data <- tryCatch({
    readRDS(input$load_data$datapath)
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Loading Data",
      "Failed to load the data. Please check the file format.",
      easyClose = TRUE
    ))
    return(NULL)
  })
  
  # If loaded_data is NULL (failed to load), exit the observer
  shiny::req(!is.null(loaded_data))
  
  # Ensure the loaded data is an sf object
  if (!inherits(loaded_data, "sf")) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Loading Data",
      "The loaded data is not a valid sf object.",
      easyClose = TRUE
    ))
    return(NULL)
  }
  
  # Update full_data with the loaded data
  full_data(loaded_data)
  
  # Provide feedback to the user
  shiny::showModal(shiny::modalDialog(
    title = "Data Loaded",
    "Your data has been successfully loaded.",
    easyClose = TRUE
  ))
  
  # Use the efficient update function for new data
  update_map_with_data("edit_map", loaded_data) |>
    mapgl::fit_bounds(loaded_data, animate = TRUE)
  
  # Clear any existing selections/highlights
  selected_tracts(NULL)
  highlighted_tracts(NULL)
})

# Trigger map resize when panels are resized
shiny::observeEvent(input$map_resize, {
  # Force map to resize/redraw
  mapgl::maplibre_proxy("edit_map") |>
    mapgl::maplibre_resize()
}, ignoreInit = TRUE)
```

## Selection Summary {data-width="200"}

```{r}
preview_data <- shiny::reactive({
  shiny::req(selected_tracts())
  replacement_data <- full_data() |>
    sf::st_drop_geometry() |>
    subset(GEOID %in% selected_tracts())
  summary_data <- replacement_data |>
    dplyr::ungroup() |>
    dplyr::summarize(
      tracts = dplyr::n(),
      population = sum(tract_pop_cur, na.rm = TRUE),
      lf = sum(tract_ASU_clf, na.rm = TRUE),
      unemp = sum(tract_ASU_unemp, na.rm = TRUE),
      ur = round(unemp/lf*100, 5)
    )
  

  list(data = replacement_data, summary = summary_data)
})

# Conditional display based on selection
shiny::conditionalPanel(
  condition = "output.has_selection == true",
  style = "height: calc(100vh - 20px); display: flex; flex-direction: column;",
  
  shiny::h3(" Summary of Selected Tracts"),
  shiny::hr(),

  # Summary statistics
  shiny::div(
    style = "background-color: #f8f9fa; padding: 15px; border-radius: 5px; margin-bottom: 15px; flex-shrink: 0;",
    shiny::textOutput("summary_count"),
    shiny::br(),
    shiny::textOutput("summary_total_pop"),
    shiny::textOutput("summary_total_labor_force"),
    shiny::textOutput("summary_avg_unemployment_rate"),
    shiny::textOutput("summary_total_unemployment")
  ),
  
  shiny::hr(style = "flex-shrink: 0;"),
  
  # Detailed table
  shiny::h4("Tract Details", style = "flex-shrink: 0;"),
  shiny::div(
    style = "flex: 1; overflow-y: auto; border: 1px solid #dee2e6; border-radius: 5px; min-height: 400px;",
    DT::dataTableOutput("summary_table")
  )
)

# No selection message
shiny::conditionalPanel(
  condition = "output.has_selection == false",
  shiny::div(
    style = "text-align: center; color: #6c757d; font-style: italic; padding: 50px 20px;",
    shiny::h4("No tracts currently selected"),
    shiny::p("Select tracts on the map or use the controls to view summary data here.")
  )
)

# Server logic for conditional display
output$has_selection <- shiny::reactive({
  length(selected_tracts()) > 0
})
shiny::outputOptions(output, "has_selection", suspendWhenHidden = FALSE)

# Summary statistics outputs using your existing preview_data reactive
output$summary_count <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Number of selected tracts:", preview_result$summary$tracts)
})

output$summary_total_pop <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total population:", format(preview_result$summary$population, big.mark = ","))
})

output$summary_total_labor_force <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total labor force:", format(preview_result$summary$lf, big.mark = ","))
})

output$summary_avg_unemployment_rate <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Average unemployment rate:", preview_result$summary$ur, "%")
})

output$summary_total_unemployment <- shiny::renderText({
  shiny::req(selected_tracts())
  preview_result <- preview_data()
  paste("Total unemployment:", format(preview_result$summary$unemp, big.mark = ","))
})

# Detailed table — interactive DT with bidirectional map-highlight sync
output$summary_table <- DT::renderDataTable({
  shiny::req(selected_tracts())

  preview_data()$data |>
    dplyr::select(GEOID, asunum, tract_pop_cur, tract_ASU_clf,
                  tract_ASU_urate, tract_ASU_unemp) |>
    dplyr::arrange(GEOID) |>
    dplyr::rename(
      ASU             = asunum,
      Population      = tract_pop_cur,
      `Labor Force`   = tract_ASU_clf,
      `Unemp. Rate %` = tract_ASU_urate,
      `Unemp.`        = tract_ASU_unemp
    ) |>
    DT::datatable(
      selection  = "multiple",
      rownames   = FALSE,
      options    = list(
        pageLength = 25,
        scrollX    = TRUE,
        dom        = "ftp",
        columnDefs = list(list(className = "dt-right", targets = c(1, 2, 3, 4, 5)))
      )
    ) |>
    DT::formatRound(columns = "Unemp. Rate %", digits = 2) |>
    DT::formatCurrency(columns = c("Population", "Labor Force", "Unemp."),
                       currency = "", digits = 0)
}, server = FALSE)

# DT row selected → highlight those tracts on the map
shiny::observeEvent(input$summary_table_rows_selected, ignoreNULL = FALSE, {
  shiny::req(selected_tracts(), preview_data())
  sel_rows  <- input$summary_table_rows_selected
  tbl_geoids <- preview_data()$data |> dplyr::arrange(GEOID) |> dplyr::pull(GEOID)
  geoids <- if (length(sel_rows) > 0) tbl_geoids[sel_rows] else character(0)
  highlighted_tracts(geoids)
  filter_expr <- if (length(geoids) == 0) {
    NULL
  } else if (length(geoids) == 1) {
    list("==", list("get", "GEOID"), as.character(geoids))
  } else {
    list("in", list("get", "GEOID"), list("literal", as.character(geoids)))
  }
  apply_overlay_filter(mapgl::maplibre_proxy("edit_map"), "highlighted", geoids)
})

# highlighted_tracts() changed externally (percentile filter / map) → reflect in DT
shiny::observe({
  hl <- highlighted_tracts()
  shiny::req(selected_tracts(), preview_data())
  tbl_geoids  <- preview_data()$data |> dplyr::arrange(GEOID) |> dplyr::pull(GEOID)
  row_indices <- which(tbl_geoids %in% hl)
  if (!identical(as.integer(shiny::isolate(input$summary_table_rows_selected)),
                 as.integer(row_indices))) {
    DT::dataTableProxy("summary_table") |> DT::selectRows(row_indices)
  }
}) |> shiny::bindEvent(highlighted_tracts(), ignoreInit = TRUE)
```

# Save and Load Data

## Controls {data-width="150"}

These buttons allow you to save the map in progress so that you can load
it again in the future to continue your work.

First, choose a save directory. The default is your current working
directory. After loading data, you can go directly to the ASU
Modification tab to continue making edits.

```{r}
# Directory selection - Simple text input approach
shiny::textInput("save_dir_path", "Save Directory Path:", value = getwd(), width = "100%")
#shiny::actionButton("browse_dir", "Browse for Directory") # non-functional
shiny::textOutput("current_save_dir")
shiny::textOutput("dir_status")
shiny::br()

shiny::actionButton("save_data", "Save Data")
shiny::fileInput("load_data", "Load Data", accept = c(".rds"))

# Reactive value to store the selected directory
save_directory <- shiny::reactiveVal(getwd())  # Default to current working directory

# Observer to handle directory path changes
shiny::observeEvent(input$save_dir_path, {
  if (!is.null(input$save_dir_path) && input$save_dir_path != "") {
    if (dir.exists(input$save_dir_path)) {
      save_directory(input$save_dir_path)
    }
  }
})

# Display current save directory
output$current_save_dir <- shiny::renderText({
  paste("Current save directory: ", save_directory())
})

output$review_save_dir <- shiny::renderText({
  paste("Current save directory: ", save_directory())
})

# Display directory status
output$dir_status <- shiny::renderText({
  if (dir.exists(save_directory())) {
    "✓ Directory exists and is accessible"
  } else {
    "⚠ Directory does not exist - it will be created when saving"
  }
})

# Observer for saving the data
shiny::observeEvent(input$save_data, {
  shiny::req(full_data())
  
  # Use the selected directory
  save_path <- file.path(save_directory(), "saved_data.rds")
  
  # Check if directory exists, create if it doesn't
  if (!dir.exists(save_directory())) {
    dir.create(save_directory(), recursive = TRUE)
  }
  
  # Save the full_data object to an RDS file
  tryCatch({
    saveRDS(full_data(), save_path)
    
    # Provide feedback to the user
    shiny::showModal(shiny::modalDialog(
      title = "Data Saved",
      paste("Your data has been saved to:", save_path),
      easyClose = TRUE
    ))
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Saving Data",
      paste("Failed to save data:", e$message),
      easyClose = TRUE
    ))
  })
})


```

## Output {data-width="850"}

```{r}
# Empty output section
```

# ASU Review and Finalization

## Controls {data-width="150"}

This screen will allow you to review your selected ASUs. It uses a
default unemployment-rate threshold of 6.45% and a population threshold
of 10,000. Update either value here when the applicable program guidance
changes.

*Generate ASU Summary* will show a table of all ASUs, and whether they
pass or fail the unemployment, population, and contiguity tests. The
qualification calculation uses the unrounded aggregate rate, even though
the displayed rate is rounded.

Note, you *may* go back and forth between the selection and review
screens.

When ready to create an output file, the *Generate LSS .txt File* button
will create a formatted batch file for upload to LSS.

```{r}
shiny::numericInput("asu_ur", "ASU Unemployment Rate Threshold:", value = 6.45, min = 0)

shiny::numericInput("asu_pop", "ASU Population Threshold:", value = 10000, min = 0)

shiny::actionButton("run_asu_summary", "Generate ASU Summary")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::textOutput("review_save_dir")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::actionButton("lss_txt", "Generate LSS .txt File")
shiny::renderUI(shiny::HTML("<hr>"))

shiny::actionButton("review_txt", "Create Summary CSV")
```

## Output {data-width="850"}

```{r}
DT::dataTableOutput("asu_review")

summarized_asu <- shiny::reactiveVal(NULL)

shiny::observeEvent(input$run_asu_summary, {

  shiny::req(full_data())

  asu_sf <- full_data() |> dplyr::filter(asunum > 0.5)

  # Contiguity check: use the island-fixed full-state neighbor graph built at
  # load time so that island tracts connected via centroid distance are handled
  # consistently with how the solver built the ASUs.
  asu_ids <- sort(unique(asu_sf$asunum))

  nb_check <- nb_adjacency()
  if (is.null(nb_check)) {
    # Fallback for sessions loaded from RDS: rebuild from current geometry.
    all_sf  <- full_data()
    raw_nb  <- lapply(
      suppressWarnings(sfdep::st_contiguity(all_sf$geometry)),
      function(x) if (length(x) == 1L && x == 0L) integer(0L) else as.integer(x)
    )
    fixed_nb <- if (all(c("INTPTLAT", "INTPTLON") %in% names(all_sf)))
      fix_nb_islands(raw_nb, as.numeric(all_sf$INTPTLAT), as.numeric(all_sf$INTPTLON))
    else raw_nb
    nb_check <- list(geoids = all_sf$GEOID, nb = fixed_nb)
    nb_adjacency(nb_check)
  }

  contiguous_flag <- vapply(asu_ids, function(id) {
    grp_geoids <- asu_sf |> dplyr::filter(asunum == id) |> dplyr::pull(GEOID)
    asu_is_contiguous(grp_geoids, nb_check)
  }, logical(1L))

  contiguity_df <- data.frame(asunum = asu_ids, contiguous = contiguous_flag)

  summary_table <- asu_sf |>
    sf::st_drop_geometry() |>
    dplyr::group_by(asunum) |>
    dplyr::summarize(
      Tracts = dplyr::n(),
      Population = sum(tract_pop_cur, na.rm = TRUE),
      `Labor Force` = sum(tract_ASU_clf, na.rm = TRUE),
      Unemployment = sum(tract_ASU_unemp, na.rm = TRUE),
      `Unemployment Rate` = round((Unemployment / `Labor Force`) * 100, 5),
      .groups = "drop"
    ) |>
    dplyr::left_join(contiguity_df, by = "asunum") |>
    dplyr::mutate(
      ur_qualified   = dplyr::if_else(`Unemployment Rate` >= input$asu_ur, TRUE, FALSE),
      pop_qualified  = dplyr::if_else(Population >= input$asu_pop, TRUE, FALSE),
      contiguous     = contiguous,
      asu_qualified  = ur_qualified & pop_qualified & contiguous
    )

  summarized_asu(summary_table)
})

output$asu_review <- DT::renderDataTable({
  tbl <- summarized_asu()
  shiny::req(tbl)

  tbl |>
    dplyr::mutate(
      ur_qualified  = ifelse(ur_qualified,  "TRUE", "FALSE"),
      pop_qualified = ifelse(pop_qualified, "TRUE", "FALSE"),
      contiguous    = ifelse(contiguous,    "TRUE", "FALSE"),
      asu_qualified = ifelse(asu_qualified, "TRUE", "FALSE")
    ) |>
    DT::datatable(
      rownames  = FALSE,
      selection = "none",
      
      options   = list(
        scrollX   = TRUE,
        pageLength = 25,
        dom        = "ftp",
        order      = list(list(0, "asc"))
      )
    ) |>
    DT::formatRound(columns = "Unemployment Rate", digits = 3) |>
    DT::formatCurrency(
      columns  = c("Population", "Labor Force", "Unemployment"),
      currency = "", digits = 0
    ) |>
    DT::formatStyle(
      columns     = c("ur_qualified", "pop_qualified", "contiguous", "asu_qualified"),
      color       = DT::styleEqual(c("TRUE", "FALSE"), c("#2e7d32", "#c62828")),
      fontWeight  = "bold"
    )
}, server = FALSE)

shiny::observeEvent(input$lss_txt, {
  
  shiny::req(full_data())

  # Function to generate text content
  generate_txt <- function() {
    df <- full_data() |>
      sf::st_drop_geometry() |>
      dplyr::select(GEOID, asunum) |>
      dplyr::filter(asunum > 0) |>
      dplyr::mutate(asunum = as.integer(asunum),
             GEOID = paste0("14000US", GEOID),
             asunum = paste0("SU", state(), sprintf("%04d", asunum)),
             GEOID = paste(asunum, GEOID, sep = " + ")) |> 
      dplyr::arrange(asunum, GEOID) |>
      dplyr::select(-asunum)

    txt_content <- utils::capture.output(utils::write.table(df, quote=FALSE, row.names=FALSE, col.names = FALSE))
    txt_content <- paste(txt_content, collapse="\n")
    return(txt_content)
  }

  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    shiny::showModal(shiny::modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  txt_content <- generate_txt()

  # Use the selected save directory
  filePath <- file.path(current_dir, "lss_batch_file.txt")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      shiny::showModal(shiny::modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the file
    writeLines(txt_content, filePath)
    
    # Show success message with file location
    shiny::showModal(shiny::modalDialog(
      title = "File Created Successfully",
      shiny::HTML(paste0(
        "<strong>LSS batch file has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Creating File",
      paste("Failed to create LSS file:", e$message),
      easyClose = TRUE
    ))
  })
})

shiny::observeEvent(input$review_txt, {

  shiny::req(full_data())

  # Function to generate CSV content
  generate_csv <- function() {
    df <- full_data() |>
      sf::st_drop_geometry() |>
      dplyr::arrange(asunum, GEOID)
    
    return(df)
  }
  
  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    shiny::showModal(shiny::modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  df <- generate_csv()

  # Use the selected save directory
  filePath <- file.path(current_dir, "ASU_Review_File.csv")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      shiny::showModal(shiny::modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the CSV file
    utils::write.csv(df, filePath, row.names = FALSE)
    
    # Show success message with file location
    shiny::showModal(shiny::modalDialog(
      title = "File Created Successfully",
      shiny::HTML(paste0(
        "<strong>ASU Review CSV has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    shiny::showModal(shiny::modalDialog(
      title = "Error Creating File",
      paste("Failed to create CSV file:", e$message),
      easyClose = TRUE
    ))
  })
})

####################
## Future Development

# Export Excel Worksheet with tract details and selections.
# Allow "Save" of data with txt/Excel files so single session is not required.
# Better table formatting.
# highlight high unemployment tracts in non-included areas.
# Check for contiguity of selected ASUs after manual edits.
# Export high-resolution images of ASUs for review.
# Use reactlog package to monitor reactive interactions in the code for cleanup. https://mastering-shiny.org/reactive-graph.html
```

# How to Use ASUbuildR

**About ASUbuildR 0.6.2**

ASUbuildR is an application built in R using Shiny and Flexdashboard to
create a point-and-click environment to build Areas of Substantial
Unemployment (ASUs). It helps analysts build, inspect, edit, validate,
and export tract-based ASU candidates. It does not replace current BLS
program guidance or analyst review.

The ASU creation logic in this program was created by a collaboration
between the states of Ohio and Nevada, with the Shiny app and Graphical
User Interface designed by Nevada’s Research & Analysis Bureau. The GitHub repository for this package can be found at <https://github.com/aisolori/ASUbuildR>

**About the ASU Process**

The ASU process typically begins when the U.S. Bureau of Labor
Statistics distributes to each state a file providing the calculated
inputs for regions within the state from which ASUs must be designated.
This is provided as an Excel sheet with a filename like `ST_asuYY.xlsx`
where ST is the state abbreviation and YY is the year for which the ASU
is being designated. **ASUbuildR** is currently designed to pull
information directly from this file, including identifying the state and
year for which data should be retrieved. Because this file is built using public information but using
preliminary LAUS data for June of the current year, archives of these
files are not publicly available but also do not contain confidential
information. Future versions of ASUbuildR may build these input files
directly, but currently the state-specific input file is required.

ASU candidates are evaluated on geographic contiguity, population, and
unemployment rate. The dashboard defaults to a population threshold of
10,000 and an operational unemployment-rate threshold of 6.45%, which
rounds to 6.5% at one decimal place. Both review thresholds are editable;
analysts should set them to the values required by current BLS guidance.

**Using ASUbuildR**

Install the package from GitHub with
`pak::pak("aisolori/ASUbuildR")`, then launch it with
`ASUbuildR::launch_ASUbuildR()`. Tract Hunter and Simple Snake need no
separate Python configuration. To use CP-SAT, run
`ASUbuildR::setup_asu_python()` once before launching the dashboard.

**Basic Navigation**

Within the app window, there are a series of tabs across the top. These
control the basic navigation within the app. The functionality of each
tab will be described in detail below.

-   Data Initialization: this is where you load the Excel file from BLS.

-   Load Initial ASU: here you use the data from BLS to generate an
    initial list of potential ASUs.

-   Modify ASU Selections: this is the primary window you will interact
    with. It allows for the manipulation and refinement of the data.

-   Save and Load Data: this allows you to save the current state of the
    modified data and maps to be loaded later. If you have already saved
    data, you can come straight here to load that map.

-   ASU Review and Finalization: this tab contains tools to review the
    summarized ASUs, to generate an output CSV file to review individual
    tracts, and to generate a batch TXT file that can be uploaded to
    BLS.

**Data Initialization**

To begin the ASU building process, click the `Browse` button on the left
side of the page. Once a file has been selected, an overview of the file
will appear on the right to confirm the file has successfully loaded.
You may review the contents of the columns to ensure that data appears
to be loaded correctly. This script assumes the columns are maintained
in a consistent order by BLS - changes to the structure of the file will
require updates to the structure of this code.

When review is complete, continue to the next tab - Load Initial ASU.

**Load Initial ASU**

There are three initial-building options:

-   **Tract Hunter** seeds and expands candidate groups, then lets you run
  additional Hunter and Combine Groups passes. It considers the configured
  population requirement while improving total unemployment in ASUs.

-   **Simple Snake** grows candidates tract by tract from high-unemployment
  seeds. It is a quick starting point but does not enforce the population
  qualification while building.

-   **CP-SAT (OR-Tools)** optimizes contiguous tract assignments against the
  selected unemployment and population thresholds. It requires the optional
  Python environment created by `ASUbuildR::setup_asu_python()`.

If the state being loaded is a New England state, you will see an additional option to select the year to download.  Because the Census Bureau only maintains NECTA-based census tracts for 2021 and earlier, using the default settings may cause the app to crash if your Excel file uses these older GEOIDs.

Choose an algorithm, then click **Load tracts & initialise**. The app downloads
the selected Census tract boundaries, joins the uploaded BLS data, and runs
the chosen builder. Runtime varies by state, algorithm, and CP-SAT time limit.
When processing finishes, the map and ASU summary appear on the right.

If using the Tract Hunter, use the **Run Hunter** and **Combine Groups** buttons to allow the algorithm to attempt to improve the existing ASUs.

The summary table provides aggregate population, labor force,
unemployment, and unemployment rate for each candidate ASU.

To begin manually manipulating the data, proceed to the next tab, Modify ASU
Selections. Or, if the number of modifications is large or the state
takes a long time to run, skip ahead to perform an initial save of the
data on Save and Load Data.

**Modify ASU Selections**

This tab is where all edits to the ASU selections are made. A number of
tools are provided to aid in the optimization of the ASUs. In general,
*selected* tracts will be outlined in red on the map.  A summary of selected tracts as well as aggregate statistics will be displayed on the right of the map.

-   Hover over the map. Hovering the mouse over the map will display
    information about that tract useful to the selection process.

-   Click on the map. Clicking on the map will select, or unselect, a
    particular tract.

-   Select ASU Number: This dropdown allows you to select all the tracts
    in a particular ASU. If you select None, this will effectively clear
    all selected tracts.

-   Change ASU Value / Update Selected Tracts: This is how you change
    the pre-assigned ASU number. Enter the numeric value in the box
    under *Change ASU Value:* and then click the *Update Selected
    Tracts* to change the value. The numeric value remains in the box,
    providing a convenient way to keep changing tracts or groups of
    tracts to the same ASU. Further, changing the ASU number to 0 is how
    selected tracts are removed from any ASUs.

-   Selection Summary: The summary and sortable tract table update
  automatically. Selecting rows in the table highlights those tracts on
  the map; map and percentile highlights are reflected in the table.

-   Percentile Threshold: This allows you to examine the tracts in your current selection with either high or low unemployment rates.  This is useful for identifying tracts outside of current ASUs that you may wish to include in an ASU or identifying tracts inside of current ASUs that you may with to drop.  Please choose whether you want to highlight Low or High unemployment tracts and enter a number in the Percentile Threshold box to highlight the top or bottom X% of tracts (X is the number you enter).  For example, if you have 200 tracts selected and choose 5 as your threshold and High unemployment, clicking the **Highlight Tracts** button will turn the color of the 10 tracts with the highest unemployment rates black.  Selecting **Clear Highlights** will clear all current highlighting.

-   Enter a GEOID / Select a Single Tract: This functionality is most
    useful when using the CSV output that this app generates on the
    **ASU Review and Finalization** tab. Given the 11-character GEOID,
    you can input that to highlight a single tract to examine where it
    fits. This can be helpful for fine-tuning - reviewing the
    highest-unemployment tracts not included in an ASU to ensure they
    cannot be connected in to improve the ASU.

-   Reset to Initial Data: This button will reset all changes you have
    made to the state of the map after the **Load Initial ASU** process.
    For when you’ve just broken everything and want to start over. Use
    with caution!

**Save and Load Data**

This tab contains some simple options to help save and load the current
state of the map. While Nevada had only 6 initial ASUs to review other
states may have over 100! This will save the background data of the app
as an RDS file, or load data from an RDS file.

To specify a directory other than your working directory, paste that
directory into the `Save Directory Path` box.

**ASU Review and Finalization**

This will typically be the last screen in the process. Set the applicable
unemployment-rate and population thresholds, then select **Generate ASU
Summary**. The table checks unemployment, population, and contiguity. Green
`TRUE` values pass a test; red `FALSE` values require review. Qualification
uses the unrounded aggregate unemployment rate.

Generate LSS .txt File and Create Summary CSV will both create files in
the directory specified in the **Save and Load Data** tab (or your
working directory). The LSS TXT file should be formatted to be able to
upload to LSS directly once the ASUs are completed. The Summary CSV file
provides a handy output of all selected tracts and a useful tool to
review the selections for high-unemployment tracts that were not
included.
