---
title: "Hyperparameter Tuning for Tract Hunter"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This script demonstrates how to evaluate different scoring weights in the
Tract Hunter workflow using the `tune_scoring()` helper.  Adjust the
parameter grids and the number of cores to leverage the resources on your
machine.

```{r, eval=FALSE}
library(ASUbuildR)

# objects you would normally pass to run_tract_hunter()
tract_list <- ...   # tract data with contiguity info
bls_df     <- ...   # BLS unemployment data

# define the grid of weights to try
unemp_power     <- c(0.8, 0.9, 1.0)
ur_weight       <- c(1.0, 1.2)
path_len_weight <- c(0, 0.05)

# launch the tuning with, say, all but one of your cores
res <- tune_scoring(tract_list, bls_df,
                    unemp_power     = unemp_power,
                    ur_weight       = ur_weight,
                    path_len_weight = path_len_weight,
                    cores = parallel::detectCores() - 1)

print(res)
```

The resulting data frame summarises the ASU counts for each parameter
combination, allowing you to identify effective scoring weights quickly.
